### 2023-10-22 final_paper.qmd
Added documents

---
title: "Final paper"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-10-22 assignment_1.qmd
Update 22.10

---
title: "Assignment 1: Data and Descriptive"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
editor: visual
bibliography: [Econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
---

```{r}
#| label: setup
#| output: false
#| echo: false

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
```

# Introduction

While national GDP and GDP per capita are vital indicators of a country's aggregate economic health, they do not shed light on how wealth or income is distributed among its residents. A high national GDP can, paradoxically, coexist with pockets of regional deprivation [@lessmann2017].

The truth of this statement becomes more evident when taking a closer look in to into sub-national data. Regional wealth disparities are of prime concern, especially when crafting policies for equitable growth [@lessmann2017]. A country's macro-level prosperity does not automatically guarantee that all its regions partake equally in this wealth. By studying smaller regions within a country, it is possible to get a more nuanced narrative about the state of regional economic disparities [@lessmann2017].

This assignment is anchored in this very premise. It seeks to acquire, process, and analyse sub-national GDP and population data for a selected subset of European countries, namely France, Denmark, Hungary, Portugal, and Slovakia, spanning the period from 2000 to 2020. The overarching aim is to calculate GDP per capita for these regions and delve into the intricacies of regional inequity. This exploration will be underpinned by various descriptive statistics, visualizations, and analytical techniques to paint a comprehensive picture of regional economic landscapes.

Using time series data observing the GDP per capita and GINI trends across regions and over time, we can discern whether certain areas are ahead or behind in economic performance.

# Literature review

"In recent decades, the regional distribution of incomes within countries has attracted considerable interest among academics and policy makers." [@lessmann2017]

Slovakia, experienced significant economic growth. Their GDP per capita grew faster than the older EU member states, albeit from a lower starting point, leading to a convergence [@fullart].

The global financial crisis and subsequent European debt crisis had significant impacts on European economies. GDP per capita declined in several countries, particularly those hit hardest by the crisis, such as Greece, Spain, and Portugal. Some regions took many years to recover to pre-crisis levels, while others, especially in Northern Europe, recovered more quickly [@lewis].

Major metropolitan areas, like Paris in France, London in the UK, and Berlin in Germany, often grew faster than other regions in their respective countries, leading to increasing spatial disparities [@kiuru2019].

Denmark maintained relatively high levels of GDP per capita throughout this period, benefiting from a diversified economy, strong institutions, and a high degree of economic openness. It did face challenges during the 2008 economic crisis but recovered relatively quickly [@dynamics].

Slovenia, which joined the EU in 2004, showed growth in GDP per capita, though it too was affected by the global economic downturn in 2008. However, it managed to recover in the subsequent years [@fullart].

Regional disparities persisted, with the Île-de-France region (which includes Paris) significantly outperforming other regions in terms of GDP per capita. The gap between urban and rural areas also remained a topic of discussion [@regional].

Towards the end of this period, in early 2020, the COVID-19 pandemic posed a new set of challenges for European economies, causing significant contractions in GDP per capita across many regions [@regional].

# Part A: Sub-national GDP and GDP per Capita

## Data Acquisition and datasets

The first thing we did in this assignment was to acquire our data. Trough Eurostat, we could download the datasets nama_10r_3gdp and demo_r_pjanggr3 as csv files, as well as filter the data by our preferences before downloading it. As per the assignment, we filtered the dataset by choosing the years 2000 to 2020. Furthermore, we selected the NUTS 3 region for the nations we were given, which were Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified the data to be in million Euro.

### GDP (nama_10r_3gdp)

The nama_10r_3gdp dataset from Eurostat provides insights into GDP at regional level using the NUTS classification system. It furnishes GDP values in both current prices and adjusted for inflation, with some figures given in purchasing power standards (PPS) to account for price level differences between countries. The data is often structured by year and region @eurostat2023.

The GDP at market prices represents the final result of production activities of resident producer units within a region or nation. It is calculated as the sum of the gross value added across various institutional sectors or industries, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries). This also balances out in the total economy production account. In terms of methodology, while national accounts compile GDP from the expenditure side, regional accounts don't adopt this approach due to the complexities of accurately mapping inter-regional flows of goods and services.

The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @eurostat2023c.

### Population (demo_r_pjanggr3)

Using the NUTS categorization once more, Eurostat's demo_r_pjanggr3 records annual population changes at the regional level. This dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender. It's also often displayed in a year-by-region format, and the data usually spans in yearly intervals @eurostat2023a.

Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project. The project covers 37 countries and is the central repository for demographic and migration-related data. Specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year. Additionally, data on marriages, divorces, and migration flows are recorded.

For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @eurostat2021.

### NUTS classification

The Nomenclature of Territorial Units for Statistics (NUTS) offers a stratified system to segment the economic territory of the EU (including the UK) to facilitate the consistent collection and harmonization of regional statistics across Europe. The NUTS regions range from NUTS 0 Country level to NUTS 3 small units such as municipalities level.

```{r}
#| echo: false
#| output: false

# This section of the code is dedicated to adding data to RStudio. We will import the datasets into the RStudio environment for further analysis, manipulation, and visualization using R.

# Data for population
pop_data <- read_csv("Raw_data/demo.gz")

# Data for GDP
gdp_data <- read_csv("Raw_data/nama.gz")
```

```{r}
#| echo: false
#| output: false
# Renaming and selecting the variables we want to keep
# We're doing this in order to make the datasets more clean and easy to read

gdp_data_01 <- gdp_data |>
  select(geo, year = TIME_PERIOD, GDP = OBS_VALUE)

pop_data_01 <- pop_data |>
  select(geo, year = TIME_PERIOD, population = OBS_VALUE)
```

## GDP per Capita Calculation

The formula for calculating GDP per Capita is as follows:

$y_i=GDP_i/population_i$

```{r}
#| echo: false
#| output: false

# Merging the datasets pop_data and gdp_data into one dataset by joining the variables "geo" and "year".

merged_data <- left_join(pop_data_01, gdp_data_01, join_by(geo, year))

# Select Only NUTS 3 Regions
# Filter for specific countries, this will be useful when we will find the GDP per capita for the countries seperately

merged_data <- merged_data |>
  filter(
    startsWith(as.character(geo), "PT") | 
    startsWith(as.character(geo), "FR") | 
    startsWith(as.character(geo), "HU") |
    startsWith(as.character(geo), "SK") | 
    startsWith(as.character(geo), "DK")
  ) %>%

# Filter rows where 'geo' column has more than 4 characters to select only NUTS3 regions
  filter(nchar(as.character(geo)) > 4) %>%

# Rename 'geo' column to 'id_nuts3'
  rename(id_nuts3 = geo)
```

```{r}
#| echo: false
#| output: false

# Calculating GDP per Capita

GDP_per_capita <- merged_data |>
  mutate(
    GDP_per_capita = round(100*GDP/population, digits = 2)
  )

# we use *100 in order to get the gdp per capita as a percentage, and digits = 2 to limit it down to two decimals
```

After calculating the GDP per capita for all NUTS 3 regions in our assigned countries, we can see that there is a large spread between the figures for the various regions. In this assignement we want to look at regional inequity; in order to do this in a valuable way we have to divide between the different countries. By doing this, we can gain important insights on regional differences that we can utilize, for instance, to discuss national policy on equity and sustainable economic development in regions.

### Descriptive statistics

In this part we will report and interpret different types of essential descriptive statistics. Measuring regional income inequality is challenging due to heterogeneity of regions. The number of regions in our data set varies largely in size and population. Since the focus of this paper is purely growth and changes in inequities over time, the variations of size and population density becomes a minor issue because the country-level territorial heterogeneity is fixed.

"Interest in income inequality has led to the development of several ways of measuring it. Two types of measures are of interest in this paper----static and dynamic. *Static measures* provide a snapshot of these inequalities at a point of time whereas the dynamic *measures capture historical trends.*" (læreboken)

In this part, we'll look at GDP per capita for our assigned countries on a NUTS 3 level. In addition, we'll use different kinds of descriptive statistics in order to further 0,37, is analyse this data. In this analysis, we'll use @wooldridge2020 for help.

By using figures, we can visualize the GDP per capita, and look at how it varies among the different regions. In these figures, a line represent one NUTS 3 region.

**Mean**

Calculate the mean to provide a representative value for a dataset, facilitating understanding of its central tendency and serving as a benchmark against which deviations and anomalies can be assessed, in later steps when building and interpreting regression models.

**MMR**

A comparison of the GRDP (gross regional domestic product) per capita of the region with the highest income to the region with the lowest income (minimum per capita GRDP) provides a measure of the range of these disparities. If this measure is small (close to 1), then it would mean that the different regions have relatively equal incomes. If this measure is large, then the interpretation is more problematic, as it does not tell us if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers. Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to comprehend, and politically powerful measure of regional income inequality.

**Standard diviation (SD)**

Calculating SD to quantify the dispersion or variability of a data set around its mean. Helping us assess the degree of uncertainty, variability, or risk associated with an economic variable or parameter, which is crucial for understanding the reliability of estimations and predictions [@wooldridge2020].

**Median**

The median serves as a robust measure of central tendency, especially when a dataset may have outliers or is skewed. Unlike the mean, the median is not influenced by extreme values and, thus, can provide a clearer picture of the "typical" value in situations where the data distribution is not symmetrical [@wooldridge2020].

### Portugal

```{r}
#| echo: false
#| output: false

# Making a dataset that only contains data from Portugal
# Use the na.omit function in order to remove NA values, cant make ggplot with NA values.

Portugal_data <- GDP_per_capita |>
  filter(grepl("PT", id_nuts3)) |>
  na.omit(Portugal_data)
```

```{r}
#| echo: false

# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Portugal.

Portugal_data_plot <- ggplot(Portugal_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
  geom_line() +
  labs(title="Figure 1: GDP per Capita for Portugal", 
       x="Year",
       y="GDP per Capita") +
  theme(legend.position="none")

# Print out the figure

print(Portugal_data_plot)
```

```{r}
#| echo: false

# Calculate summary statistics for Portugal, chosing only GDP per capita 
summary_stats_portugal <- sapply(Portugal_data["GDP_per_capita"], function(x) c(
  mean = mean(x),
  median = median(x),
  std_dev = sd(x),
  minimum = min(x),
  maximum = max(x)
))

# Print the summary statistics
print(summary_stats_portugal)
```

By looking at figure for Portugal, we can see that the GDP per capita in Portugal's regions appears to be fairly consistent. There is however some regional variability. We can see that the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas. Since Lisbon is the capital of Portugal, there is probably a higher concentration of industries, making it a economic center (which again makes the GDP per capita higher).

To continue, we can see that the mean is a little higher that the median, something that might indicate that regions like Lisbon are pulling up the average. If we compare the standard derivation for Portugal with the other countries, we'll see that is fairly low in comparison. This might mean that there is not a lot of variability between the GDP per capita across different regions in Portugal. The gap between minimum and maximum is also low compared to other countries, something that'll also show us that the economic disparity in Portugal might not be as high as it is in other countries.

### France

```{r}
#| echo: false
#| output: false

# Making a dataset that only contains data from France

France_data <- GDP_per_capita |>
  filter(grepl("FR", id_nuts3)) |>
  na.omit(France_data)
```

```{r}
#| echo: false

# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in France.

France_data_plot <- ggplot(France_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
  geom_line() +
  labs(title="Figure 1: GDP per Capita for France", 
       x="Year",
       y="GDP per Capita") +
  theme(legend.position="none")

print(France_data_plot)
```

```{r}
#| echo: false

# Calculate summary statistics for France, chosing only GDP per capita 
summary_stats_france <- sapply(France_data["GDP_per_capita"], function(x) c(
  mean = mean(x),
  median = median(x),
  std_dev = sd(x),
  minimum = min(x),
  maximum = max(x)
))

# Print the summary statistics
print(summary_stats_france)
```

When looking at the figure for France, we can right away see that there are some regions that have a much higher GDP per capita compared to the other regions. The region that has the highest GDP per capita for all years is the île-de-France region, one that also includes Paris. This significant difference between the regions with the highest GDP per capita and the lowest, shows us that there is a high concentration of economic activity and wealth in a few urban regions. Similar to Portugal, we can also again see that there is a difference between urban and rural regions.

Just as in Portugal, there is also a higher mean in France as well. Something that is different from the data in France compared to Portugal, is that the standard derivation is higher, and the difference between minimum and maximum is large. This strengthens what we have look at earlier in the figure, with some regions having a high concentration of wealth.

### Hungary

```{r}
#| echo: false
#| output: false

# Making a dataset that only contains data from Hungary

Hungary_data <- GDP_per_capita |>
  filter(grepl("HU", id_nuts3)) |>
  na.omit(Hungary_data)
```

```{r}
#| echo: false

# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Hungary.

Hungary_data_plot <- ggplot(Hungary_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
  geom_line() +
  labs(title="Figure 1: GDP per Capita for Hungary", 
       x="Year",
       y="GDP per Capita") +
  theme(legend.position="none")

print(Hungary_data_plot)
```

```{r}
#| echo: false

# Calculate summary statistics for Hungary, chosing only GDP per capita 
summary_stats_hungary <- sapply(Hungary_data["GDP_per_capita"], function(x) c(
  mean = mean(x),
  median = median(x),
  std_dev = sd(x),
  minimum = min(x),
  maximum = max(x)
))

# Print the summary statistics
print(summary_stats_hungary)
```

In Hungary, most of the regions have similar GDP per Capita. One region that sticks out by having a higher value, is the region of Budapest, Hungary's biggest city.

We have here as well an mean that is larger than the median, high standard derivation, and a large gap between minimum and maximum.

### Slovakia

```{r}
#| echo: false
#| output: false

# Making a dataset that only contains data from Slovakia

Slovakia_data <- GDP_per_capita |>
  filter(grepl("SK", id_nuts3)) |>
  na.omit(Slovakia_data)
```

```{r}
#| echo: false

# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Slovakia.

Slovakia_data_plot <- ggplot(Slovakia_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
  geom_line() +
  labs(title="Figure 1: GDP per Capita for Slovakia", 
       x="Year",
       y="GDP per Capita") +
  theme(legend.position="none")

print(Slovakia_data_plot)
```

```{r}
#| echo: false

# Calculate summary statistics for Slovakia, chosing only GDP per capita 
summary_stats_slovakia <- sapply(Slovakia_data["GDP_per_capita"], function(x) c(
  mean = mean(x),
  median = median(x),
  std_dev = sd(x),
  minimum = min(x),
  maximum = max(x)
))

# Print the summary statistics
print(summary_stats_slovakia)
```

In Slovakia as well, we have one region that has a much higher GDP per capita than the rest of the regions. This region is Bratislava, which is the biggest city and capital, something that might point to this city being the economic capital of Slovakia as well.

Slovakia has also a mean higher that the median, and a large gap between minimum and maximum. In addition, the standard derivation is pretty high, meaning that there is some regions (or one region in this case) that is far away from the rest of the regions when it comes to economic development.

### Denmark

```{r}
#| echo: false
#| output: false

# Making a dataset that only contains data from Denmark

Denmark_data <- GDP_per_capita |>
  filter(grepl("DK", id_nuts3)) |>
  na.omit(Denmark_data)
```

```{r}
#| echo: false

# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Denmark.

Denmark_data_plot <- ggplot(Denmark_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
  geom_line() +
  labs(title="Figure 1: GDP per Capita for Denmark", 
       x="Year",
       y="GDP per Capita") +
  theme(legend.position="none")

print(Denmark_data_plot)
```

```{r}
#| echo: false

# Calculate summary statistics for Denmark, chosing only GDP per capita 
summary_stats_denmark <- sapply(Denmark_data["GDP_per_capita"], function(x) c(
  mean = mean(x),
  median = median(x),
  std_dev = sd(x),
  minimum = min(x),
  maximum = max(x)
))

# Print the summary statistics
print(summary_stats_denmark)
```

Lastly, we have Denmark. We can see similar pattern here as well, with the capital Copenhagen being one of the regions with the highest GDP per capita.

We can also see the same as the previous countries, with the mean being higher than the median, which shows us that regions like Copenhagen might drag the mean up by being much larger than the rest of the regions.

# Part B: Regional Inequity

## Gini Coefficient Calculation

With the use of the NUTS3 GDP per capita data and this formula:

$GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|$

we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.

The gini coefficient can help us measure inequality in a distribution, as is therefore a useful tool for us to use when we look at regional inequity. The closer the gini coefficient is to 1, the bigger the inequality is; a number closer to 0 equals equality. When looking at the gini coefficient for NUTS 2 regions, we also get a better overview over differences in income between different regions, and it also makes it easier to find the reasons as to why there is a difference between the regions @hasell2023.

```{r}
#| echo: false
#| output: false

# Before we can find the Gini coefficient for NUTS2 regions, we have to generate NUTS 2 and 0 identifiers

# Generate NUTS 2 and 0 Identifiers
main_data <- GDP_per_capita |> 
  mutate(
    id_nuts2 = substr(id_nuts3, 1, 4),
    id_nuts0 = substr(id_nuts3, 1, 2)
  ) 
```

```{r}
#| echo: false
#| output: false

# Calculating gini
gini <- main_data |>
  filter(population != 0) |>
  filter(population != "NA") |>
  filter(GDP_per_capita != 0) |>
  filter(GDP_per_capita != "NA") |>
  group_by(id_nuts2, year) |>
  mutate(gini_n2 = gini.wtd(GDP_per_capita, weights = population)) |>
  ungroup()

# Slice function to remove numbers that get repeated. 
gini_nuts_2 <- gini |>
  group_by(id_nuts2, year) |>
  slice(1) |>
  ungroup()

# Since some NUTS 2 and 3 regions are the same (like Copenhagen for example), some gini coefficients will be 0. We should remove them before doing further analysis and data presentation

# Filter data
gini_filtered <- gini_nuts_2 |>
  filter(gini_n2 != 0)
```

After calculating the gini coefficients, we can see that there are some similarities to the data we got from GDP per capita for NUTS 3 regions. In order to see these similarities better, as well as look for other important aspects that can be provided trough the calculations, we will visualize the data in three different ways.

```{r}
#| echo: false

# Creating a bar chart
ggplot(gini_filtered, aes(x = id_nuts2, y = gini_n2, fill = id_nuts0)) +
  geom_bar(stat = "identity") +
  ggtitle("Gini Coefficient by NUTS 2 Region") +
  xlab("NUTS 2 Region ID") +
  ylab("Gini Coefficient") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
#| echo: false

# Creating a time-series plot
ggplot(gini_filtered, aes(x = year, y = gini_n2, color = id_nuts2)) +
  geom_line() +
  ggtitle("Time-series Plot of Gini Coefficient by NUTS 2 Region") +
  xlab("Year") +
  ylab("Gini Coefficient")
```

```{r}
#| echo: false

# Creating a box plot
ggplot(gini_filtered, aes(x = id_nuts0, y = GDP_per_capita)) +
  geom_boxplot() +
  ggtitle("Distribution of GDP per Capita by Country") +
  xlab("Country") +
  ylab("GDP per Capita")
```

## Discussion

# References

### 2023-10-22 assignment_1.tex
Update 22.10

--- 2023-10-22 assignment_1.qmd
@@
----
-title: "Assignment 1: Data and Descriptive"
-author: "Anine Therese Karlsen & Mona Lisa Jones"
-format: pdf
-abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
-Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
-editor: visual
-bibliography: [Econometrics.bib]
-page-layout: full
-number-sections: true
-papersize: a4
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  a4paper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
-```{r}
-#| label: setup
-#| output: false
-#| echo: false
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-library(tidyverse)
-library(tinytex)
-library(ggplot2)
-library(readxl)
-library(magrittr)
-library(vtable)
-library(lubridate)
-library(dineq)
-```
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-# Introduction
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Assignment 1: Data and Descriptive},
+  pdfauthor={Anine Therese Karlsen \& Mona Lisa Jones},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-While national GDP and GDP per capita are vital indicators of a country's aggregate economic health, they do not shed light on how wealth or income is distributed among its residents.
-A high national GDP can, paradoxically, coexist with pockets of regional deprivation [@lessmann2017].
+\title{Assignment 1: Data and Descriptive}
+\author{Anine Therese Karlsen \& Mona Lisa Jones}
+\date{}
 
-The truth of this statement becomes more evident when taking a closer look in to into sub-national data.
-Regional wealth disparities are of prime concern, especially when crafting policies for equitable growth [@lessmann2017].
-A country's macro-level prosperity does not automatically guarantee that all its regions partake equally in this wealth.
-By studying smaller regions within a country, it is possible to get a more nuanced narrative about the state of regional economic disparities [@lessmann2017].
+\begin{document}
+\maketitle
+\begin{abstract}
+This assignment aims to acquire, process, and analyze sub-national GDP
+and population data for a subset of European countries. Calculate GDP
+per capita and explore regional inequity using various descriptive
+statistics and visualizations.
+\end{abstract}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
 
-This assignment is anchored in this very premise.
-It seeks to acquire, process, and analyse sub-national GDP and population data for a selected subset of European countries, namely France, Denmark, Hungary, Portugal, and Slovakia, spanning the period from 2000 to 2020.
-The overarching aim is to calculate GDP per capita for these regions and delve into the intricacies of regional inequity.
-This exploration will be underpinned by various descriptive statistics, visualizations, and analytical techniques to paint a comprehensive picture of regional economic landscapes.
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Using time series data observing the GDP per capita and GINI trends across regions and over time, we can discern whether certain areas are ahead or behind in economic performance.
+While national GDP and GDP per capita are vital indicators of a
+country's aggregate economic health, they do not shed light on how
+wealth or income is distributed among its residents. A high national GDP
+can, paradoxically, coexist with pockets of regional deprivation
+(Lessmann and Seidel 2017).
 
-# Literature review
+The truth of this statement becomes more evident when taking a closer
+look in to into sub-national data. Regional wealth disparities are of
+prime concern, especially when crafting policies for equitable growth
+(Lessmann and Seidel 2017). A country's macro-level prosperity does not
+automatically guarantee that all its regions partake equally in this
+wealth. By studying smaller regions within a country, it is possible to
+get a more nuanced narrative about the state of regional economic
+disparities (Lessmann and Seidel 2017).
 
-"In recent decades, the regional distribution of incomes within countries has attracted considerable interest among academics and policy makers." [@lessmann2017]
+This assignment is anchored in this very premise. It seeks to acquire,
+process, and analyse sub-national GDP and population data for a selected
+subset of European countries, namely France, Denmark, Hungary, Portugal,
+and Slovakia, spanning the period from 2000 to 2020. The overarching aim
+is to calculate GDP per capita for these regions and delve into the
+intricacies of regional inequity. This exploration will be underpinned
+by various descriptive statistics, visualizations, and analytical
+techniques to paint a comprehensive picture of regional economic
+landscapes.
 
-Slovakia, experienced significant economic growth. Their GDP per capita grew faster than the older EU member states, albeit from a lower starting point, leading to a convergence [@fullart].
+Using time series data observing the GDP per capita and GINI trends
+across regions and over time, we can discern whether certain areas are
+ahead or behind in economic performance.
 
-The global financial crisis and subsequent European debt crisis had significant impacts on European economies.
-GDP per capita declined in several countries, particularly those hit hardest by the crisis, such as Greece, Spain, and Portugal.
-Some regions took many years to recover to pre-crisis levels, while others, especially in Northern Europe, recovered more quickly [@lewis].
+\hypertarget{literature-review}{%
+\section{Literature review}\label{literature-review}}
 
-Major metropolitan areas, like Paris in France, London in the UK, and Berlin in Germany, often grew faster than other regions in their respective countries, leading to increasing spatial disparities [@kiuru2019].
+``In recent decades, the regional distribution of incomes within
+countries has attracted considerable interest among academics and policy
+makers.'' (Lessmann and Seidel 2017)
 
-Denmark maintained relatively high levels of GDP per capita throughout this period, benefiting from a diversified economy, strong institutions, and a high degree of economic openness.
-It did face challenges during the 2008 economic crisis but recovered relatively quickly [@dynamics].
+Slovakia, experienced significant economic growth. Their GDP per capita
+grew faster than the older EU member states, albeit from a lower
+starting point, leading to a convergence ({``Full Article: The Causal
+Impact of Economic Growth on Material Use in Europe,''} n.d.).
 
-Slovenia, which joined the EU in 2004, showed growth in GDP per capita, though it too was affected by the global economic downturn in 2008.
-However, it managed to recover in the subsequent years [@fullart].
+The global financial crisis and subsequent European debt crisis had
+significant impacts on European economies. GDP per capita declined in
+several countries, particularly those hit hardest by the crisis, such as
+Greece, Spain, and Portugal. Some regions took many years to recover to
+pre-crisis levels, while others, especially in Northern Europe,
+recovered more quickly (Lewis and Verhoeven, n.d.).
 
-Regional disparities persisted, with the Île-de-France region (which includes Paris) significantly outperforming other regions in terms of GDP per capita.
-The gap between urban and rural areas also remained a topic of discussion [@regional].
+Major metropolitan areas, like Paris in France, London in the UK, and
+Berlin in Germany, often grew faster than other regions in their
+respective countries, leading to increasing spatial disparities (Kiuru
+and Inkinen 2019).
 
-Towards the end of this period, in early 2020, the COVID-19 pandemic posed a new set of challenges for European economies, causing significant contractions in GDP per capita across many regions [@regional].
+Denmark maintained relatively high levels of GDP per capita throughout
+this period, benefiting from a diversified economy, strong institutions,
+and a high degree of economic openness. It did face challenges during
+the 2008 economic crisis but recovered relatively quickly ({``Dynamics
+of Technological Innovation, Energy Consumption, Energy Price and
+Economic Growth in Denmark - Murad - 2019 - Environmental Progress \&
+Sustainable Energy - Wiley Online Library,''} n.d.).
 
-# Part A: Sub-national GDP and GDP per Capita
+Slovenia, which joined the EU in 2004, showed growth in GDP per capita,
+though it too was affected by the global economic downturn in 2008.
+However, it managed to recover in the subsequent years ({``Full Article:
+The Causal Impact of Economic Growth on Material Use in Europe,''}
+n.d.).
 
-## Data Acquisition and datasets
+Regional disparities persisted, with the Île-de-France region (which
+includes Paris) significantly outperforming other regions in terms of
+GDP per capita. The gap between urban and rural areas also remained a
+topic of discussion ({``Regional Inequality in Europe: Evidence, Theory
+and Policy Implications \textbar{} Journal of Economic Geography
+\textbar{} Oxford Academic,''} n.d.).
+
+Towards the end of this period, in early 2020, the COVID-19 pandemic
+posed a new set of challenges for European economies, causing
+significant contractions in GDP per capita across many regions
+({``Regional Inequality in Europe: Evidence, Theory and Policy
+Implications \textbar{} Journal of Economic Geography \textbar{} Oxford
+Academic,''} n.d.).
+
+\hypertarget{part-a-sub-national-gdp-and-gdp-per-capita}{%
+\section{Part A: Sub-national GDP and GDP per
+Capita}\label{part-a-sub-national-gdp-and-gdp-per-capita}}
+
+\hypertarget{data-acquisition-and-datasets}{%
+\subsection{Data Acquisition and
+datasets}\label{data-acquisition-and-datasets}}
 
 The first thing we did in this assignment was to acquire our data.
-Trough Eurostat, we could download the datasets nama_10r_3gdp and demo_r_pjanggr3 as csv files, as well as filter the data by our preferences before downloading it.
-As per the assignment, we filtered the dataset by choosing the years 2000 to 2020.
-Furthermore, we selected the NUTS 3 region for the nations we were given, which were Portugal, France, Hungary, Slovakia and Denmark.
-Finally, we specified the data to be in million Euro.
+Trough Eurostat, we could download the datasets nama\_10r\_3gdp and
+demo\_r\_pjanggr3 as csv files, as well as filter the data by our
+preferences before downloading it. As per the assignment, we filtered
+the dataset by choosing the years 2000 to 2020. Furthermore, we selected
+the NUTS 3 region for the nations we were given, which were Portugal,
+France, Hungary, Slovakia and Denmark. Finally, we specified the data to
+be in million Euro.
 
-### GDP (nama_10r_3gdp)
+\hypertarget{gdp-nama_10r_3gdp}{%
+\subsubsection{GDP (nama\_10r\_3gdp)}\label{gdp-nama_10r_3gdp}}
 
-The nama_10r_3gdp dataset from Eurostat provides insights into GDP at regional level using the NUTS classification system.
-It furnishes GDP values in both current prices and adjusted for inflation, with some figures given in purchasing power standards (PPS) to account for price level differences between countries.
-The data is often structured by year and region @eurostat2023.
+The nama\_10r\_3gdp dataset from Eurostat provides insights into GDP at
+regional level using the NUTS classification system. It furnishes GDP
+values in both current prices and adjusted for inflation, with some
+figures given in purchasing power standards (PPS) to account for price
+level differences between countries. The data is often structured by
+year and region Eurostat (2023a).
 
-The GDP at market prices represents the final result of production activities of resident producer units within a region or nation.
-It is calculated as the sum of the gross value added across various institutional sectors or industries, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries).
-This also balances out in the total economy production account.
-In terms of methodology, while national accounts compile GDP from the expenditure side, regional accounts don't adopt this approach due to the complexities of accurately mapping inter-regional flows of goods and services.
+The GDP at market prices represents the final result of production
+activities of resident producer units within a region or nation. It is
+calculated as the sum of the gross value added across various
+institutional sectors or industries, augmented by taxes and reduced by
+subsidies on products (which are not allocated to specific sectors or
+industries). This also balances out in the total economy production
+account. In terms of methodology, while national accounts compile GDP
+from the expenditure side, regional accounts don't adopt this approach
+due to the complexities of accurately mapping inter-regional flows of
+goods and services.
 
-The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @eurostat2023c.
+The different measures for the regional GDP are absolute figures in €
+and Purchasing Power Standards (PPS), figures per inhabitant and
+relative data compared to the EU Member States average Eurostat (2023b).
 
-### Population (demo_r_pjanggr3)
+\hypertarget{population-demo_r_pjanggr3}{%
+\subsubsection{Population
+(demo\_r\_pjanggr3)}\label{population-demo_r_pjanggr3}}
 
-Using the NUTS categorization once more, Eurostat's demo_r_pjanggr3 records annual population changes at the regional level.
-This dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender.
-It's also often displayed in a year-by-region format, and the data usually spans in yearly intervals @eurostat2023a.
+Using the NUTS categorization once more, Eurostat's demo\_r\_pjanggr3
+records annual population changes at the regional level. This dataset
+includes information on births, deaths, net migration, and may also
+include demographic information on age and gender. It's also often
+displayed in a year-by-region format, and the data usually spans in
+yearly intervals Eurostat (2023c).
 
-Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project.
-The project covers 37 countries and is the central repository for demographic and migration-related data.
-Specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year.
-Additionally, data on marriages, divorces, and migration flows are recorded.
+Eurostat's primary source for yearly demographic data at the regional
+level stems from the Unified Demography (Unidemo) project. The project
+covers 37 countries and is the central repository for demographic and
+migration-related data. Specific metrics gathered under UNIDEMO
+encompass population counts at the close of the calendar year and events
+such as births and deaths occurring within that year. Additionally, data
+on marriages, divorces, and migration flows are recorded.
 
-For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @eurostat2021.
+For the purpose of this research, the demographic data references the
+NUTS 2016 classification, which provides a detailed breakdown of the
+European Union's territory Eurostat (2021).
 
-### NUTS classification
+\hypertarget{nuts-classification}{%
+\subsubsection{NUTS classification}\label{nuts-classification}}
 
-The Nomenclature of Territorial Units for Statistics (NUTS) offers a stratified system to segment the economic territory of the EU (including the UK) to facilitate the consistent collection and harmonization of regional statistics across Europe.
-The NUTS regions range from NUTS 0 Country level to NUTS 3 small units such as municipalities level.
+The Nomenclature of Territorial Units for Statistics (NUTS) offers a
+stratified system to segment the economic territory of the EU (including
+the UK) to facilitate the consistent collection and harmonization of
+regional statistics across Europe. The NUTS regions range from NUTS 0
+Country level to NUTS 3 small units such as municipalities level. ~
 
-```{r}
-#| echo: false
-#| output: false
-
-# This section of the code is dedicated to adding data to RStudio. We will import the datasets into the RStudio environment for further analysis, manipulation, and visualization using R.
-
-# Data for population
-pop_data <- read_csv("Raw_data/demo.gz")
-
-# Data for GDP
-gdp_data <- read_csv("Raw_data/nama.gz")
-```
-
-```{r}
-#| echo: false
-#| output: false
-# Renaming and selecting the variables we want to keep
-# We're doing this in order to make the datasets more clean and easy to read
-
-gdp_data_01 <- gdp_data |>
-  select(geo, year = TIME_PERIOD, GDP = OBS_VALUE)
-
-pop_data_01 <- pop_data |>
-  select(geo, year = TIME_PERIOD, population = OBS_VALUE)
-```
-
-## GDP per Capita Calculation
+\hypertarget{gdp-per-capita-calculation}{%
+\subsection{GDP per Capita
+Calculation}\label{gdp-per-capita-calculation}}
 
@@
 
-$y_i=GDP_i/population_i$
+\(y_i=GDP_i/population_i\)
 
-```{r}
-#| echo: false
-#| output: false
+After calculating the GDP per capita for all NUTS 3 regions in our
+assigned countries, we can see that there is a large spread between the
+figures for the various regions. In this assignement we want to look at
+regional inequity; in order to do this in a valuable way we have to
+divide between the different countries. By doing this, we can gain
+important insights on regional differences that we can utilize, for
+instance, to discuss national policy on equity and sustainable economic
+development in regions.
 
-# Merging the datasets pop_data and gdp_data into one dataset by joining the variables "geo" and "year".
+\hypertarget{descriptive-statistics}{%
+\subsubsection{Descriptive statistics}\label{descriptive-statistics}}
 
-merged_data <- left_join(pop_data_01, gdp_data_01, join_by(geo, year))
+In this part we will report and interpret different types of essential
+descriptive statistics. Measuring regional income inequality is
+challenging due to heterogeneity of regions. The number of regions in
+our data set varies largely in size and population. Since the focus of
+this paper is purely growth and changes in inequities over time, the
+variations of size and population density becomes a minor issue because
+the country-level territorial heterogeneity is fixed.
 
-# Select Only NUTS 3 Regions
-# Filter for specific countries, this will be useful when we will find the GDP per capita for the countries seperately
+``Interest in income inequality has led to the development of several
+ways of measuring it. Two types of measures are of interest in this
+paper----static and dynamic. \emph{Static measures} provide a snapshot
+of these inequalities at a point of time whereas the dynamic
+\emph{measures capture historical trends.}'' (læreboken)
 
-merged_data <- merged_data |>
-  filter(
-    startsWith(as.character(geo), "PT") | 
-    startsWith(as.character(geo), "FR") | 
-    startsWith(as.character(geo), "HU") |
-    startsWith(as.character(geo), "SK") | 
-    startsWith(as.character(geo), "DK")
-  ) %>%
+In this part, we'll look at GDP per capita for our assigned countries on
+a NUTS 3 level. In addition, we'll use different kinds of descriptive
+statistics in order to further 0,37, is analyse this data. In this
+analysis, we'll use Wooldridge (2020) for help.
 
-# Filter rows where 'geo' column has more than 4 characters to select only NUTS3 regions
-  filter(nchar(as.character(geo)) > 4) %>%
+By using figures, we can visualize the GDP per capita, and look at how
+it varies among the different regions. In these figures, a line
+represent one NUTS 3 region.
 
-# Rename 'geo' column to 'id_nuts3'
-  rename(id_nuts3 = geo)
-```
+\textbf{Mean}
 
-```{r}
-#| echo: false
-#| output: false
+Calculate the mean to provide a representative value for a dataset,
+facilitating understanding of its central tendency and serving as a
+benchmark against which deviations and anomalies can be assessed, in
+later steps when building and interpreting regression models.
 
-# Calculating GDP per Capita
+\textbf{MMR}
 
-GDP_per_capita <- merged_data |>
-  mutate(
-    GDP_per_capita = round(100*GDP/population, digits = 2)
-  )
+A comparison of the GRDP (gross regional domestic product) per capita of
+the region with the highest income to the region with the lowest income
+(minimum per capita GRDP) provides a measure of the range of these
+disparities. If this measure is small (close to 1), then it would mean
+that the different regions have relatively equal incomes. If this
+measure is large, then the interpretation is more problematic, as it
+does not tell us if the high ratio is due to substantial variation in
+the distribution of per capita GDRPs or the presence of outliers.
+Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to
+comprehend, and politically powerful measure of regional income
+inequality.
 
-# we use *100 in order to get the gdp per capita as a percentage, and digits = 2 to limit it down to two decimals
-```
+\textbf{Standard diviation (SD)}
 
-After calculating the GDP per capita for all NUTS 3 regions in our assigned countries, we can see that there is a large spread between the figures for the various regions.
-In this assignement we want to look at regional inequity; in order to do this in a valuable way we have to divide between the different countries.
-By doing this, we can gain important insights on regional differences that we can utilize, for instance, to discuss national policy on equity and sustainable economic development in regions.
+Calculating SD to quantify the dispersion or variability of a data set
+around its mean. Helping us assess the degree of uncertainty,
+variability, or risk associated with an economic variable or parameter,
+which is crucial for understanding the reliability of estimations and
+predictions (Wooldridge 2020).
 
-### Descriptive statistics
+\textbf{Median}
 
-In this part we will report and interpret different types of essential descriptive statistics.
-Measuring regional income inequality is challenging due to heterogeneity of regions.
-The number of regions in our data set varies largely in size and population.
-Since the focus of this paper is purely growth and changes in inequities over time, the variations of size and population density becomes a minor issue because the country-level territorial heterogeneity is fixed.
+The median serves as a robust measure of central tendency, especially
+when a dataset may have outliers or is skewed. Unlike the mean, the
+median is not influenced by extreme values and, thus, can provide a
+clearer picture of the ``typical'' value in situations where the data
+distribution is not symmetrical (Wooldridge 2020).
 
-"Interest in income inequality has led to the development of several ways of measuring it.
-Two types of measures are of interest in this paper----static and dynamic.
-*Static measures* provide a snapshot of these inequalities at a point of time whereas the dynamic *measures capture historical trends.*" (læreboken)
+\hypertarget{portugal}{%
+\subsubsection{Portugal}\label{portugal}}
 
-In this part, we'll look at GDP per capita for our assigned countries on a NUTS 3 level.
-In addition, we'll use different kinds of descriptive statistics in order to further 0,37, is analyse this data.
-In this analysis, we'll use @wooldridge2020 for help.
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-7-1.pdf}
 
-By using figures, we can visualize the GDP per capita, and look at how it varies among the different regions. In these figures, a line represent one NUTS 3 region.
+\begin{verbatim}
+        GDP_per_capita
+mean         1.4185524
+median       1.3900000
+std_dev      0.3702905
+minimum      0.6500000
+maximum      2.7200000
+\end{verbatim}
 
-**Mean**
+By looking at figure for Portugal, we can see that the GDP per capita in
+Portugal's regions appears to be fairly consistent. There is however
+some regional variability. We can see that the regions around the big
+cities like Lisbon have a higher GDP per capita compared to some more
+rural areas. Since Lisbon is the capital of Portugal, there is probably
+a higher concentration of industries, making it a economic center (which
+again makes the GDP per capita higher).
 
-Calculate the mean to provide a representative value for a dataset, facilitating understanding of its central tendency and serving as a benchmark against which deviations and anomalies can be assessed, in later steps when building and interpreting regression models.
+To continue, we can see that the mean is a little higher that the
+median, something that might indicate that regions like Lisbon are
+pulling up the average. If we compare the standard derivation for
+Portugal with the other countries, we'll see that is fairly low in
+comparison. This might mean that there is not a lot of variability
+between the GDP per capita across different regions in Portugal. The gap
+between minimum and maximum is also low compared to other countries,
+something that'll also show us that the economic disparity in Portugal
+might not be as high as it is in other countries.
 
-**MMR**
+\hypertarget{france}{%
+\subsubsection{France}\label{france}}
 
-A comparison of the GRDP (gross regional domestic product) per capita of the region with the highest income to the region with the lowest income (minimum per capita GRDP) provides a measure of the range of these disparities.
-If this measure is small (close to 1), then it would mean that the different regions have relatively equal incomes.
-If this measure is large, then the interpretation is more problematic, as it does not tell us if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers.
-Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to comprehend, and politically powerful measure of regional income inequality.
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-10-1.pdf}
 
-**Standard diviation (SD)**
+\begin{verbatim}
+        GDP_per_capita
+mean          2.630444
+median        2.440000
+std_dev       1.053767
+minimum       0.830000
+maximum      11.580000
+\end{verbatim}
 
-Calculating SD to quantify the dispersion or variability of a data set around its mean.
-Helping us assess the degree of uncertainty, variability, or risk associated with an economic variable or parameter, which is crucial for understanding the reliability of estimations and predictions [@wooldridge2020].
-
-**Median**
-
-The median serves as a robust measure of central tendency, especially when a dataset may have outliers or is skewed.
-Unlike the mean, the median is not influenced by extreme values and, thus, can provide a clearer picture of the "typical" value in situations where the data distribution is not symmetrical [@wooldridge2020].
-
-### Portugal
-
-```{r}
-#| echo: false
-#| output: false
-
-# Making a dataset that only contains data from Portugal
-# Use the na.omit function in order to remove NA values, cant make ggplot with NA values.
-
-Portugal_data <- GDP_per_capita |>
-  filter(grepl("PT", id_nuts3)) |>
-  na.omit(Portugal_data)
-```
-
-```{r}
-#| echo: false
-
-# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Portugal.
-
-Portugal_data_plot <- ggplot(Portugal_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
-  geom_line() +
-  labs(title="Figure 1: GDP per Capita for Portugal", 
-       x="Year",
-       y="GDP per Capita") +
-  theme(legend.position="none")
-
-# Print out the figure
-
-print(Portugal_data_plot)
-```
-
-```{r}
-#| echo: false
-
-# Calculate summary statistics for Portugal, chosing only GDP per capita 
-summary_stats_portugal <- sapply(Portugal_data["GDP_per_capita"], function(x) c(
-  mean = mean(x),
-  median = median(x),
-  std_dev = sd(x),
-  minimum = min(x),
-  maximum = max(x)
-))
-
-# Print the summary statistics
-print(summary_stats_portugal)
-```
-
-By looking at figure for Portugal, we can see that the GDP per capita in Portugal's regions appears to be fairly consistent.
-There is however some regional variability.
-We can see that the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas.
-Since Lisbon is the capital of Portugal, there is probably a higher concentration of industries, making it a economic center (which again makes the GDP per capita higher).
-
-To continue, we can see that the mean is a little higher that the median, something that might indicate that regions like Lisbon are pulling up the average.
-If we compare the standard derivation for Portugal with the other countries, we'll see that is fairly low in comparison.
-This might mean that there is not a lot of variability between the GDP per capita across different regions in Portugal.
-The gap between minimum and maximum is also low compared to other countries, something that'll also show us that the economic disparity in Portugal might not be as high as it is in other countries.
-
-### France
-
-```{r}
-#| echo: false
-#| output: false
-
-# Making a dataset that only contains data from France
-
-France_data <- GDP_per_capita |>
-  filter(grepl("FR", id_nuts3)) |>
-  na.omit(France_data)
-```
-
-```{r}
-#| echo: false
-
-# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in France.
-
-France_data_plot <- ggplot(France_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
-  geom_line() +
-  labs(title="Figure 1: GDP per Capita for France", 
-       x="Year",
-       y="GDP per Capita") +
-  theme(legend.position="none")
-
-print(France_data_plot)
-```
-
-```{r}
-#| echo: false
-
-# Calculate summary statistics for France, chosing only GDP per capita 
-summary_stats_france <- sapply(France_data["GDP_per_capita"], function(x) c(
-  mean = mean(x),
-  median = median(x),
-  std_dev = sd(x),
-  minimum = min(x),
-  maximum = max(x)
-))
-
-# Print the summary statistics
-print(summary_stats_france)
-```
-
-When looking at the figure for France, we can right away see that there are some regions that have a much higher GDP per capita compared to the other regions.
-The region that has the highest GDP per capita for all years is the île-de-France region, one that also includes Paris.
-This significant difference between the regions with the highest GDP per capita and the lowest, shows us that there is a high concentration of economic activity and wealth in a few urban regions.
-Similar to Portugal, we can also again see that there is a difference between urban and rural regions.
+When looking at the figure for France, we can right away see that there
+are some regions that have a much higher GDP per capita compared to the
+other regions. The region that has the highest GDP per capita for all
+years is the île-de-France region, one that also includes Paris. This
+significant difference between the regions with the highest GDP per
+capita and the lowest, shows us that there is a high concentration of
+economic activity and wealth in a few urban regions. Similar to
+Portugal, we can also again see that there is a difference between urban
+and rural regions.
 
 Just as in Portugal, there is also a higher mean in France as well.
-Something that is different from the data in France compared to Portugal, is that the standard derivation is higher, and the difference between minimum and maximum is large.
-This strengthens what we have look at earlier in the figure, with some regions having a high concentration of wealth.
+Something that is different from the data in France compared to
+Portugal, is that the standard derivation is higher, and the difference
+between minimum and maximum is large. This strengthens what we have look
+at earlier in the figure, with some regions having a high concentration
+of wealth.
 
-### Hungary
+\hypertarget{hungary}{%
+\subsubsection{Hungary}\label{hungary}}
 
-```{r}
-#| echo: false
-#| output: false
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-13-1.pdf}
 
-# Making a dataset that only contains data from Hungary
+\begin{verbatim}
+        GDP_per_capita
+mean         0.8598000
+median       0.7650000
+std_dev      0.4059723
+minimum      0.3200000
+maximum      3.1300000
+\end{verbatim}
 
-Hungary_data <- GDP_per_capita |>
-  filter(grepl("HU", id_nuts3)) |>
-  na.omit(Hungary_data)
-```
+In Hungary, most of the regions have similar GDP per Capita. One region
+that sticks out by having a higher value, is the region of Budapest,
+Hungary's biggest city.
 
-```{r}
-#| echo: false
+We have here as well an mean that is larger than the median, high
+standard derivation, and a large gap between minimum and maximum.
 
-# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Hungary.
+\hypertarget{slovakia}{%
+\subsubsection{Slovakia}\label{slovakia}}
 
-Hungary_data_plot <- ggplot(Hungary_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
-  geom_line() +
-  labs(title="Figure 1: GDP per Capita for Hungary", 
-       x="Year",
-       y="GDP per Capita") +
-  theme(legend.position="none")
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-16-1.pdf}
 
-print(Hungary_data_plot)
-```
+\begin{verbatim}
+        GDP_per_capita
+mean         1.2501948
+median       1.0950000
+std_dev      0.8018259
+minimum      0.3000000
+maximum      4.0200000
+\end{verbatim}
 
-```{r}
-#| echo: false
+In Slovakia as well, we have one region that has a much higher GDP per
+capita than the rest of the regions. This region is Bratislava, which is
+the biggest city and capital, something that might point to this city
+being the economic capital of Slovakia as well.
 
-# Calculate summary statistics for Hungary, chosing only GDP per capita 
-summary_stats_hungary <- sapply(Hungary_data["GDP_per_capita"], function(x) c(
-  mean = mean(x),
-  median = median(x),
-  std_dev = sd(x),
-  minimum = min(x),
-  maximum = max(x)
-))
+Slovakia has also a mean higher that the median, and a large gap between
+minimum and maximum. In addition, the standard derivation is pretty
+high, meaning that there is some regions (or one region in this case)
+that is far away from the rest of the regions when it comes to economic
+development.
 
-# Print the summary statistics
-print(summary_stats_hungary)
-```
+\hypertarget{denmark}{%
+\subsubsection{Denmark}\label{denmark}}
 
-In Hungary, most of the regions have similar GDP per Capita. One region that sticks out by having a higher value, is the region of Budapest, Hungary's biggest city.
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-19-1.pdf}
 
-We have here as well an mean that is larger than the median, high standard derivation, and a large gap between minimum and maximum.
+\begin{verbatim}
+        GDP_per_capita
+mean          4.419221
+median        4.000000
+std_dev       1.343933
+minimum       2.760000
+maximum       8.420000
+\end{verbatim}
 
-### Slovakia
+Lastly, we have Denmark. We can see similar pattern here as well, with
+the capital Copenhagen being one of the regions with the highest GDP per
+capita.
 
-```{r}
-#| echo: false
-#| output: false
+We can also see the same as the previous countries, with the mean being
+higher than the median, which shows us that regions like Copenhagen
+might drag the mean up by being much larger than the rest of the
+regions.
 
-# Making a dataset that only contains data from Slovakia
+\hypertarget{part-b-regional-inequity}{%
+\section{Part B: Regional Inequity}\label{part-b-regional-inequity}}
 
-Slovakia_data <- GDP_per_capita |>
-  filter(grepl("SK", id_nuts3)) |>
-  na.omit(Slovakia_data)
-```
-
-```{r}
-#| echo: false
-
-# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Slovakia.
-
-Slovakia_data_plot <- ggplot(Slovakia_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
-  geom_line() +
-  labs(title="Figure 1: GDP per Capita for Slovakia", 
-       x="Year",
-       y="GDP per Capita") +
-  theme(legend.position="none")
-
-print(Slovakia_data_plot)
-```
-
-```{r}
-#| echo: false
-
-# Calculate summary statistics for Slovakia, chosing only GDP per capita 
-summary_stats_slovakia <- sapply(Slovakia_data["GDP_per_capita"], function(x) c(
-  mean = mean(x),
-  median = median(x),
-  std_dev = sd(x),
-  minimum = min(x),
-  maximum = max(x)
-))
-
-# Print the summary statistics
-print(summary_stats_slovakia)
-```
-
-In Slovakia as well, we have one region that has a much higher GDP per capita than the rest of the regions.
-This region is Bratislava, which is the biggest city and capital, something that might point to this city being the economic capital of Slovakia as well.
-
-Slovakia has also a mean higher that the median, and a large gap between minimum and maximum.
-In addition, the standard derivation is pretty high, meaning that there is some regions (or one region in this case) that is far away from the rest of the regions when it comes to economic development.
-
-### Denmark
-
-```{r}
-#| echo: false
-#| output: false
-
-# Making a dataset that only contains data from Denmark
-
-Denmark_data <- GDP_per_capita |>
-  filter(grepl("DK", id_nuts3)) |>
-  na.omit(Denmark_data)
-```
-
-```{r}
-#| echo: false
-
-# Making a figure using ggplot that shows us variations of GDP per capita across the different NUTS 3 regions in Denmark.
-
-Denmark_data_plot <- ggplot(Denmark_data, aes(x = year, y = GDP_per_capita, color = id_nuts3)) +
-  geom_line() +
-  labs(title="Figure 1: GDP per Capita for Denmark", 
-       x="Year",
-       y="GDP per Capita") +
-  theme(legend.position="none")
-
-print(Denmark_data_plot)
-```
-
-```{r}
-#| echo: false
-
-# Calculate summary statistics for Denmark, chosing only GDP per capita 
-summary_stats_denmark <- sapply(Denmark_data["GDP_per_capita"], function(x) c(
-  mean = mean(x),
-  median = median(x),
-  std_dev = sd(x),
-  minimum = min(x),
-  maximum = max(x)
-))
-
-# Print the summary statistics
-print(summary_stats_denmark)
-```
-
-Lastly, we have Denmark. We can see similar pattern here as well, with the capital Copenhagen being one of the regions with the highest GDP per capita.
-
-We can also see the same as the previous countries, with the mean being higher than the median, which shows us that regions like Copenhagen might drag the mean up by being much larger than the rest of the regions.
-
-# Part B: Regional Inequity
-
-## Gini Coefficient Calculation
+\hypertarget{gini-coefficient-calculation}{%
+\subsection{Gini Coefficient
+Calculation}\label{gini-coefficient-calculation}}
 
@@
 
-$GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|$
+\(GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|\)
 
-we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.
+we will compute the population-weighted GDP Gini coefficient for each
+European NUTS2 region in our assigned countries.
 
-The gini coefficient can help us measure inequality in a distribution, as is therefore a useful tool for us to use when we look at regional inequity.
-The closer the gini coefficient is to 1, the bigger the inequality is; a number closer to 0 equals equality.
-When looking at the gini coefficient for NUTS 2 regions, we also get a better overview over differences in income between different regions, and it also makes it easier to find the reasons as to why there is a difference between the regions @hasell2023.
+The gini coefficient can help us measure inequality in a distribution,
+as is therefore a useful tool for us to use when we look at regional
+inequity. The closer the gini coefficient is to 1, the bigger the
+inequality is; a number closer to 0 equals equality. When looking at the
+gini coefficient for NUTS 2 regions, we also get a better overview over
+differences in income between different regions, and it also makes it
+easier to find the reasons as to why there is a difference between the
+regions Hasell and Roser (2023).
 
-```{r}
-#| echo: false
-#| output: false
+After calculating the gini coefficients, we can see that there are some
+similarities to the data we got from GDP per capita for NUTS 3 regions.
+In order to see these similarities better, as well as look for other
+important aspects that can be provided trough the calculations, we will
+visualize the data in three different ways.
 
-# Before we can find the Gini coefficient for NUTS2 regions, we have to generate NUTS 2 and 0 identifiers
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-23-1.pdf}
 
-# Generate NUTS 2 and 0 Identifiers
-main_data <- GDP_per_capita |> 
-  mutate(
-    id_nuts2 = substr(id_nuts3, 1, 4),
-    id_nuts0 = substr(id_nuts3, 1, 2)
-  ) 
-```
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-24-1.pdf}
 
-```{r}
-#| echo: false
-#| output: false
+\includegraphics{assignment_1_files/figure-pdf/unnamed-chunk-25-1.pdf}
 
-# Calculating gini
-gini <- main_data |>
-  filter(population != 0) |>
-  filter(population != "NA") |>
-  filter(GDP_per_capita != 0) |>
-  filter(GDP_per_capita != "NA") |>
-  group_by(id_nuts2, year) |>
-  mutate(gini_n2 = gini.wtd(GDP_per_capita, weights = population)) |>
-  ungroup()
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
-# Slice function to remove numbers that get repeated. 
-gini_nuts_2 <- gini |>
-  group_by(id_nuts2, year) |>
-  slice(1) |>
-  ungroup()
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
-# Since some NUTS 2 and 3 regions are the same (like Copenhagen for example), some gini coefficients will be 0. We should remove them before doing further analysis and data presentation
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-dynamics}{}}%
+{``Dynamics of Technological Innovation, Energy Consumption, Energy
+Price and Economic Growth in Denmark - Murad - 2019 - Environmental
+Progress \& Sustainable Energy - Wiley Online Library.''} n.d.
+\url{https://aiche.onlinelibrary.wiley.com/doi/full/10.1002/ep.12905}.
 
-# Filter data
-gini_filtered <- gini_nuts_2 |>
-  filter(gini_n2 != 0)
-```
+\leavevmode\vadjust pre{\hypertarget{ref-eurostat2021}{}}%
+Eurostat. 2021. {``Population Change - {Demographic} Balance and Crude
+Rates at Regional Level ({NUTS} 3)~ (Demo\_r\_gind3).''}
+https://ec.europa.eu/eurostat/cache/metadata/en/demo\_r\_gind3\_esms.htm.
 
-After calculating the gini coefficients, we can see that there are some similarities to the data we got from GDP per capita for NUTS 3 regions.
-In order to see these similarities better, as well as look for other important aspects that can be provided trough the calculations, we will visualize the data in three different ways.
+\leavevmode\vadjust pre{\hypertarget{ref-eurostat2023}{}}%
+---------. 2023a. {``Gross Domestic Product ({GDP}) at Current Market
+Prices by {NUTS} 3 Regions.''}
+https://ec.europa.eu/eurostat/databrowser/view/nama\_10r\_3gdp/default/table?lang=en.
 
-```{r}
-#| echo: false
+\leavevmode\vadjust pre{\hypertarget{ref-eurostat2023c}{}}%
+---------. 2023b. {``Regional Economic Accounts (Reg\_eco10).''}
+https://ec.europa.eu/eurostat/cache/metadata/en/reg\_eco10\_esms.htm.
 
-# Creating a bar chart
-ggplot(gini_filtered, aes(x = id_nuts2, y = gini_n2, fill = id_nuts0)) +
-  geom_bar(stat = "identity") +
-  ggtitle("Gini Coefficient by NUTS 2 Region") +
-  xlab("NUTS 2 Region ID") +
-  ylab("Gini Coefficient") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
+\leavevmode\vadjust pre{\hypertarget{ref-eurostat2023a}{}}%
+---------. 2023c. {``Population on 1 {January} by Broad Age Group, Sex
+and {NUTS} 3 Region.''}
+https://ec.europa.eu/eurostat/databrowser/view/demo\_r\_pjanaggr3/default/table?lang=en.
 
-```{r}
-#| echo: false
+\leavevmode\vadjust pre{\hypertarget{ref-fullart}{}}%
+{``Full Article: The Causal Impact of Economic Growth on Material Use in
+Europe.''} n.d.
+\url{https://www.tandfonline.com/doi/full/10.1080/21606544.2017.1325780}.
 
-# Creating a time-series plot
-ggplot(gini_filtered, aes(x = year, y = gini_n2, color = id_nuts2)) +
-  geom_line() +
-  ggtitle("Time-series Plot of Gini Coefficient by NUTS 2 Region") +
-  xlab("Year") +
-  ylab("Gini Coefficient")
-```
+\leavevmode\vadjust pre{\hypertarget{ref-hasell2023}{}}%
+Hasell, Joe, and Max Roser. 2023. {``Measuring Inequality: {What} Is the
+{Gini} Coefficient?''} \emph{Our World in Data}, July.
 
-```{r}
-#| echo: false
+\leavevmode\vadjust pre{\hypertarget{ref-kiuru2019}{}}%
+Kiuru, Juho, and Tommi Inkinen. 2019. {``E-Capital and Economic Growth
+in European Metropolitan Areas: Applying Social Media Messaging in
+Technology-Based Urban Analysis.''} \emph{Journal of Urban Technology}
+26 (2): 67--88. \url{https://doi.org/10.1080/10630732.2019.1579513}.
 
-# Creating a box plot
-ggplot(gini_filtered, aes(x = id_nuts0, y = GDP_per_capita)) +
-  geom_boxplot() +
-  ggtitle("Distribution of GDP per Capita by Country") +
-  xlab("Country") +
-  ylab("GDP per Capita")
-```
+\leavevmode\vadjust pre{\hypertarget{ref-lessmann2017}{}}%
+Lessmann, Christian, and André Seidel. 2017. {``Regional Inequality,
+Convergence, and Its Determinants -- a View from Outer Space.''}
+\emph{European Economic Review} 92 (February): 110--32.
+\url{https://doi.org/10.1016/j.euroecorev.2016.11.009}.
 
-## Discussion
+\leavevmode\vadjust pre{\hypertarget{ref-lewis}{}}%
+Lewis, Maureen, and Marijn Verhoeven. n.d. {``Financial Crises and
+Social Spending: The Impact of the 2008-2009 Crisis.''}
+\url{https://doi.org/10.2139/ssrn.1670341}.
 
-# References
+\leavevmode\vadjust pre{\hypertarget{ref-regional}{}}%
+{``Regional Inequality in Europe: Evidence, Theory and Policy
+Implications \textbar{} Journal of Economic Geography \textbar{} Oxford
+Academic.''} n.d.
+\url{https://academic.oup.com/joeg/article/19/2/273/4989323}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wooldridge2020}{}}%
+Wooldridge, Jeffrey M. 2020. \emph{Introductory {Econometrics} - {A
+Modern Approach}}. 7th ed. {Cengage}.
+
+\end{CSLReferences}
+
+\end{document}

### 2023-10-22 assignment_2.qmd
Update 22.10

---
title: "Assignment 2: Cross-sectional estimates"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
editor: visual
bibliography: [Econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
---

```{r}
#| label: setup
#| echo: false
#| output: false

library(tidyverse)
library(ggplot2)
library(magrittr)
library(dineq)
library(readxl)
```

```{r}
#| echo: false
#| output: false

# Adding data into dataset

# Data for population
pop_data <- read_csv("Raw_data/demo.gz")

# Data for GDP
gdp_data <- read_csv("Raw_data/nama.gz")
```

# Part A: Growth and Inequity

## 1. Model Estimation

## 2. Model Diagnostics

## 3. Visualization

# Part B: Exploring Other Determinants of Inequity

## 1. Data acquisition

## 2. Multiple Linear Regression Model

## 3. Model Interpretation

### 2023-10-24 MSB104_GR_1_Final_Assignment_research_article.qmd
Final document uploaded

--- 2023-10-22 assignment_1.qmd
@@
 ---
-title: "Assignment 1: Data and Descriptive"
+title: "Regional Inequalities and economic growth; cross section and times series analysis"
 author: "Anine Therese Karlsen & Mona Lisa Jones"
@@
 abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
-Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
+Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations.Furthermore examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
 editor: visual
 bibliography: [Econometrics.bib]
+output-file: Assignment1.qmd
 page-layout: full
@@
 library(tinytex)
+library(dplyr)
 library(ggplot2)
@@
 
-The truth of this statement becomes more evident when taking a closer look in to into sub-national data.
+The truth of this statement becomes more evident when taking a closer look at sub-national data.
 Regional wealth disparities are of prime concern, especially when crafting policies for equitable growth [@lessmann2017].
@@
 
-This assignment is anchored in this very premise.
-It seeks to acquire, process, and analyse sub-national GDP and population data for a selected subset of European countries, namely France, Denmark, Hungary, Portugal, and Slovakia, spanning the period from 2000 to 2020.
-The overarching aim is to calculate GDP per capita for these regions and delve into the intricacies of regional inequity.
-This exploration will be underpinned by various descriptive statistics, visualizations, and analytical techniques to paint a comprehensive picture of regional economic landscapes.
-
-Using time series data observing the GDP per capita and GINI trends across regions and over time, we can discern whether certain areas are ahead or behind in economic performance.
+Using time series data and cross sectional observations we investigate the GDP per capita and GINI trends in Portugal, France, Hungary, Slovenia and Denmark in the time period 2000 to 2020.
+The aims is to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
+Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations
 
@@
 
-"In recent decades, the regional distribution of incomes within countries has attracted considerable interest among academics and policy makers." [@lessmann2017]
+Regional growth and inequality have been important topics of research in Europe for many years.
+In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
+Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
 
-Slovakia, experienced significant economic growth. Their GDP per capita grew faster than the older EU member states, albeit from a lower starting point, leading to a convergence [@fullart].
+When it comes to economic growth, several studies refer to numerous influencing factors.
+@gennaioli2014 and @crescenzi2012 states that regional growth, much like national growth, is significantly influenced by geography and human capital.
+Geography impacts trade, resource availability, and susceptibility to natural events.
+Human capital, encompassing skills and knowledge, drives innovation and productivity.
+Furthermore, @gennaioli2014 mentions that wealthier nations like Denmark in our case, tend to experience faster regional convergence, suggesting national prosperity aids regional growth [@gennaioli2014].
+Convergence implies that poorer regions grow faster than richer ones, gradually reducing disparities in wealth and income.
+Also referred to as beta convergence in @lessmann2017 paper.
 
-The global financial crisis and subsequent European debt crisis had significant impacts on European economies.
-GDP per capita declined in several countries, particularly those hit hardest by the crisis, such as Greece, Spain, and Portugal.
-Some regions took many years to recover to pre-crisis levels, while others, especially in Northern Europe, recovered more quickly [@lewis].
+In the context of Europe, and our selected countries we might expect to see a wide variety of regional economic performances, especially within Hungary and Slovakia and between the western European countries.
+Capital market regulations also play a fundamental role, with effective frameworks promoting quicker regional convergence [@gennaioli2014].
+Taken this in to consideration we might expect a faster growth rate after the joining of the EU in 2003 and 2004.
+That said the opening of the free labor market might also influence the migration of labor in Slovakia and Hungary.
+Not to mention the economic crisis hitting shortly after our eastern European countries joined the union.
 
-Major metropolitan areas, like Paris in France, London in the UK, and Berlin in Germany, often grew faster than other regions in their respective countries, leading to increasing spatial disparities [@kiuru2019].
+Whilst their recent EU membership might have posed certain challenges for Eastern European countries during the financial crisis, it also offered tools and resources that aided in their recovery in Slovakia and Hungary.
+As to the western Europeans such as Denmark, Portugal and France we expect to see a sharp decline in the economic growth.
+Likewise, with the recent covid19 pandemic in terms of financial aspects.
 
-Denmark maintained relatively high levels of GDP per capita throughout this period, benefiting from a diversified economy, strong institutions, and a high degree of economic openness.
-It did face challenges during the 2008 economic crisis but recovered relatively quickly [@dynamics].
+Regarding the effects of the economic crisis, another paper notes that it led to a significant increase in regional unemployment rates across Europe.
+Reporting some regions experiencing unemployment rates as high as 30% [@iammarino2019].
+The authors also discuss how the crisis exposed the vulnerability of regions that were heavily reliant on a single industry or sector, and how sheltered economies were better able to weather the storm [@iammarino2019].
+Another affect that surprised us was that the economic crisis led to increase inequality.
+The study "*the effects of financial crisis on regional inequality*" indicate any type of financial crisis results in higher income inequality [@nguyen2022].
 
-Slovenia, which joined the EU in 2004, showed growth in GDP per capita, though it too was affected by the global economic downturn in 2008.
-However, it managed to recover in the subsequent years [@fullart].
+@lessmann2017 regional inequality study address the importance of studying regional inequality.
+In particular, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
+The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.
+Furthermore, the paper provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
+The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
+@lessmann2017 paper also finds that there is an N-shaped relationship between regional inequality and economic development, which means that regional inequality increases again after the inverted U-shaped pattern has been completed [@lessmann2017].
 
-Regional disparities persisted, with the Île-de-France region (which includes Paris) significantly outperforming other regions in terms of GDP per capita.
-The gap between urban and rural areas also remained a topic of discussion [@regional].
+Lastly, @iammarino2019 discusses regional inequalities in European countries.
+It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium.
+The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019].
+Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
+On the other hand, France has been a subject of discussion and policymaking for many years, regarding its pockets of extreme wealth.
+(finner ikke kilde)
 
-Towards the end of this period, in early 2020, the COVID-19 pandemic posed a new set of challenges for European economies, causing significant contractions in GDP per capita across many regions [@regional].
+# Data
 
-# Part A: Sub-national GDP and GDP per Capita
-
-## Data Acquisition and datasets
-
-The first thing we did in this assignment was to acquire our data.
-Trough Eurostat, we could download the datasets nama_10r_3gdp and demo_r_pjanggr3 as csv files, as well as filter the data by our preferences before downloading it.
-As per the assignment, we filtered the dataset by choosing the years 2000 to 2020.
-Furthermore, we selected the NUTS 3 region for the nations we were given, which were Portugal, France, Hungary, Slovakia and Denmark.
+Trough Eurostat, we download the datasets nama_10r_3gdp and demo_r_pjanggr3 as csv files and filted the data by our preferences before downloading it.
+Furthermore, we filtered the dataset by choosing the years 2000 to 2020.
+Then, we selected the NUTS 3 region for Portugal, France, Hungary, Slovakia and Denmark.
 Finally, we specified the data to be in million Euro.
 
-### GDP (nama_10r_3gdp)
+### GDP
 
-The nama_10r_3gdp dataset from Eurostat provides insights into GDP at regional level using the NUTS classification system.
-It furnishes GDP values in both current prices and adjusted for inflation, with some figures given in purchasing power standards (PPS) to account for price level differences between countries.
-The data is often structured by year and region @eurostat2023.
+The GDP dataset provides insights into GDP at regional level using the NUTS classification system.
+It furnishes GDP values in both current prices and adjusted for inflation, with figures given in purchasing power standards (PPS) to account for price level differences between countries.
+The data is mostly structured by year and region @eurostat2023.
 
 The GDP at market prices represents the final result of production activities of resident producer units within a region or nation.
-It is calculated as the sum of the gross value added across various institutional sectors or industries, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries).
+It is calculated as the sum of the gross value added across various institutional sectors or industries.
+Furthermore, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries) @eurostat2023c.
 This also balances out in the total economy production account.
@@
 
-### Population (demo_r_pjanggr3)
+### Population
 
-Using the NUTS categorization once more, Eurostat's demo_r_pjanggr3 records annual population changes at the regional level.
-This dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender.
-It's also often displayed in a year-by-region format, and the data usually spans in yearly intervals @eurostat2023a.
+Eurostat's records annual population data with NUTS classification.
+Our dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender.
+Displayed in a year-by-region format, with yearly interval updates @eurostat2023a.
 
 Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project.
-The project covers 37 countries and is the central repository for demographic and migration-related data.
-Specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year.
-Additionally, data on marriages, divorces, and migration flows are recorded.
+The project covers 37 countries and is the central repository for demographic and migration-related data @eurostat2023a.
+As well as, specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year @eurostat2023a.
+Additionally, data on marriages, divorces, and migration flows are recorded .
 
@@
 
-The Nomenclature of Territorial Units for Statistics (NUTS) offers a stratified system to segment the economic territory of the EU (including the UK) to facilitate the consistent collection and harmonization of regional statistics across Europe.
-The NUTS regions range from NUTS 0 Country level to NUTS 3 small units such as municipalities level.
+The Nomenclature of Territorial Units for Statistics (NUTS) offers a stratified system to segment the economic territory of the EU and UK to facilitate the consistent collection and harmonization of regional statistics across Europe.
+The NUTS regions range from NUTS 0 Country level to NUTS 3 small units such as municipalities level @eurostat2023a.
 
@@
 ```
+
+# Econometric Approach
+
+In this part we will report and interpret different types of essential descriptive statistics.
+Measuring regional income inequality is challenging due to heterogeneity of regions [@lessmann2017].
+The number of regions in our data set varies largely in size and population.
+Since the focus of this paper is purely growth and changes in inequities over time, the variations of size and population density becomes a minor issue because the country-level territorial heterogeneity is fixed [@lessmann2017].
+
+"Interest in income inequality has led to the development of several ways of measuring it.
+Two types of measures are of interest in this paper----static and dynamic.
+*Static measures* provide a snapshot (cross sectional) of these inequalities at a point of time whereas the dynamic *measures capture historical trends* (Paneldata)*.*" [@wooldridge2020]
+
+By using figures, we can visualize the GDP per capita, and look at how it varies among the different regions. In these figures, a line represent one NUTS 3 region.
+
+### Descriptive statistics
+
+**Mean**
+
+We calculate the mean to provide a representative value for the dataset, facilitating understanding of its central tendency and serving as a benchmark against which deviations and anomalies can be assessed, in later steps when building and interpreting regression models [@wooldridge2020].
+
+**MMR**
+
+Comparing the GRDP (gross regional domestic product) per capita of the region with the highest income to the region with the lowest income (minimum per capita GRDP) provides a measure of the range of these disparities.
+If this measure is small (close to 1), then it would mean that the different regions have relatively equal incomes [@wooldridge2020].
+If this measure is large, then the interpretation is more problematic, as it does not tell us if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers.
+Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to comprehend, and politically powerful measure of regional income inequality.
+
+**Standard diviation (SD)**
+
+Calculating SD to quantify the dispersion or variability of a data set around its mean.
+Helping us assess the degree of uncertainty, variability, or risk associated with an economic variable or parameter, which is crucial for understanding the reliability of estimations and predictions [@wooldridge2020].
+
+**Median**
+
+The median serves as a robust measure of central tendency, especially when a dataset may have outliers or is skewed.
+Unlike the mean, the median is not influenced by extreme values and, thus, can provide a clearer picture of the "typical" value in situations where the data distribution is not symmetrical [@wooldridge2020].
+
+### **Time series (Panel data)**
+
+### Simple linear regression
+
+### Multiple linear regression
+
+# Part 1A: Sub-national GDP and GDP per Capita
 
@@
 
-After calculating the GDP per capita for all NUTS 3 regions in our assigned countries, we can see that there is a large spread between the figures for the various regions.
-In this assignement we want to look at regional inequity; in order to do this in a valuable way we have to divide between the different countries.
-By doing this, we can gain important insights on regional differences that we can utilize, for instance, to discuss national policy on equity and sustainable economic development in regions.
+After calculating the GDP per capita for all NUTS 3 regions countries, we can see that there is a large spread between the figures for the various regions.
+We want to look at regional inequity; in order to do this in a valuable way, countries are divide.
+Furthermore, gaining important insights on regional differences and utilize.
+Later to discuss national policy on equity and sustainable economic development in regions.
 
-### Descriptive statistics
-
-In this part we will report and interpret different types of essential descriptive statistics.
-Measuring regional income inequality is challenging due to heterogeneity of regions.
-The number of regions in our data set varies largely in size and population.
-Since the focus of this paper is purely growth and changes in inequities over time, the variations of size and population density becomes a minor issue because the country-level territorial heterogeneity is fixed.
-
-"Interest in income inequality has led to the development of several ways of measuring it.
-Two types of measures are of interest in this paper----static and dynamic.
-*Static measures* provide a snapshot of these inequalities at a point of time whereas the dynamic *measures capture historical trends.*" (læreboken)
-
-In this part, we'll look at GDP per capita for our assigned countries on a NUTS 3 level.
-In addition, we'll use different kinds of descriptive statistics in order to further 0,37, is analyse this data.
-In this analysis, we'll use @wooldridge2020 for help.
-
-By using figures, we can visualize the GDP per capita, and look at how it varies among the different regions. In these figures, a line represent one NUTS 3 region.
-
-**Mean**
-
-Calculate the mean to provide a representative value for a dataset, facilitating understanding of its central tendency and serving as a benchmark against which deviations and anomalies can be assessed, in later steps when building and interpreting regression models.
-
-**MMR**
-
-A comparison of the GRDP (gross regional domestic product) per capita of the region with the highest income to the region with the lowest income (minimum per capita GRDP) provides a measure of the range of these disparities.
-If this measure is small (close to 1), then it would mean that the different regions have relatively equal incomes.
-If this measure is large, then the interpretation is more problematic, as it does not tell us if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers.
-Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to comprehend, and politically powerful measure of regional income inequality.
-
-**Standard diviation (SD)**
-
-Calculating SD to quantify the dispersion or variability of a data set around its mean.
-Helping us assess the degree of uncertainty, variability, or risk associated with an economic variable or parameter, which is crucial for understanding the reliability of estimations and predictions [@wooldridge2020].
-
-**Median**
-
-The median serves as a robust measure of central tendency, especially when a dataset may have outliers or is skewed.
-Unlike the mean, the median is not influenced by extreme values and, thus, can provide a clearer picture of the "typical" value in situations where the data distribution is not symmetrical [@wooldridge2020].
-
-### Portugal
+#
 
@@
 
-When looking at the figure for France, we can right away see that there are some regions that have a much higher GDP per capita compared to the other regions.
-The region that has the highest GDP per capita for all years is the île-de-France region, one that also includes Paris.
+When looking at the figure for France, we see some regions have a much higher GDP per capita compared to the other regions.
+The regions with the highest GDP per capita for all years is the île-de-France region, one that also includes Paris.
 This significant difference between the regions with the highest GDP per capita and the lowest, shows us that there is a high concentration of economic activity and wealth in a few urban regions.
-Similar to Portugal, we can also again see that there is a difference between urban and rural regions.
+Similar to Portugal, we see differences between urban and rural regions.
 
 Just as in Portugal, there is also a higher mean in France as well.
-Something that is different from the data in France compared to Portugal, is that the standard derivation is higher, and the difference between minimum and maximum is large.
-This strengthens what we have look at earlier in the figure, with some regions having a high concentration of wealth.
+On the contary the data in France has higher standard derivation, and the difference between minimum and maximum is larger.
+This strengthens earlier figures showing, some regions having a high concentration of wealth.
 
@@
 
-In Hungary, most of the regions have similar GDP per Capita. One region that sticks out by having a higher value, is the region of Budapest, Hungary's biggest city.
+In Hungary, most of the regions have similar GDP per Capita. One region that sticks out by having a higher value, is the region of Budapest, the capital.
 
-We have here as well an mean that is larger than the median, high standard derivation, and a large gap between minimum and maximum.
+This case also record the mean as higher vale than the median, high standard derivation, and a large gap between minimum and maximum.
 
@@
 
-In Slovakia as well, we have one region that has a much higher GDP per capita than the rest of the regions.
-This region is Bratislava, which is the biggest city and capital, something that might point to this city being the economic capital of Slovakia as well.
+Additionally, Slovakia, also have one region with much higher GDP per capita than the rest of the regions.
+This region is Bratislava, which is the biggest city and the capital, something that might point to this city being the economic centre of Slovakia as well.
 
-Slovakia has also a mean higher that the median, and a large gap between minimum and maximum.
-In addition, the standard derivation is pretty high, meaning that there is some regions (or one region in this case) that is far away from the rest of the regions when it comes to economic development.
+Slovakia follows the trend with higher mean than the median, and a large gap between minimum and maximum.
+In addition, the standard derivation is significantly high, meaning that there is some regions (or one region in this case) that is further away from the rest of the regions in terms of economic development.
 
@@
 
-Lastly, we have Denmark. We can see similar pattern here as well, with the capital Copenhagen being one of the regions with the highest GDP per capita.
+Lastly, we see similar pattern in Denmark, with the capital Copenhagen being one of the regions with the highest GDP per capita.
 
-We can also see the same as the previous countries, with the mean being higher than the median, which shows us that regions like Copenhagen might drag the mean up by being much larger than the rest of the regions.
+Whith mean higher than the median, showing that regions like Copenhagen possible dragging the mean up by population volume.
 
-# Part B: Regional Inequity
+# Part 1B: Regional Inequity
 
@@
 
-## Discussion
+# Part 2A: Growth and Inequity
+
+### **Data acquisition**
+
+Firstly, the data is summerised for the year 2010 at the NUTS2 regional level, focusing on economic metrics.
+Processing the dataset containing the NUTS3 level.
+Thereafter, computing the total GDP and total population for each NUTS2 region and year.
+Then, derive the GDP per capita at the NUTS2 level by dividing the total GDP by the total population again.
+Furthermore, narrowing down the dataset to observations from the year 2010 where the Gini coefficient is positive.
+Then adding, logarithmic transformations to linearize relationships or to reduce the impact of extreme values.
+
+The next step is making a data frame of NUTS2 2010 and grouping by NUTS0 country level variables, as well as allocating NUTS2 to selected countries.
+Lastly presenting data in **table XXX** showing a count of NUTS2 regions within each top-level region or country in the 2010 data.
+
+**#Esempel:Looking at the descriptive statistics we see in average a very low level of inequality within all regions.
+Even at the maximum inequality is modest with 0.35 spacial in comparison with other countries, see for example Lessmann and Seidel ([2017](https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course%20files/html/Assignments_example.html?context_id=10747~22507&context_type=Course&download=1&id=107470000002033066&inline=1#ref-lessmann2017regional)) for such a comparison.Studding the temporal component of our data we see that most regions follow a common trend in there level of inequality with modest changes over time.
+See for example the NUTS2 regions of France in [Figure 4](https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course%20files/html/Assignments_example.html?context_id=10747~22507&context_type=Course&download=1&id=107470000002033066&inline=1#fig-regional_gini_time).
+There are however some severe exceptions ...#**
+
+The next step is, estimating a Linear Regression for All Countries using cross sectional data from 2010, NUTS2.
+
+### **Cross Sectional Analysis**
+
+With cross-sectional data analysis we create a snapshot of the year 2010.
+Cross-sectional data is simpler to manage and interpret than time-series or panel data.
+With data from only one time point, we avoid complications arising from temporal dynamics.
+Cross-sectional data allows for the comparison of different regions at the same time, which can be crucial for identifying disparities or differences between the regions [@wooldridge2020].
+
+### **Simple linear regression model**
+
+We use simple linear regression to model the relationship with the GINI as the dependent variable and the natural logarithm of GDP per capita as the independent variable.
+Capturing the relationship between regional development and regional inequality for all regions in 2010.
+
+**Model assumptions**
+
+"The relationship between our dependent and independent variables is linear, ensuring a clear and direct connection between them.
+Each observation operates independently of the others, emphasizing the unique contribution of every data point.
+Additionally, we expect homoscedasticity, implying that the variance of the residuals remains consistent regardless of the independent variable's level.
+It's also crucial that, for any specified value of X, Y maintains a normal distribution.
+And although more pertinent to multiple regression, it's worth noting the absence of multicollinearity, ensuring that no two predictors are closely correlated."
+
+**Model specification**
+
+**###Vis formel##**
+
+Y is the dependent variable (hva som perdikeres).
+
+X is the independent variable (input).
+
+β0​ is the intercept.
+
+β1​ is the slope of the line.
+
+ε represents the residuals or error in the prediction.
+
+-   **Intercept (**β0​): represents the value of Y when X is 0.
+
+    **Slope (**β1​): Indicates the change in Y for a one-unit change in X.
+
+"In linear regression, the objective is straightforward: determine the line that best represents the observed data.
+Furthermore, this 'goodness of fit' line is characterized by coefficients β0​ and β1, which represent the intercept and slope, respectively [@wooldridge2020].
+The optimal line is one that minimizes the sum of the squared differences between the observed values and the predictions made by the model [@wooldridge2020].
+This approach, known as the Ordinary Least Squares (OLS), has a solid mathematical foundation [@wooldridge2020].
+By taking the sum of these squared differences and setting their partial derivatives with respect to β0​ and β1 to zero, we can derive the formula that give us these coefficients.
+And once these coefficients are in hand, they provide a direct equation for the line that best fits our data set [@wooldridge2020].
+One of the strengths of OLS is its closed-form solution, meaning we can directly compute the coefficients from the data without iterative procedures [@wooldridge2020].
+Moreover, under the foundamental assumptions of linear regression, OLS stands out as the Best Linear Unbiased Estimator (BLUE), emphasizing its reliability and efficiency in estimating the true underlying relationship [@wooldridge2020]."
+
+### Individual country cases
+
+#### Portugal
+
+#### Denmark
+
+#### France
+
+#### Hungary
+
+#### Slovakia
+
+### **Outlier discussion**
+
+**##Eksepel....When looking at the data it is obvious that we have too many regions with zero inequality.
+This seems strange.
+The reason for this is .....
+In the subsequent analysis it therefore would make scene to exclude these observations to not bias our finding.**
+
+**Looking at specific countries we observe allso further abnormalities like in ....
+where ...
+As thees outliers are not do to a bias in our data we shold keep them in our data.
+However we need to test later the sensitivity of our findings regarding the presences of these borderline observations.
+...####**
+
+# Part 2B: Exploring Other Determinants of Inequity
+
+### I. Data Acquisition
+
+### II. Multiple Linear Regression Model
+
+### III. Model Interpretation
+
+# Discussion
 

### 2023-10-24 MSB104_GR_1_Final_Assignment_research_article.qmd
litteratur liste

@@
 
+### Cross sectional analysis
+
 ### Simple linear regression
@@
 
-After calculating the GDP per capita for all NUTS 3 regions countries, we can see that there is a large spread between the figures for the various regions.
+After calculating the GDP per capita for all NUTS 3 regions within the chosen countries, we can see that there is a large spread between the figures for the various regions.
 We want to look at regional inequity; in order to do this in a valuable way, countries are divide.
@@
 
-#
+### Portugal
 

### 2023-10-24 MSB104_GR_1_Final_Assignment_research_article.qmd
litteratur igjen

@@
 editor: visual
-bibliography: [Econometrics.bib]
+bibliography: [Regional_inequality_23]
 output-file: Assignment1.qmd

### 2023-11-01 MSB104_GR_1_Final_Assignment_research_article.qmd
code

--- 2023-10-24 MSB104_GR_1_Final_Assignment_research_article.qmd
@@
 editor: visual
-bibliography: [Econometrics.bib]
+bibliography: [Regional_inequality_23.bib]
 output-file: Assignment1.qmd

### 2023-11-01 MSB104_GR_1_Final_Assignment_research_article.tex
code

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  a4paper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Regional Inequalities and economic growth; cross section and times series analysis},
  pdfauthor={Anine Therese Karlsen \& Mona Lisa Jones},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Regional Inequalities and economic growth; cross section and
times series analysis}
\author{Anine Therese Karlsen \& Mona Lisa Jones}
\date{}

\begin{document}
\maketitle
\begin{abstract}
This assignment aims to acquire, process, and analyze sub-national GDP
and population data for a subset of European countries. Calculate GDP
per capita and explore regional inequity using various descriptive
statistics and visualizations.Furthermore examining the relationship
between regional development and inequality, employing cross sectional
estimation techniques for the year 2010.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

While national GDP and GDP per capita are vital indicators of a
country's aggregate economic health, they do not shed light on how
wealth or income is distributed among its residents. A high national GDP
can, paradoxically, coexist with pockets of regional deprivation
(\textbf{lessmann2017?}).

The truth of this statement becomes more evident when taking a closer
look at sub-national data. Regional wealth disparities are of prime
concern, especially when crafting policies for equitable growth
(\textbf{lessmann2017?}). A country's macro-level prosperity does not
automatically guarantee that all its regions partake equally in this
wealth. By studying smaller regions within a country, it is possible to
get a more nuanced narrative about the state of regional economic
disparities (\textbf{lessmann2017?}).

Using time series data and cross sectional observations we investigate
the GDP per capita and GINI trends in Portugal, France, Hungary,
Slovenia and Denmark in the time period 2000 to 2020. The aims is to
acquire, process, and analyze sub-national GDP and population data for a
subset of European countries. Calculate GDP per capita and explore
regional inequity using various descriptive statistics and
visualizations

\hypertarget{literature-review}{%
\section{Literature review}\label{literature-review}}

Regional growth and inequality have been important topics of research in
Europe for many years. In recent decades, the European Union has
implemented policies aimed at reducing regional disparities and
promoting economic growth across its member states. Furthermore, despite
these efforts, significant disparities in income and economic
development persist across regions in Europe (\textbf{lessmann2017?}).

When it comes to economic growth, several studies refer to numerous
influencing factors. (\textbf{gennaioli2014?}) and
(\textbf{crescenzi2012?}) states that regional growth, much like
national growth, is significantly influenced by geography and human
capital. Geography impacts trade, resource availability, and
susceptibility to natural events. Human capital, encompassing skills and
knowledge, drives innovation and productivity. Furthermore,
(\textbf{gennaioli2014?}) mentions that wealthier nations like Denmark
in our case, tend to experience faster regional convergence, suggesting
national prosperity aids regional growth (\textbf{gennaioli2014?}).
Convergence implies that poorer regions grow faster than richer ones,
gradually reducing disparities in wealth and income. Also referred to as
beta convergence in (\textbf{lessmann2017?}) paper.

In the context of Europe, and our selected countries we might expect to
see a wide variety of regional economic performances, especially within
Hungary and Slovakia and between the western European countries. Capital
market regulations also play a fundamental role, with effective
frameworks promoting quicker regional convergence
(\textbf{gennaioli2014?}). Taken this in to consideration we might
expect a faster growth rate after the joining of the EU in 2003 and
2004. That said the opening of the free labor market might also
influence the migration of labor in Slovakia and Hungary. Not to mention
the economic crisis hitting shortly after our eastern European countries
joined the union.

Whilst their recent EU membership might have posed certain challenges
for Eastern European countries during the financial crisis, it also
offered tools and resources that aided in their recovery in Slovakia and
Hungary. As to the western Europeans such as Denmark, Portugal and
France we expect to see a sharp decline in the economic growth.
Likewise, with the recent covid19 pandemic in terms of financial
aspects.

Regarding the effects of the economic crisis, another paper notes that
it led to a significant increase in regional unemployment rates across
Europe. Reporting some regions experiencing unemployment rates as high
as 30\% (\textbf{iammarino2019?}). The authors also discuss how the
crisis exposed the vulnerability of regions that were heavily reliant on
a single industry or sector, and how sheltered economies were better
able to weather the storm (\textbf{iammarino2019?}). Another affect that
surprised us was that the economic crisis led to increase inequality.
The study ``\emph{the effects of financial crisis on regional
inequality}'' indicate any type of financial crisis results in higher
income inequality (\textbf{nguyen2022?}).

(\textbf{lessmann2017?}) regional inequality study address the
importance of studying regional inequality. In particular, its potential
consequences, such as political tensions that can undermine social and
political stability (\textbf{lessmann2017?}). The paper also discusses
the relationship between regional inequality and personal income
inequality and conflict risk. Furthermore, the paper provides us with
insights into the inverted U-shaped relationship between regional
inequality and the level of economic development in different country
groups (\textbf{lessmann2017?}). The inverted U-shape also referred to
as Kuznets Curve suggests that as an economy develops, regional
inequality initially rises, reaches a peak, and then starts to decline.
(\textbf{lessmann2017?}) paper also finds that there is an N-shaped
relationship between regional inequality and economic development, which
means that regional inequality increases again after the inverted
U-shaped pattern has been completed (\textbf{lessmann2017?}).

Lastly, (\textbf{iammarino2019?}) discusses regional inequalities in
European countries. It presents evidence that inter-regional inequality
has been increasing in the European Union since the turn of the
millennium. The authors argue that this is due to the existence of
several groups of regional economies in Europe that are structurally
very different from one another (\textbf{iammarino2019?}). Additionally,
the authors argue that there are some countries in the EU that are more
evenly developed than others, and that a map of under performance or
over-performance means less in a high-income but evenly developed
country, such as in Denmark in our case (\textbf{iammarino2019?}). On
the other hand, France has been a subject of discussion and policymaking
for many years, regarding its pockets of extreme wealth. (finner ikke
kilde)

\hypertarget{data}{%
\section{Data}\label{data}}

Trough Eurostat, we download the datasets nama\_10r\_3gdp and
demo\_r\_pjanggr3 as csv files and filted the data by our preferences
before downloading it. Furthermore, we filtered the dataset by choosing
the years 2000 to 2020. Then, we selected the NUTS 3 region for
Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified
the data to be in million Euro.

\hypertarget{gdp}{%
\subsubsection{GDP}\label{gdp}}

The GDP dataset provides insights into GDP at regional level using the
NUTS classification system. It furnishes GDP values in both current
prices and adjusted for inflation, with figures given in purchasing
power standards (PPS) to account for price level differences between
countries. The data is mostly structured by year and region
(\textbf{eurostat2023?}).

The GDP at market prices represents the final result of production
activities of resident producer units within a region or nation. It is
calculated as the sum of the gross value added across various
institutional sectors or industries. Furthermore, augmented by taxes and
reduced by subsidies on products (which are not allocated to specific
sectors or industries) (\textbf{eurostat2023c?}). This also balances out
in the total economy production account. In terms of methodology, while
national accounts compile GDP from the expenditure side, regional
accounts don't adopt this approach due to the complexities of accurately
mapping inter-regional flows of goods and services.

The different measures for the regional GDP are absolute figures in €
and Purchasing Power Standards (PPS), figures per inhabitant and
relative data compared to the EU Member States average
(\textbf{eurostat2023c?}).

\hypertarget{population}{%
\subsubsection{Population}\label{population}}

Eurostat's records annual population data with NUTS classification. Our
dataset includes information on births, deaths, net migration, and may
also include demographic information on age and gender. Displayed in a
year-by-region format, with yearly interval updates
(\textbf{eurostat2023a?}).

Eurostat's primary source for yearly demographic data at the regional
level stems from the Unified Demography (Unidemo) project. The project
covers 37 countries and is the central repository for demographic and
migration-related data (\textbf{eurostat2023a?}). As well as, specific
metrics gathered under UNIDEMO encompass population counts at the close
of the calendar year and events such as births and deaths occurring
within that year (\textbf{eurostat2023a?}). Additionally, data on
marriages, divorces, and migration flows are recorded .

For the purpose of this research, the demographic data references the
NUTS 2016 classification, which provides a detailed breakdown of the
European Union's territory (\textbf{eurostat2021?}).

\hypertarget{nuts-classification}{%
\subsubsection{NUTS classification}\label{nuts-classification}}

The Nomenclature of Territorial Units for Statistics (NUTS) offers a
stratified system to segment the economic territory of the EU and UK to
facilitate the consistent collection and harmonization of regional
statistics across Europe. The NUTS regions range from NUTS 0 Country
level to NUTS 3 small units such as municipalities level
(\textbf{eurostat2023a?}). ~

\hypertarget{econometric-approach}{%
\section{Econometric Approach}\label{econometric-approach}}

In this part we will report and interpret different types of essential
descriptive statistics. Measuring regional income inequality is
challenging due to heterogeneity of regions (\textbf{lessmann2017?}).
The number of regions in our data set varies largely in size and
population. Since the focus of this paper is purely growth and changes
in inequities over time, the variations of size and population density
becomes a minor issue because the country-level territorial
heterogeneity is fixed (\textbf{lessmann2017?}).

``Interest in income inequality has led to the development of several
ways of measuring it. Two types of measures are of interest in this
paper----static and dynamic. \emph{Static measures} provide a snapshot
(cross sectional) of these inequalities at a point of time whereas the
dynamic \emph{measures capture historical trends} (Paneldata)\emph{.}''
(\textbf{wooldridge2020?})

By using figures, we can visualize the GDP per capita, and look at how
it varies among the different regions. In these figures, a line
represent one NUTS 3 region.

\hypertarget{descriptive-statistics}{%
\subsubsection{Descriptive statistics}\label{descriptive-statistics}}

\textbf{Mean}

We calculate the mean to provide a representative value for the dataset,
facilitating understanding of its central tendency and serving as a
benchmark against which deviations and anomalies can be assessed, in
later steps when building and interpreting regression models
(\textbf{wooldridge2020?}).

\textbf{MMR}

Comparing the GRDP (gross regional domestic product) per capita of the
region with the highest income to the region with the lowest income
(minimum per capita GRDP) provides a measure of the range of these
disparities. If this measure is small (close to 1), then it would mean
that the different regions have relatively equal incomes
(\textbf{wooldridge2020?}). If this measure is large, then the
interpretation is more problematic, as it does not tell us if the high
ratio is due to substantial variation in the distribution of per capita
GDRPs or the presence of outliers. Nevertheless, maximum to minimum
ratio (MMR) provides a quick, easy to comprehend, and politically
powerful measure of regional income inequality.

\textbf{Standard diviation (SD)}

Calculating SD to quantify the dispersion or variability of a data set
around its mean. Helping us assess the degree of uncertainty,
variability, or risk associated with an economic variable or parameter,
which is crucial for understanding the reliability of estimations and
predictions (\textbf{wooldridge2020?}).

\textbf{Median}

The median serves as a robust measure of central tendency, especially
when a dataset may have outliers or is skewed. Unlike the mean, the
median is not influenced by extreme values and, thus, can provide a
clearer picture of the ``typical'' value in situations where the data
distribution is not symmetrical (\textbf{wooldridge2020?}).

\hypertarget{time-series-panel-data}{%
\subsubsection{\texorpdfstring{\textbf{Time series (Panel
data)}}{Time series (Panel data)}}\label{time-series-panel-data}}

\hypertarget{cross-sectional-analysis}{%
\subsubsection{Cross sectional
analysis}\label{cross-sectional-analysis}}

\hypertarget{simple-linear-regression}{%
\subsubsection{Simple linear
regression}\label{simple-linear-regression}}

\hypertarget{multiple-linear-regression}{%
\subsubsection{Multiple linear
regression}\label{multiple-linear-regression}}

\hypertarget{part-1a-sub-national-gdp-and-gdp-per-capita}{%
\section{Part 1A: Sub-national GDP and GDP per
Capita}\label{part-1a-sub-national-gdp-and-gdp-per-capita}}

\hypertarget{gdp-per-capita-calculation}{%
\subsection{GDP per Capita
Calculation}\label{gdp-per-capita-calculation}}

The formula for calculating GDP per Capita is as follows:

\(y_i=GDP_i/population_i\)

After calculating the GDP per capita for all NUTS 3 regions within the
chosen countries, we can see that there is a large spread between the
figures for the various regions. We want to look at regional inequity;
in order to do this in a valuable way, countries are divide.
Furthermore, gaining important insights on regional differences and
utilize. Later to discuss national policy on equity and sustainable
economic development in regions.

\hypertarget{portugal}{%
\subsubsection{Portugal}\label{portugal}}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-7-1.pdf}

\begin{verbatim}
        GDP_per_capita
mean         1.4185524
median       1.3900000
std_dev      0.3702905
minimum      0.6500000
maximum      2.7200000
\end{verbatim}

By looking at figure for Portugal, we can see that the GDP per capita in
Portugal's regions appears to be fairly consistent. There is however
some regional variability. We can see that the regions around the big
cities like Lisbon have a higher GDP per capita compared to some more
rural areas. Since Lisbon is the capital of Portugal, there is probably
a higher concentration of industries, making it a economic center (which
again makes the GDP per capita higher).

To continue, we can see that the mean is a little higher that the
median, something that might indicate that regions like Lisbon are
pulling up the average. If we compare the standard derivation for
Portugal with the other countries, we'll see that is fairly low in
comparison. This might mean that there is not a lot of variability
between the GDP per capita across different regions in Portugal. The gap
between minimum and maximum is also low compared to other countries,
something that'll also show us that the economic disparity in Portugal
might not be as high as it is in other countries.

\hypertarget{france}{%
\subsubsection{France}\label{france}}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-10-1.pdf}

\begin{verbatim}
        GDP_per_capita
mean          2.630444
median        2.440000
std_dev       1.053767
minimum       0.830000
maximum      11.580000
\end{verbatim}

When looking at the figure for France, we see some regions have a much
higher GDP per capita compared to the other regions. The regions with
the highest GDP per capita for all years is the île-de-France region,
one that also includes Paris. This significant difference between the
regions with the highest GDP per capita and the lowest, shows us that
there is a high concentration of economic activity and wealth in a few
urban regions. Similar to Portugal, we see differences between urban and
rural regions.

Just as in Portugal, there is also a higher mean in France as well. On
the contary the data in France has higher standard derivation, and the
difference between minimum and maximum is larger. This strengthens
earlier figures showing, some regions having a high concentration of
wealth.

\hypertarget{hungary}{%
\subsubsection{Hungary}\label{hungary}}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-13-1.pdf}

\begin{verbatim}
        GDP_per_capita
mean         0.8598000
median       0.7650000
std_dev      0.4059723
minimum      0.3200000
maximum      3.1300000
\end{verbatim}

In Hungary, most of the regions have similar GDP per Capita. One region
that sticks out by having a higher value, is the region of Budapest, the
capital.

This case also record the mean as higher vale than the median, high
standard derivation, and a large gap between minimum and maximum.

\hypertarget{slovakia}{%
\subsubsection{Slovakia}\label{slovakia}}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-16-1.pdf}

\begin{verbatim}
        GDP_per_capita
mean         1.2501948
median       1.0950000
std_dev      0.8018259
minimum      0.3000000
maximum      4.0200000
\end{verbatim}

Additionally, Slovakia, also have one region with much higher GDP per
capita than the rest of the regions. This region is Bratislava, which is
the biggest city and the capital, something that might point to this
city being the economic centre of Slovakia as well.

Slovakia follows the trend with higher mean than the median, and a large
gap between minimum and maximum. In addition, the standard derivation is
significantly high, meaning that there is some regions (or one region in
this case) that is further away from the rest of the regions in terms of
economic development.

\hypertarget{denmark}{%
\subsubsection{Denmark}\label{denmark}}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-19-1.pdf}

\begin{verbatim}
        GDP_per_capita
mean          4.419221
median        4.000000
std_dev       1.343933
minimum       2.760000
maximum       8.420000
\end{verbatim}

Lastly, we see similar pattern in Denmark, with the capital Copenhagen
being one of the regions with the highest GDP per capita.

Whith mean higher than the median, showing that regions like Copenhagen
possible dragging the mean up by population volume.

\hypertarget{part-1b-regional-inequity}{%
\section{Part 1B: Regional Inequity}\label{part-1b-regional-inequity}}

\hypertarget{gini-coefficient-calculation}{%
\subsection{Gini Coefficient
Calculation}\label{gini-coefficient-calculation}}

With the use of the NUTS3 GDP per capita data and this formula:

\(GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|\)

we will compute the population-weighted GDP Gini coefficient for each
European NUTS2 region in our assigned countries.

The gini coefficient can help us measure inequality in a distribution,
as is therefore a useful tool for us to use when we look at regional
inequity. The closer the gini coefficient is to 1, the bigger the
inequality is; a number closer to 0 equals equality. When looking at the
gini coefficient for NUTS 2 regions, we also get a better overview over
differences in income between different regions, and it also makes it
easier to find the reasons as to why there is a difference between the
regions (\textbf{hasell2023?}).

After calculating the gini coefficients, we can see that there are some
similarities to the data we got from GDP per capita for NUTS 3 regions.
In order to see these similarities better, as well as look for other
important aspects that can be provided trough the calculations, we will
visualize the data in three different ways.

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-23-1.pdf}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-24-1.pdf}

\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-25-1.pdf}

\hypertarget{part-2a-growth-and-inequity}{%
\section{Part 2A: Growth and
Inequity}\label{part-2a-growth-and-inequity}}

\hypertarget{data-acquisition}{%
\subsubsection{\texorpdfstring{\textbf{Data
acquisition}}{Data acquisition}}\label{data-acquisition}}

Firstly, the data is summerised for the year 2010 at the NUTS2 regional
level, focusing on economic metrics. Processing the dataset containing
the NUTS3 level. Thereafter, computing the total GDP and total
population for each NUTS2 region and year. Then, derive the GDP per
capita at the NUTS2 level by dividing the total GDP by the total
population again. Furthermore, narrowing down the dataset to
observations from the year 2010 where the Gini coefficient is positive.
Then adding, logarithmic transformations to linearize relationships or
to reduce the impact of extreme values.

The next step is making a data frame of NUTS2 2010 and grouping by NUTS0
country level variables, as well as allocating NUTS2 to selected
countries. Lastly presenting data in \textbf{table XXX} showing a count
of NUTS2 regions within each top-level region or country in the 2010
data.

\textbf{\#Esempel:Looking at the descriptive statistics we see in
average a very low level of inequality within all regions. Even at the
maximum inequality is modest with 0.35 spacial in comparison with other
countries, see for example Lessmann and Seidel
(\href{https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course\%20files/html/Assignments_example.html?context_id=10747~22507\&context_type=Course\&download=1\&id=107470000002033066\&inline=1\#ref-lessmann2017regional}{2017})
for such a comparison.Studding the temporal component of our data we see
that most regions follow a common trend in there level of inequality
with modest changes over time. See for example the NUTS2 regions of
France in
\href{https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course\%20files/html/Assignments_example.html?context_id=10747~22507\&context_type=Course\&download=1\&id=107470000002033066\&inline=1\#fig-regional_gini_time}{Figure~4}.
There are however some severe exceptions \ldots\#}

The next step is, estimating a Linear Regression for All Countries using
cross sectional data from 2010, NUTS2.

\hypertarget{cross-sectional-analysis-1}{%
\subsubsection{\texorpdfstring{\textbf{Cross Sectional
Analysis}}{Cross Sectional Analysis}}\label{cross-sectional-analysis-1}}

With cross-sectional data analysis we create a snapshot of the year
2010. Cross-sectional data is simpler to manage and interpret than
time-series or panel data. With data from only one time point, we avoid
complications arising from temporal dynamics. Cross-sectional data
allows for the comparison of different regions at the same time, which
can be crucial for identifying disparities or differences between the
regions (\textbf{wooldridge2020?}).

\hypertarget{simple-linear-regression-model}{%
\subsubsection{\texorpdfstring{\textbf{Simple linear regression
model}}{Simple linear regression model}}\label{simple-linear-regression-model}}

We use simple linear regression to model the relationship with the GINI
as the dependent variable and the natural logarithm of GDP per capita as
the independent variable. Capturing the relationship between regional
development and regional inequality for all regions in 2010.

\textbf{Model assumptions}

``The relationship between our dependent and independent variables is
linear, ensuring a clear and direct connection between them. Each
observation operates independently of the others, emphasizing the unique
contribution of every data point. Additionally, we expect
homoscedasticity, implying that the variance of the residuals remains
consistent regardless of the independent variable's level. It's also
crucial that, for any specified value of X, Y maintains a normal
distribution. And although more pertinent to multiple regression, it's
worth noting the absence of multicollinearity, ensuring that no two
predictors are closely correlated.''

\textbf{Model specification}

\textbf{\#\#\#Vis formel\#\#}

Y is the dependent variable (hva som perdikeres).

X is the independent variable (input).

β0\hspace{0pt} is the intercept.

β1\hspace{0pt} is the slope of the line.

ε represents the residuals or error in the prediction.

\begin{itemize}
\item
  \textbf{Intercept (}β0\hspace{0pt}): represents the value of Y when X
  is 0.

  \textbf{Slope (}β1\hspace{0pt}): Indicates the change in Y for a
  one-unit change in X.
\end{itemize}

``In linear regression, the objective is straightforward: determine the
line that best represents the observed data. Furthermore, this `goodness
of fit' line is characterized by coefficients β0\hspace{0pt} and β1,
which represent the intercept and slope, respectively
(\textbf{wooldridge2020?}). The optimal line is one that minimizes the
sum of the squared differences between the observed values and the
predictions made by the model (\textbf{wooldridge2020?}). This approach,
known as the Ordinary Least Squares (OLS), has a solid mathematical
foundation (\textbf{wooldridge2020?}). By taking the sum of these
squared differences and setting their partial derivatives with respect
to β0\hspace{0pt} and β1 to zero, we can derive the formula that give us
these coefficients. And once these coefficients are in hand, they
provide a direct equation for the line that best fits our data set
(\textbf{wooldridge2020?}). One of the strengths of OLS is its
closed-form solution, meaning we can directly compute the coefficients
from the data without iterative procedures (\textbf{wooldridge2020?}).
Moreover, under the foundamental assumptions of linear regression, OLS
stands out as the Best Linear Unbiased Estimator (BLUE), emphasizing its
reliability and efficiency in estimating the true underlying
relationship (\textbf{wooldridge2020?}).''

\hypertarget{individual-country-cases}{%
\subsubsection{Individual country
cases}\label{individual-country-cases}}

\hypertarget{portugal-1}{%
\paragraph{Portugal}\label{portugal-1}}

\hypertarget{denmark-1}{%
\paragraph{Denmark}\label{denmark-1}}

\hypertarget{france-1}{%
\paragraph{France}\label{france-1}}

\hypertarget{hungary-1}{%
\paragraph{Hungary}\label{hungary-1}}

\hypertarget{slovakia-1}{%
\paragraph{Slovakia}\label{slovakia-1}}

\hypertarget{outlier-discussion}{%
\subsubsection{\texorpdfstring{\textbf{Outlier
discussion}}{Outlier discussion}}\label{outlier-discussion}}

\textbf{\#\#Eksepel\ldots.When looking at the data it is obvious that we
have too many regions with zero inequality. This seems strange. The
reason for this is \ldots.. In the subsequent analysis it therefore
would make scene to exclude these observations to not bias our finding.}

\textbf{Looking at specific countries we observe allso further
abnormalities like in \ldots. where \ldots{} As thees outliers are not
do to a bias in our data we shold keep them in our data. However we need
to test later the sensitivity of our findings regarding the presences of
these borderline observations. \ldots\#\#\#\#}

\hypertarget{part-2b-exploring-other-determinants-of-inequity}{%
\section{Part 2B: Exploring Other Determinants of
Inequity}\label{part-2b-exploring-other-determinants-of-inequity}}

\hypertarget{i.-data-acquisition}{%
\subsubsection{I. Data Acquisition}\label{i.-data-acquisition}}

\hypertarget{ii.-multiple-linear-regression-model}{%
\subsubsection{II. Multiple Linear Regression
Model}\label{ii.-multiple-linear-regression-model}}

\hypertarget{iii.-model-interpretation}{%
\subsubsection{III. Model
Interpretation}\label{iii.-model-interpretation}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}

### 2023-11-01 assignment_1.qmd
code

@@
 
+...
+
 ```{r}
@@
 ```
+
+...
 
@@
 
+...
+
 ## Discussion
 
-# References
+What does the GINI really tell us? (Look at paper from page 116)
+
+# Assignment 2: Cross-sectional Estimates
+
+## 2A: Growth and Inequity
+
+### 1. Model Estimation
+
+```{r}
+data2010_nuts2 <- gini_nuts_2 %>%
+ group_by(id_nuts2, year) %>%
+ mutate( gdp_n2 = sum(GDP, na.rm = T),
+ pop_n2 = sum(population, na.rm = T),
+GDP_per_capita = gdp_n2/pop_n2) %>%
+ ungroup() %>%
+ distinct(year, id_nuts0, id_nuts2, GDP_per_capita, gini_n2, gdp_n2, pop_n2) %>% 
+  filter(year == 2010, gini_n2 > 0)
+
+data2010_nuts2 <- data2010_nuts2 %>% 
+  mutate(log_GDP_pc = log(GDP_per_capita))
+```

### 2023-11-01 MSB104_GR_1_Final_Assignment_research_article.qmd
mer text

--- 2023-10-24 MSB104_GR_1_Final_Assignment_research_article.qmd
@@
 abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
-Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations.Furthermore examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
+Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
 editor: visual
-bibliography: [Econometrics.bib]
+bibliography: [Regional_inequality_23.bib]
 output-file: Assignment1.qmd
@@
 The aims is to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
-Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations
 
-# Literature review
+The primary objective is to gain knowledge of econometric methods used for science in business through our given topic economic growth and inequality measures.
+Both GDP per capita and the Gini coefficient are vital indicators in the field of economics and are frequently used in a wide array of studies to analyse and interpret the economic conditions and social disparities of different regions or countries.
+Their analysis provides crucial insights for policymakers, economists, and researchers in shaping and evaluating the effectiveness of socio-economic policies.
 
-Regional growth and inequality have been important topics of research in Europe for many years.
+The research question ***"What are the regional economic growth and inequality measurements from 2000 to 2020, in selected European countries and how does transport infrastructure affect the results?"*** is key in guiding our study.
+Setting a solid foundation for a comprehensive study that not only measure our given parameters, but also critically examines the role of transportation infrastructure play in these dyniamics.
+@lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs as a determinant of regional inequality transportation costs have the potential to reduce regional inequalities.
+
+The sructure of our study is methodically designed to facilitate this exploration.
+In part 1A and B we calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations such as panel data (***Assignment_1***).
+
+In part 2A and B we use cross sectional method to examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).
+
+# Literature Review: Regional Economic Growth, Inequality, and Transport Infrastructure (2000-2020)
+
+Why some places grow and prosper is a fundamental question in social science research.
+This question motivated Adam Smith\'s work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A.
+(1863).
+An inquiry into the nature and causes of the wealth of nations.
+Edinburgh: A.
+& C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*.
+Furthermore,* regional growth and inequality have been important topics of research in Europe for many years.
 In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
@@
 
-When it comes to economic growth, several studies refer to numerous influencing factors.
-@gennaioli2014 and @crescenzi2012 states that regional growth, much like national growth, is significantly influenced by geography and human capital.
-Geography impacts trade, resource availability, and susceptibility to natural events.
-Human capital, encompassing skills and knowledge, drives innovation and productivity.
-Furthermore, @gennaioli2014 mentions that wealthier nations like Denmark in our case, tend to experience faster regional convergence, suggesting national prosperity aids regional growth [@gennaioli2014].
-Convergence implies that poorer regions grow faster than richer ones, gradually reducing disparities in wealth and income.
-Also referred to as beta convergence in @lessmann2017 paper.
+**Economic Growth and Regional Disparities**
 
-In the context of Europe, and our selected countries we might expect to see a wide variety of regional economic performances, especially within Hungary and Slovakia and between the western European countries.
-Capital market regulations also play a fundamental role, with effective frameworks promoting quicker regional convergence [@gennaioli2014].
-Taken this in to consideration we might expect a faster growth rate after the joining of the EU in 2003 and 2004.
-That said the opening of the free labor market might also influence the migration of labor in Slovakia and Hungary.
-Not to mention the economic crisis hitting shortly after our eastern European countries joined the union.
+Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
+Lessmann and Seidel\'s research offers insights into these dynamics, highlighting the persistence of income disparities across regions despite overarching economic advancement [@lessmann2017].
+Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.\'s findings [@gennaioli2014].
+However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia showcasing different growth trajectories, particularly after their EU accession.
 
-Whilst their recent EU membership might have posed certain challenges for Eastern European countries during the financial crisis, it also offered tools and resources that aided in their recovery in Slovakia and Hungary.
-As to the western Europeans such as Denmark, Portugal and France we expect to see a sharp decline in the economic growth.
-Likewise, with the recent covid19 pandemic in terms of financial aspects.
+**Financial crisis, and EU membership for Slovakia and Hungary**
 
-Regarding the effects of the economic crisis, another paper notes that it led to a significant increase in regional unemployment rates across Europe.
-Reporting some regions experiencing unemployment rates as high as 30% [@iammarino2019].
-The authors also discuss how the crisis exposed the vulnerability of regions that were heavily reliant on a single industry or sector, and how sheltered economies were better able to weather the storm [@iammarino2019].
+The 2008 financial crisis significantly disrupted regional growth and exacerbated income inequalities across Europe.
+Nguyen et al.
+highlight that financial turmoil typically leads to increased income disparity, a trend observed globally [@nguyen2022].
+
+Focusing on Europe, we expect varied regional economic performances, especially when comparing Eastern European countries such as Hungary and Slovakia to their Western counterparts.
+Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential acceleration in growth for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014].
+However, the financial crisis and subsequent labor market shifts may have influenced these trends.
+
+EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic.
+These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.
+
+**Regional inequality**
+
 Another affect that surprised us was that the economic crisis led to increase inequality.
@@
 The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.
+
+**Determinants of regional inequality**
+
+In Lessmann and Seidel\'s exploration of regional inequality, the intricacies of this phenomenon unfold, revealing a multifaceted landscape.
+The study delineates how economic development carves an N-shaped trajectory in regional disparities, illustrating an initial surge in inequality, followed by a decrease, and an eventual uptick again, paralleling the region\'s economic progress.
+Geographic characteristics stand out as pivotal, with the natural and topographical elements of a region significantly swaying income distributions [@lessmann2017].
+
+Urbanization emerges as a crucial determinant, where bustling, populous urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
+The regulatory landscape too plays a formidable role; robust market regulations, particularly within the realms of capital and labour, are highlighted as vital catalysts fostering regional convergence and mitigating the chasms of inequality [@lessmann2017].
+Collectively, these determinants underscore the complexity of regional inequality, prompting a call for comprehensive and nuanced strategies to navigate and address its multifarious dimensions.
+
+*Transport infrastructure*
+
+Transport infrastructure plays a pivotal role in shaping regional economic landscapes.
+Improved connectivity fosters economic activities, reduces travel time, and enhances accessibility to markets and resources.
+Studies indicates that regions with robust transport infrastructure tend to attract more investments, generate employment, and ultimately contribute to regional prosperity.
+
+**Expected Inverted U- and N shaped curves**
+
 Furthermore, the paper provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
 The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
+
 @lessmann2017 paper also finds that there is an N-shaped relationship between regional inequality and economic development, which means that regional inequality increases again after the inverted U-shaped pattern has been completed [@lessmann2017].
 
-Lastly, @iammarino2019 discusses regional inequalities in European countries.
+@iammarino2019 discusses regional inequalities in European countries.
 It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium.
@@
 Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
-On the other hand, France has been a subject of discussion and policymaking for many years, regarding its pockets of extreme wealth.
+On the other hand, France has been a subject of discussion and policy making for many years, regarding its pockets of extreme wealth.
 (finner ikke kilde)
+
+Lastly, It is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure emerging as a significant determinant.
+Earlier studies provides us with a foundational understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities.
+The integration of transport infrastructure into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
 
@@
 
-### **Outlier discussion**
+### **Outline discussion**
 

### 2023-11-01 MSB104_GR_1_Final_Assignment_research_article.tex
mer text

@@
 \makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 This assignment aims to acquire, process, and analyze sub-national GDP
-and population data for a subset of European countries. Calculate GDP
-per capita and explore regional inequity using various descriptive
-statistics and visualizations.Furthermore examining the relationship
+and population data for a subset of European countries. Firstly
+(assignment 1), Calculateing GDP per capita and exploreing regional
+inequity using various descriptive statistics and
+visualizations.Secondly (assignment 2), examining the relationship
 between regional development and inequality, employing cross sectional
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 can, paradoxically, coexist with pockets of regional deprivation
-(\textbf{lessmann2017?}).
+(Lessmann and Seidel 2017).
 
@@
 concern, especially when crafting policies for equitable growth
-(\textbf{lessmann2017?}). A country's macro-level prosperity does not
+(Lessmann and Seidel 2017). A country's macro-level prosperity does not
 automatically guarantee that all its regions partake equally in this
@@
 get a more nuanced narrative about the state of regional economic
-disparities (\textbf{lessmann2017?}).
+disparities (Lessmann and Seidel 2017).
 
@@
 acquire, process, and analyze sub-national GDP and population data for a
-subset of European countries. Calculate GDP per capita and explore
+subset of European countries.
+
+The primary objective is to gain knowledge of econometric methods used
+for science in business through our given topic economic growth and
+inequality measures. Both GDP per capita and the Gini coefficient are
+vital indicators in the field of economics and are frequently used in a
+wide array of studies to analyse and interpret the economic conditions
+and social disparities of different regions or countries. Their analysis
+provides crucial insights for policymakers, economists, and researchers
+in shaping and evaluating the effectiveness of socio-economic policies.
+
+The research question \textbf{\emph{``What are the regional economic
+growth and inequality measurements from 2000 to 2020, in selected
+European countries and how does transport infrastructure affect the
+results?''}} is key in guiding our study. Setting a solid foundation for
+a comprehensive study that not only measure our given parameters, but
+also critically examines the role of transportation infrastructure play
+in these dyniamics. Lessmann and Seidel (2017) states that
+transportation costs play an important role in agglomeration and income,
+and reduced transportation costs as a determinant of regional inequality
+transportation costs have the potential to reduce regional inequalities.
+
+The sructure of our study is methodically designed to facilitate this
+exploration. In part 1A and B we calculate GDP per capita and explore
 regional inequity using various descriptive statistics and
-visualizations
+visualizations such as panel data (\textbf{\emph{Assignment\_1}}).
 
-\hypertarget{literature-review}{%
-\section{Literature review}\label{literature-review}}
+In part 2A and B we use cross sectional method to examining the
+relationship between regional development and inequality, employing
+cross sectional estimation techniques for the year 2010
+(\textbf{\emph{Assignment\_2}}).
 
-Regional growth and inequality have been important topics of research in
-Europe for many years. In recent decades, the European Union has
-implemented policies aimed at reducing regional disparities and
-promoting economic growth across its member states. Furthermore, despite
-these efforts, significant disparities in income and economic
-development persist across regions in Europe (\textbf{lessmann2017?}).
+\hypertarget{literature-review-regional-economic-growth-inequality-and-transport-infrastructure-2000-2020}{%
+\section{Literature Review: Regional Economic Growth, Inequality, and
+Transport Infrastructure
+(2000-2020)}\label{literature-review-regional-economic-growth-inequality-and-transport-infrastructure-2000-2020}}
 
-When it comes to economic growth, several studies refer to numerous
-influencing factors. (\textbf{gennaioli2014?}) and
-(\textbf{crescenzi2012?}) states that regional growth, much like
-national growth, is significantly influenced by geography and human
-capital. Geography impacts trade, resource availability, and
-susceptibility to natural events. Human capital, encompassing skills and
-knowledge, drives innovation and productivity. Furthermore,
-(\textbf{gennaioli2014?}) mentions that wealthier nations like Denmark
-in our case, tend to experience faster regional convergence, suggesting
-national prosperity aids regional growth (\textbf{gennaioli2014?}).
-Convergence implies that poorer regions grow faster than richer ones,
-gradually reducing disparities in wealth and income. Also referred to as
-beta convergence in (\textbf{lessmann2017?}) paper.
+Why some places grow and prosper is a fundamental question in social
+science research. This question motivated Adam Smith\textquotesingle s
+work, \emph{An Inquiry into the Nature and Causes of the Wealth of
+Nations,} (Smith
+\href{https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4\#ref-CR32}{1863})
+and has been a major influence of the \emph{new economic geography}
+(Feldman 2014)\emph{. Furthermore,} regional growth and inequality have
+been important topics of research in Europe for many years. In recent
+decades, the European Union has implemented policies aimed at reducing
+regional disparities and promoting economic growth across its member
+states. Furthermore, despite these efforts, significant disparities in
+income and economic development persist across regions in Europe
+(Lessmann and Seidel 2017).
 
-In the context of Europe, and our selected countries we might expect to
-see a wide variety of regional economic performances, especially within
-Hungary and Slovakia and between the western European countries. Capital
-market regulations also play a fundamental role, with effective
-frameworks promoting quicker regional convergence
-(\textbf{gennaioli2014?}). Taken this in to consideration we might
-expect a faster growth rate after the joining of the EU in 2003 and
-2004. That said the opening of the free labor market might also
-influence the migration of labor in Slovakia and Hungary. Not to mention
-the economic crisis hitting shortly after our eastern European countries
-joined the union.
+\textbf{Economic Growth and Regional Disparities}
 
-Whilst their recent EU membership might have posed certain challenges
-for Eastern European countries during the financial crisis, it also
-offered tools and resources that aided in their recovery in Slovakia and
-Hungary. As to the western Europeans such as Denmark, Portugal and
-France we expect to see a sharp decline in the economic growth.
-Likewise, with the recent covid19 pandemic in terms of financial
-aspects.
+Regional growth in Europe exhibits a complex pattern, influenced
+significantly by factors such as geography, human capital, and economic
+policies. Lessmann and Seidel\textquotesingle s research offers insights
+into these dynamics, highlighting the persistence of income disparities
+across regions despite overarching economic advancement (Lessmann and
+Seidel 2017). Wealthier nations like Denmark tend to experience quicker
+regional convergence, a phenomenon supported by Gennaioli et
+al.\textquotesingle s findings (Gennaioli et al. 2014). However, this is
+not uniformly experienced across all European countries, with Eastern
+European nations such as Hungary and Slovakia showcasing different
+growth trajectories, particularly after their EU accession.
 
-Regarding the effects of the economic crisis, another paper notes that
-it led to a significant increase in regional unemployment rates across
-Europe. Reporting some regions experiencing unemployment rates as high
-as 30\% (\textbf{iammarino2019?}). The authors also discuss how the
-crisis exposed the vulnerability of regions that were heavily reliant on
-a single industry or sector, and how sheltered economies were better
-able to weather the storm (\textbf{iammarino2019?}). Another affect that
-surprised us was that the economic crisis led to increase inequality.
-The study ``\emph{the effects of financial crisis on regional
-inequality}'' indicate any type of financial crisis results in higher
-income inequality (\textbf{nguyen2022?}).
+\textbf{Financial crisis, and EU membership for Slovakia and Hungary}
 
-(\textbf{lessmann2017?}) regional inequality study address the
+The 2008 financial crisis significantly disrupted regional growth and
+exacerbated income inequalities across Europe. Nguyen et al.~highlight
+that financial turmoil typically leads to increased income disparity, a
+trend observed globally (Nguyen 2022).
+
+Focusing on Europe, we expect varied regional economic performances,
+especially when comparing Eastern European countries such as Hungary and
+Slovakia to their Western counterparts. Given the role of capital market
+regulations in promoting regional convergence, outlined by Gennaioli et
+al., we anticipate a potential acceleration in growth for Eastern
+European countries following their EU accession in 2003 and 2004
+(Gennaioli et al. 2014). However, the financial crisis and subsequent
+labor market shifts may have influenced these trends.
+
+EU membership provided Slovakia and Hungary with resources for economic
+recovery, while Western European countries, including Denmark, Portugal,
+and France, likely experienced significant economic declines due to the
+financial crisis, a pattern possibly mirrored in the recent COVID-19
+pandemic. These complexities highlight the importance of a detailed
+analysis to understand regional economic growth and inequality in
+challenging times.
+
+\textbf{Regional inequality}
+
+Another affect that surprised us was that the economic crisis led to
+increase inequality. The study ``\emph{the effects of financial crisis
+on regional inequality}'' indicate any type of financial crisis results
+in higher income inequality (Nguyen 2022).
+
+Lessmann and Seidel (2017) regional inequality study address the
 importance of studying regional inequality. In particular, its potential
 consequences, such as political tensions that can undermine social and
-political stability (\textbf{lessmann2017?}). The paper also discusses
+political stability (Lessmann and Seidel 2017). The paper also discusses
 the relationship between regional inequality and personal income
-inequality and conflict risk. Furthermore, the paper provides us with
-insights into the inverted U-shaped relationship between regional
-inequality and the level of economic development in different country
-groups (\textbf{lessmann2017?}). The inverted U-shape also referred to
-as Kuznets Curve suggests that as an economy develops, regional
-inequality initially rises, reaches a peak, and then starts to decline.
-(\textbf{lessmann2017?}) paper also finds that there is an N-shaped
+inequality and conflict risk.
+
+\textbf{Determinants of regional inequality}
+
+In Lessmann and Seidel\textquotesingle s exploration of regional
+inequality, the intricacies of this phenomenon unfold, revealing a
+multifaceted landscape. The study delineates how economic development
+carves an N-shaped trajectory in regional disparities, illustrating an
+initial surge in inequality, followed by a decrease, and an eventual
+uptick again, paralleling the region\textquotesingle s economic
+progress. Geographic characteristics stand out as pivotal, with the
+natural and topographical elements of a region significantly swaying
+income distributions (Lessmann and Seidel 2017).
+
+Urbanization emerges as a crucial determinant, where bustling, populous
+urban centres usually enjoy elevated income levels, contrasting starkly
+with the often-stark disparities witnessed in more rural and secluded
+areas (Lessmann and Seidel 2017). The regulatory landscape too plays a
+formidable role; robust market regulations, particularly within the
+realms of capital and labour, are highlighted as vital catalysts
+fostering regional convergence and mitigating the chasms of inequality
+(Lessmann and Seidel 2017). Collectively, these determinants underscore
+the complexity of regional inequality, prompting a call for
+comprehensive and nuanced strategies to navigate and address its
+multifarious dimensions.
+
+\emph{Transport infrastructure}
+
+Transport infrastructure plays a pivotal role in shaping regional
+economic landscapes. Improved connectivity fosters economic activities,
+reduces travel time, and enhances accessibility to markets and
+resources. Studies indicates that regions with robust transport
+infrastructure tend to attract more investments, generate employment,
+and ultimately contribute to regional prosperity.
+
+\textbf{Expected Inverted U- and N shaped curves}
+
+Furthermore, the paper provides us with insights into the inverted
+U-shaped relationship between regional inequality and the level of
+economic development in different country groups (Lessmann and Seidel
+2017). The inverted U-shape also referred to as Kuznets Curve suggests
+that as an economy develops, regional inequality initially rises,
+reaches a peak, and then starts to decline.
+
+Lessmann and Seidel (2017) paper also finds that there is an N-shaped
 relationship between regional inequality and economic development, which
 means that regional inequality increases again after the inverted
-U-shaped pattern has been completed (\textbf{lessmann2017?}).
+U-shaped pattern has been completed (Lessmann and Seidel 2017).
 
-Lastly, (\textbf{iammarino2019?}) discusses regional inequalities in
-European countries. It presents evidence that inter-regional inequality
-has been increasing in the European Union since the turn of the
-millennium. The authors argue that this is due to the existence of
-several groups of regional economies in Europe that are structurally
-very different from one another (\textbf{iammarino2019?}). Additionally,
-the authors argue that there are some countries in the EU that are more
-evenly developed than others, and that a map of under performance or
-over-performance means less in a high-income but evenly developed
-country, such as in Denmark in our case (\textbf{iammarino2019?}). On
-the other hand, France has been a subject of discussion and policymaking
-for many years, regarding its pockets of extreme wealth. (finner ikke
-kilde)
+(\textbf{iammarino2019?}) discusses regional inequalities in European
+countries. It presents evidence that inter-regional inequality has been
+increasing in the European Union since the turn of the millennium. The
+authors argue that this is due to the existence of several groups of
+regional economies in Europe that are structurally very different from
+one another (\textbf{iammarino2019?}). Additionally, the authors argue
+that there are some countries in the EU that are more evenly developed
+than others, and that a map of under performance or over-performance
+means less in a high-income but evenly developed country, such as in
+Denmark in our case (\textbf{iammarino2019?}). On the other hand, France
+has been a subject of discussion and policy making for many years,
+regarding its pockets of extreme wealth. (finner ikke kilde)
+
+Lastly, It is evident that regional economic growth and inequality in
+Europe from 2000 to 2020 have been influenced by a constellation of
+factors, with transport infrastructure emerging as a significant
+determinant. Earlier studies provides us with a foundational
+understanding of these dynamics, while also highlighting the nuanced
+impacts of economic crises on regional disparities. The integration of
+transport infrastructure into this discourse adds an additional layer of
+complexity, underscoring its potential as a catalyst for regional
+development and equality.
 
@@
 descriptive statistics. Measuring regional income inequality is
-challenging due to heterogeneity of regions (\textbf{lessmann2017?}).
+challenging due to heterogeneity of regions (Lessmann and Seidel 2017).
 The number of regions in our data set varies largely in size and
@@
 becomes a minor issue because the country-level territorial
-heterogeneity is fixed (\textbf{lessmann2017?}).
+heterogeneity is fixed (Lessmann and Seidel 2017).
 
@@
 
-\hypertarget{outlier-discussion}{%
-\subsubsection{\texorpdfstring{\textbf{Outlier
-discussion}}{Outlier discussion}}\label{outlier-discussion}}
+\hypertarget{outline-discussion}{%
+\subsubsection{\texorpdfstring{\textbf{Outline
+discussion}}{Outline discussion}}\label{outline-discussion}}
 
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-feldman2014}{}}%
+Feldman, Maryann P. 2014. {``The Character of Innovative Places:
+Entrepreneurial Strategy, Economic Development, and Prosperity.''}
+\emph{Small Business Economics} 43 (1): 9--20.
+\url{https://doi.org/10.1007/s11187-014-9574-4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gennaioli2014}{}}%
+Gennaioli, Nicola, Rafael La Porta, Florencio Lopez De Silanes, and
+Andrei Shleifer. 2014. {``Growth in Regions.''} \emph{Journal of
+Economic Growth} 19 (3): 259--309.
+\url{https://doi.org/10.1007/s10887-014-9105-9}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lessmann2017}{}}%
+Lessmann, Christian, and André Seidel. 2017. {``Regional Inequality,
+Convergence, and Its Determinants -- a View from Outer Space.''}
+\emph{European Economic Review} 92 (February): 110--32.
+\url{https://doi.org/10.1016/j.euroecorev.2016.11.009}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
+Nguyen, Thanh Cong. 2022. {``The Effects of Financial Crisis on Income
+Inequality.''} \emph{Development Policy Review} 40 (6): e12600.
+\url{https://doi.org/10.1111/dpr.12600}.
+
+\end{CSLReferences}
 

### 2023-11-02 MSB104_GR_1_Final_Assignment_research_article.qmd
Add files via upload

@@
 
-Using time series data and cross sectional observations we investigate the GDP per capita and GINI trends in Portugal, France, Hungary, Slovenia and Denmark in the time period 2000 to 2020.
-The aims is to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
+Using time series data and cross-sectional observations we investigate Economic growth and Inequality trends in Portugal, France, Hungary, Slovenia and Denmark in the time period 2000 to 2020.
+The aims are to acquire, process, and analyse sub-national GDP, GINI and population data for a subset of European countries.
 
-The primary objective is to gain knowledge of econometric methods used for science in business through our given topic economic growth and inequality measures.
-Both GDP per capita and the Gini coefficient are vital indicators in the field of economics and are frequently used in a wide array of studies to analyse and interpret the economic conditions and social disparities of different regions or countries.
+The primary objective is to gain knowledge of econometric methods used for science in business through our given topic.
+Subsequently, econometric terms and definitions are explained in greater detail than it normally would in similar papers.
+Using GDP per capita and the Gini coefficient.
+Moreover, both are vital indicators in the field of economics and are frequently used, in a wide array of studies.
+Furthermore, to analyse and interpret the economic conditions and social disparities of different regions or countries.
 Their analysis provides crucial insights for policymakers, economists, and researchers in shaping and evaluating the effectiveness of socio-economic policies.
 
-The research question ***"What are the regional economic growth and inequality measurements from 2000 to 2020, in selected European countries and how does transport infrastructure affect the results?"*** is key in guiding our study.
-Setting a solid foundation for a comprehensive study that not only measure our given parameters, but also critically examines the role of transportation infrastructure play in these dyniamics.
-@lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs as a determinant of regional inequality transportation costs have the potential to reduce regional inequalities.
+The research question ***"What are the regional economic growth and inequality measurements from 2000 to 2020, in selected European countries and how does other*** ***determinants affect the results?"*** is key in guiding our study.
+Setting a solid foundation for a comprehensive study that not only measure our given parameters, but also a critically examination of the role of determinants such as transportation infrastructure and education.
+@lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs as a determinant of regional inequality.
 
 The sructure of our study is methodically designed to facilitate this exploration.
-In part 1A and B we calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations such as panel data (***Assignment_1***).
+In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***).
 
-In part 2A and B we use cross sectional method to examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).
+In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).
 
-# Literature Review: Regional Economic Growth, Inequality, and Transport Infrastructure (2000-2020)
+Lastly, potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions.
+The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017).
+Despite these precautions, there may have been instances where incorrect information was provided, and some errors could have been overlooked.
+The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.
+
+# Literature Review: Regional Economic Growth, Inequality and determines (2000-2020)
 
 Why some places grow and prosper is a fundamental question in social science research.
-This question motivated Adam Smith\'s work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A.
+This question motivated Adam Smith's work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A.
 (1863).
@@
 Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
-Lessmann and Seidel\'s research offers insights into these dynamics, highlighting the persistence of income disparities across regions despite overarching economic advancement [@lessmann2017].
-Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.\'s findings [@gennaioli2014].
-However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia showcasing different growth trajectories, particularly after their EU accession.
+Lessmann and Seidel's research offers insights into these dynamics, highlighting the persistence of income disparities, across regions despite overarching economic advancement [@lessmann2017].
+Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014].
+However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia we may expect different growth trajectories, particularly after their EU accession.
 
@@
 Nguyen et al.
-highlight that financial turmoil typically leads to increased income disparity, a trend observed globally [@nguyen2022].
+highlight that financial turmoil typically leads to increased income disparity; a trend observed globally [@nguyen2022].
 
 Focusing on Europe, we expect varied regional economic performances, especially when comparing Eastern European countries such as Hungary and Slovakia to their Western counterparts.
-Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential acceleration in growth for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014].
-However, the financial crisis and subsequent labor market shifts may have influenced these trends.
+Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential growth acceleration.
+In particular for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014].
+However, the financial crisis and subsequent labour market shifts may also have influenced these trends.
 
@@
 @lessmann2017 regional inequality study address the importance of studying regional inequality.
-In particular, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
+Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
 The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.
@@
 
-In Lessmann and Seidel\'s exploration of regional inequality, the intricacies of this phenomenon unfold, revealing a multifaceted landscape.
-The study delineates how economic development carves an N-shaped trajectory in regional disparities, illustrating an initial surge in inequality, followed by a decrease, and an eventual uptick again, paralleling the region\'s economic progress.
 Geographic characteristics stand out as pivotal, with the natural and topographical elements of a region significantly swaying income distributions [@lessmann2017].
 
-Urbanization emerges as a crucial determinant, where bustling, populous urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
-The regulatory landscape too plays a formidable role; robust market regulations, particularly within the realms of capital and labour, are highlighted as vital catalysts fostering regional convergence and mitigating the chasms of inequality [@lessmann2017].
-Collectively, these determinants underscore the complexity of regional inequality, prompting a call for comprehensive and nuanced strategies to navigate and address its multifarious dimensions.
+Urbanization emerges as a crucial determinant, where bustling, populous around urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
 
-*Transport infrastructure*
+*Transport infrastructure and Education*
 
-Transport infrastructure plays a pivotal role in shaping regional economic landscapes.
+Transport infrastructure plays a central role in shaping regional economic landscapes.
 Improved connectivity fosters economic activities, reduces travel time, and enhances accessibility to markets and resources.
@@
 
+[@lessmann2017] discusses the determinant of Education in the context of human capital.
+The study mentions that human capital is the most important determinant of differences in regional development within countries.
+The study also notes that human capital creates positive externalities, which can spill over regional boundaries.
+Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
+Lastly, promoting internal migration.
+
 **Expected Inverted U- and N shaped curves**
 
-Furthermore, the paper provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
+Furthermore, [@lessmann2017] provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
 The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
@@
 The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019].
-Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
-On the other hand, France has been a subject of discussion and policy making for many years, regarding its pockets of extreme wealth.
-(finner ikke kilde)
+Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
+On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.
 
-Lastly, It is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure emerging as a significant determinant.
-Earlier studies provides us with a foundational understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities.
-The integration of transport infrastructure into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
+Overall, it is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure and education emerging as significant determinants.
+Earlier studies provide us with a fundamental understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities.
+The integration of transport infrastructure and education into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
 
@@
 
-Trough Eurostat, we download the datasets nama_10r_3gdp and demo_r_pjanggr3 as csv files and filted the data by our preferences before downloading it.
+Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files and filted the data by our preferences before downloading it.
 Furthermore, we filtered the dataset by choosing the years 2000 to 2020.
-Then, we selected the NUTS 3 region for Portugal, France, Hungary, Slovakia and Denmark.
+Then, we selected the NUTS requiered levels region for Portugal, France, Hungary, Slovakia and Denmark.
 Finally, we specified the data to be in million Euro.
 
-### GDP
+## GDP
 
@@
 It furnishes GDP values in both current prices and adjusted for inflation, with figures given in purchasing power standards (PPS) to account for price level differences between countries.
-The data is mostly structured by year and region @eurostat2023.
+The data is mostly structured by year and region @database.
 
@@
 It is calculated as the sum of the gross value added across various institutional sectors or industries.
-Furthermore, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries) @eurostat2023c.
+Furthermore, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries) @database.
 This also balances out in the total economy production account.
@@
 
-The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @eurostat2023c.
+The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @database.
 
-### Population
+## Population
 
@@
 Our dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender.
-Displayed in a year-by-region format, with yearly interval updates @eurostat2023a.
+Displayed in a year-by-region format, with yearly interval updates @database.
 
 Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project.
-The project covers 37 countries and is the central repository for demographic and migration-related data @eurostat2023a.
-As well as, specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year @eurostat2023a.
+The project covers 37 countries and is the central repository for demographic and migration-related data @database.
+As well as, specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year @database.
 Additionally, data on marriages, divorces, and migration flows are recorded .
 
-For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @eurostat2021.
+For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @database.
 
-### NUTS classification
+## Demographic (Nuts 2)
+
+Statistics on population change (demo_r_mlifexp) and on population structure was collected from Euro Stat [@database].
+Used to monitor demographic behavior within political, economic, social and cultural contexts at NUTS 2 level.
+Same policy and demographic classifications as previous population data set applies.
+
+## Transportation Infrastructure (Nuts 2)
+
+This data-set (tran_r_net) presents recent data on the inland transport network of the European Union (EU), EFTA and candidate countries presenting motorways, railways and inland waterways [@database].
+Furthermore used as a varable of of interest (MLR) in our section growth and inequality.
+Moreover, to excamine if evolution of the transport network is closely linked to the general development of the economy.
+
+The collected data are compiled by the competent national authorities.
+Reported annually be the National Statistical Office, the Civil Aviation Authorities (air transport), the Ministry of Transport (inland waterways, railway and road networks, victims), and the National Maritime Administration (maritime transport) [@database].
+
+## Education (Nuts 2)
+
+This data set (edat_lfse_04) contains population by educational attainment level, sex and NUTS 2 regions (%).
+Including data on the highest level of education successfully completed by the individuals of a given population.
+Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labour status of young people by years since completion of highest level of education.
+Calculated as annual averages of quarterly EU Labour Force Survey data (EU-LFS).
+
+## NUTS classification
 
@@
 
-### Descriptive statistics
+## Descriptive statistics
 
@@
 
-# Part 1A: Sub-national GDP and GDP per Capita
+## Part 1A: Sub-national GDP and GDP per Capita
 
-## GDP per Capita Calculation
+### GDP per Capita Calculation
 
@@
 
-### Portugal
+### GDP per capita Portugal
 
@@
 There is however some regional variability.
-We can see that the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas.
+We can see signs of urbanisazation where the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas.
 Since Lisbon is the capital of Portugal, there is probably a higher concentration of industries, making it a economic center (which again makes the GDP per capita higher).
@@
 
-### France
+### GDP per capita France
 
@@
 
-### Hungary
+### GDP per capita Hungary
 
@@
 
-### Slovakia
+### GDP per capita Slovakia
 
@@
 
-### Denmark
+### GDP per capita Denmark
 
@@
 
-# Part 1B: Regional Inequity
+## Part 1B: Regional Inequity
 
-## Gini Coefficient Calculation
+### Gini Coefficient Calculation
+
+In this part we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.
+
+The gini coefficient can help us measure inequality in a distribution, as is therefore a useful tool for us to use when we look at regional inequity.
+The closer the gini coefficient is to 1, the bigger the inequality is; a number closer to 0 equals equality.
+When looking at the gini coefficient for NUTS 2 regions, we also get a better overview over differences in income between different regions, and it also makes it easier to find the reasons as to why there is a difference between the regions @hasell2023.
 
@@
 $GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|$
-
-we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.
-
-The gini coefficient can help us measure inequality in a distribution, as is therefore a useful tool for us to use when we look at regional inequity.
-The closer the gini coefficient is to 1, the bigger the inequality is; a number closer to 0 equals equality.
-When looking at the gini coefficient for NUTS 2 regions, we also get a better overview over differences in income between different regions, and it also makes it easier to find the reasons as to why there is a difference between the regions @hasell2023.
 
@@
 
-# Part 2A: Growth and Inequity
+## Part 2A: Growth and Inequity
 
@@
 
+In this part of the paper, we will carry out a simple regression model and explore the effect of regional economic development, represented by 𝑦𝑗 (GDP per capita), on regional inequality, represented by 𝐺𝐼𝑁 𝐼𝑊𝑗.
+We will do this to gain an understanding of the connection between economic growth indicators such as GDP per capita and inequality might have.
+We will explore if higher GDP per capita may lead to less or greater inequality, and gain an understanding to what extent these variables are related.
+
+Unlike the traditional Gini coefficient, which treats all individuals equally regardless of the population size of the region they reside in, the weighted Gini considers the population size of each region, assigning more weight to regions with larger populations.
+
+In the context of regional inequality, this is particularly important because it ensures that the income disparities in more populous regions have a proportionally larger impact on the overall measure of inequality.
+For instance, if a country has one region with a very high level of income per capita but a small population, and another region with a lower level of income per capita but a large population, the weighted Gini coefficient would reflect the inequality experienced by a larger portion of the country's population, providing a more accurate picture of the national income distribution.
+
 We use simple linear regression to model the relationship with the GINI as the dependent variable and the natural logarithm of GDP per capita as the independent variable.
@@
 
-**###Vis formel##**
+$$
+Y_i = β_1 + β_2X_i + ε_i
+$$
 
-Y is the dependent variable (hva som perdikeres).
+$Y_i$ represent the dependent/ explained variable
 
-X is the independent variable (input).
+$X_i$ represent the independent/ explanatory variable
 
-β0​ is the intercept.
+$β_1$ represent the intercept/ constant
 
-β1​ is the slope of the line.
+$β_2$ represent the slope coefficient
 
-ε represents the residuals or error in the prediction.
+$ε_i$ represents the residuals or error in the prediction.
 
--   **Intercept (**β0​): represents the value of Y when X is 0.
+**Intercept (**β0​): represents the value of Y when X is 0.
 
@@
 
-"In linear regression, the objective is straightforward: determine the line that best represents the observed data.
-Furthermore, this 'goodness of fit' line is characterized by coefficients β0​ and β1, which represent the intercept and slope, respectively [@wooldridge2020].
-The optimal line is one that minimizes the sum of the squared differences between the observed values and the predictions made by the model [@wooldridge2020].
-This approach, known as the Ordinary Least Squares (OLS), has a solid mathematical foundation [@wooldridge2020].
-By taking the sum of these squared differences and setting their partial derivatives with respect to β0​ and β1 to zero, we can derive the formula that give us these coefficients.
-And once these coefficients are in hand, they provide a direct equation for the line that best fits our data set [@wooldridge2020].
-One of the strengths of OLS is its closed-form solution, meaning we can directly compute the coefficients from the data without iterative procedures [@wooldridge2020].
-Moreover, under the foundamental assumptions of linear regression, OLS stands out as the Best Linear Unbiased Estimator (BLUE), emphasizing its reliability and efficiency in estimating the true underlying relationship [@wooldridge2020]."
+#### **Goodness of fit**
 
-### Individual country cases
+The goodness of fit in a simple linear regression model measures how well the regression line approximates the real data points.
+The regression line that best represents the data according to the least squares criterion, which minimizes the sum of the squared vertical distances of the points from the line.
 
-#### Portugal
+R\^2 is a statistical measure that represents the proportion of the variance for the dependent variable that's explained by the independent variable.
+It ranges from 0 to 1.
+A higher R\^2 value indicates a better fit of the model to the data.
 
-#### Denmark
+R\^2 = 0 The model does not explain any of the variability of the response data around its mean.
 
-#### France
+R\^2 = 1 The model explains all the variability of the response data around its mean.
 
-#### Hungary
+If the assumptions for a simple linear regression are met, it indicates a good fit for the model.
+[Plotting the data can provide a visual indication of the goodness of fit.]{.underline} The points should fall around a straight line without clear patterns in the residuals.
+As it is crucial to verify these assumptions before proceeding with interpreting the results of the regression analysis.
 
-#### Slovakia
+**Residuals vs.
+Fitted Values Plot** to check for homoscedasticity and linearity.
+Ideally, this plot shows no pattern; the residuals are randomly scattered around the horizontal line at zero.
+If there's a pattern (like a curve or systematic spread of residuals), it suggests non-linearity or heteroscedasticity.
 
-### **Outline discussion**
+**Normal Q-Q Plot** to check if the residuals are approximately normally distributed.
+The points should fall roughly along a straight line.
+Deviations from a straight line suggest deviations from normality.
 
-**##Eksepel....When looking at the data it is obvious that we have too many regions with zero inequality.
-This seems strange.
-The reason for this is .....
-In the subsequent analysis it therefore would make scene to exclude these observations to not bias our finding.**
+**Scale-Location Plot** is another way to check for homoscedasticity.
+The spread of the residuals should be roughly the same across all levels of the fitted values.
+If the spread increases or decreases with the fitted values, it suggests heteroscedasticity.
 
-**Looking at specific countries we observe allso further abnormalities like in ....
-where ...
-As thees outliers are not do to a bias in our data we shold keep them in our data.
-However we need to test later the sensitivity of our findings regarding the presences of these borderline observations.
-...####**
+**Residuals vs.
+Leverage Plot** to identify influential observations that might unduly influence the regression.
+Points with high leverage or large residuals (outliers) can be identified.
+Cook's distance lines may be added to help identify points that have a large influence on the model.
 
-# Part 2B: Exploring Other Determinants of Inequity
+## Part 2B: Exploring Other Determinants of Inequity
 
@@
 
+Se data section for more specifications on the added population nut2, transport infrastructure and education.
+
 ### II. Multiple Linear Regression Model
+
+Multiple Linear Regression (MLR) extends simple linear regression to incorporate multiple explanatory variables, allowing us to examine how multiple factors impact a dependent variable.
+Choosing a data set from the year 2010 that consists of various regions (FRANCE?), with data on each region's economic indicators, demographic variables, and other factors.
+Our aim is to understand how these variables collectively affect regional inequality.
+
+### **Assumptions**
+
+**I. Linearity**
+
+The relationship between the independent variable X and the dependent variable Y is linear.
+This means that changes in X are associated with proportional changes in Y.
+A straight line should provide a good fit to the data points when plotted on a graph.
+
+You can create a scatter plot of Y versus X and visually inspect whether a straight line could well represent the relationship.
+Alternatively, you can plot the residuals versus the fitted values and check for any obvious patterns or non-linearity.
+
+**II. Independence:**
+
+Observations are independent of each other. The value of Y for one observation should not depend on the value of Y for any other observation.
+
+This assumption is more about study design and data collection.
+Ensure that your observations are not correlated in time or space.
+For example, if you're analyzing economic data across regions, make sure that the regions are not influencing each other.
+
+**III. Homoscedasticity (Equal Variance):**
+
+The variance of the residuals (the errors) is constant across all levels of X.
+This means that the spread or "width" of the residuals should remain roughly the same across all values of the independent variable.
+
+A plot of residuals versus fitted values should show a random scatter and not display any funnel-like shapes (wider at one end).
+
+**IV. Normality of Residuals:**
+
+The residuals (the differences between observed and predicted values) are normally distributed. This assumption is particularly important for hypothesis testing and creating confidence intervals.
+
+A Quantile-Quantile (Q-Q) plot of the residuals can show if they follow a normal distribution. The points should fall roughly along a straight line.
+
+**V. No Perfect Multicollinearity (Specific to Multiple Regression)**
+
+In multiple regression settings, this assumes that no independent variable is a perfect linear function of any other independent variables.
+While this is more pertinent to multiple linear regression, it is crucial there because high correlation between independent variables can lead to unstable coefficient estimates.
+
+Checking the variance inflation factor (VIF) for each variable; a VIF above 5-10 indicates a problematic amount of collinearity.
+
+### Model specification
+
+**Understanding the Coefficients**
+
+Intercept beta_0 Represents the expected value of the dependent variable when all independent variables are set to zero.
+Interpretation is often nonsensical in multiple regression if there is no meaningful condition where all predictors are zero.
+
+Slope Coefficients beta_1, beta_2, \..., \\beta_k \\))\*\*: Represent the expected change in the dependent variable for a one-unit change in the respective independent variable, holding all other variables constant.
+
+### Ordinary Least Squares (OLS) Estimation
+
+"The OLS method is employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
+This technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
+The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the data-set, without necessitating iterative procedures.
+
+Furthermore, when the classical linear regression assumptions are met, OLS guarantees that the estimators are BLUE, ensuring their unbiasedness and efficiency.
+This is particularly crucial in econometric analysis, where the precision and reliability of parameter estimates are paramount for policy implications and economic interpretations.
+The foundation assumptions of linear regression, including linearity, independence, homoscedasticity, and normality of residuals, are prerequisites for OLS to attain these desirable properties.
+It is imperative, therefore, to conduct diagnostic tests and assess the validity of these assumptions to ensure the robustness of the OLS estimators used in our analysis."
+
+#### **Model Validation with Cross-Sectional Data**
+
+~~- \*\*Residual Analysis\*\*: Examine residuals for patterns to check for violations of linearity, independence, and homoscedasticity.~~
+
+~~- \*\*Outlier and Influence Diagnostics\*\*: Identify regions with high leverage or large residuals to ensure they are not disproportionately affecting the model.~~
+
+~~- \*\*Multicollinearity\*\*: Given that MLR includes multiple predictors, check for multicollinearity to ensure that the predictors are not too highly correlated with each other.~~
+
+~~\#### \*\*Predictive Power and Goodness of Fit\*\*~~
+
+~~- \*\*\\( R\^2 \\) and Adjusted \\( R\^2 \\)\*\*: Provide information on how well the independent variables explain the variation in the dependent variable.~~
+
+~~- \*\*F-Test\*\*: Tests whether at least one of the independent variables is statistically significant in explaining variation in the dependent variable.~~
+
+~~\#### \*\*Conclusion\*\*~~
+
+~~- The multiple linear regression model, estimated using OLS, allows for the comprehensive analysis of how various factors impact regional inequality.~~
+
+~~- Validating the assumptions of the OLS and properly interpreting the coefficients are vital steps in ensuring the reliability of the model.~~
+
+~~- The use of cross-sectional data from the year 2010 ensures that the analysis is grounded in a specific time context, making the results more interpretable and applicable.~~
+
+#### MLR Portugal
+
+#### MLR Denmark
+
+#### MLR France
+
+#### MLR Hungary
+
+#### MLR Slovakia
 

### 2023-11-02 assignment_1.qmd

@@
 library(dineq)
+library(vtable)
+library(dineq)
+library(dplyr)
+library(huxtable)
+library(lmtest)
+library(plm)
+library(readr) 
 ```
@@
 
+## **I. Data Collection**
+
+```{r}
+#using R and the dplyr package to manipulate a dataset and calculate GDP per capita at the NUTS2 region level for the year 2010
+
+#GDP_Per_Capita_NUTS2 <- GDP_Per_Capita %>%
+  #filter(Year == 2010) %>%
+  #select(Year, Population, GDP, NUTS2, GINI_NUTS2, NUTS) %>%
+  #group_by(NUTS2, Year) %>%
+  #mutate(GDP_NUTS2 = sum(GDP), Population_NUTS2 = sum(Population)) %>%
+  #ungroup() %>%
+  #distinct(NUTS, NUTS2, Year, GDP_NUTS2, Population_NUTS2, GINI_NUTS2) %>%
+  #mutate(GDP_capita_NUTS2 = (GDP_NUTS2 * 1000000)/Population_NUTS2)
+
+```
+
+```{r}
+
+#using R and the dplyr package to load and manipulate three different datasets related to demographics, transport infrastructure, and education at the NUTS2 regional level.
+Below is a breakdown of the steps and transformations you are performing on each dataset:
+
+#Demographics_NUTS2 <- read.csv('demo_r_mlifexp__custom_3767848_linear.csv') %>%
+  #select(geo, TIME_PERIOD, OBS_VALUE) %>%
+  #mutate(Life_expectency = OBS_VALUE, Year = TIME_PERIOD, NUTS2 = geo) %>%
+  #select(NUTS2, Year, Life_expectency)
+
+#Trans_infrastructure_NUTS2 <- read.csv('tran_r_net__custom_3761027_linear.csv') %>%
+  #select(geo, TIME_PERIOD, OBS_VALUE, tra_infr) %>%
+  #mutate(Number_of_MWAY_and_RL = OBS_VALUE, Year = TIME_PERIOD, NUTS2 = geo, #MWAY_or_RL = tra_infr) %>%
+  #select(NUTS2, Year, MWAY_or_RL, Number_of_MWAY_and_RL)
+
+#Education_NUTS2 <- read.csv('edat_lfse_04__custom_3767870_linear.csv') %>%
+  #filter(age == "Y25-64") %>%
+  #select(geo, TIME_PERIOD, OBS_VALUE) %>%
+  #mutate(Tertiary_educ = OBS_VALUE, Year = TIME_PERIOD, NUTS2 = geo) %>%
+  #select(NUTS2, Year, Tertiary_educ)
+```
+
+```{r}
+#Integrating different datasets into a comprehensive dataframe named Inequity_Data, and subsequently creating separate dataframes for specific countries based on their NUTS (Nomenclature of Units for Territorial Statistics) codes and the year 2010.
+Here's a breakdown of the code:
+
+Inequity_Data <- GDP_Per_Capita_NUTS2 %>%
+  left_join(Education_NUTS2, by = c("Year", "NUTS2")) %>%
+  left_join(Demographics_NUTS2, by = c("Year", "NUTS2")) %>%
+  left_join(Trans_infrastructure_NUTS2, by = c("Year", "NUTS2")) %>%
+  mutate(log_GDP_capita_NUTS2 = log(GDP_capita_NUTS2)) %>%
+  filter(Year == 2010 & GINI_NUTS2 > 0)
+
+AT2010 <- Inequity_Data %>%
+  filter(NUTS == "PT")
+BE2010 <- Inequity_Data %>%
+  filter(NUTS == "FR")
+BG2010 <- Inequity_Data %>%
+  filter(NUTS == "HU")
+HR2010 <- Inequity_Data %>%
+  filter(NUTS == "SK")
+IT2010 <- Inequity_Data %>%
+  filter(NUTS == "DK")
+```
+
+## **II Model specification**
+
+**Simple Linear Regression**
+
+```{r}
+#creating linear regression models for seven different datasets, each corresponding to a different country or the entire dataset, to analyze the relationship between regional inequality (measured by the GINI_NUTS2 variable) and per capita GDP (measured by the GDP_capita_NUTS2 variable) for the year 2010.
+
+#lm1 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = PT2010)
+#lm2 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = FR2010)
+#lm3 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = HU2010)
+#lm4 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = SK2010)
+#lm5 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = DK2010)
+#lm6 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = Inequity_Data)
+```
+
+Table XXX
+
+```{r}
+#using the huxreg() function to create a table summarizing the results of several linear regression models.
+The huxreg() function is typically used with the huxtable package in R, which is a tool for creating LaTeX and HTML tables from R objects.
+
+#huxreg(list("PT" = lm1, "FR" = lm2,  "HU" = lm3, "SK" = lm4, "DK" = lm5, "CC" = lm6), statistics = c(N = "nobs", R2 = "r.squared"), coefs = c(GDP_capita_NUTS2 = "GDP_capita_NUTS2", const.
+= "(Intercept)"), note = "Note: {stars} T statistics in brackets.")
+```
+
+PLOTTING
+
+```{r}
+#creating a scatter plot of the GINI_NUTS2 against GDP_capita_NUTS2 using data from Inequity_Data.
+In addition to the scatter plot, a regression line (calculated from a previous simple linear regression model, presumably) is added to the plot.
+
+#| label: fig-17
+#| fig-cap: "Gini & GDP per capita CC"
+#Inequity_Data %>%
+  #ggplot(aes(x = GDP_capita_NUTS2, y = GINI_NUTS2)) +
+  #geom_point(color = "maroon", alpha = 0.7) +
+  #geom_abline(intercept = b0_GDP_on_GINI_IT, slope = b1_GDP_on_GINI_IT, color = #"blue", alpha = 0.5) +
+  #labs(x = "GDP per Capita", y = "GINI") +
+  #theme_bw()
+
+```
+
+## **III. Residuales / model interpetation?**
+
+```{r}
+#predicting the GINI coefficient based on GDP per capita for all the regions in Inequity_Data using the lm_GDP_on_GINI model, and then adding these predicted values and residuals to the Inequity_Data dataframe.
+
+# Fit a simple linear regression model
+#lm_GDP_on_GINI <- lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = Inequity_Data)
+
+# Predict the GINI coefficient and add predictions and residuals to the dataframe
+#Inequity_Data$Predicted <- predict(lm_GDP_on_GINI)
+#Inequity_Data$Residuals <- residuals(lm_GDP_on_GINI)
+
+#Inequity_Data$Predicted <- predict(lm_GDP_on_GINI)
+#Inequity_Data$Residuals <- residuals(lm_GDP_on_GINI)
+```
+
+```{r}
+#| label: fig-18
+#| fig-cap: "Gini & GDP per capita CC with residuals"
+
+#predicting the GINI coefficient based on GDP per capita for all the regions in Inequity_Data using the lm_GDP_on_GINI model, and then adding these predicted values and residuals to the Inequity_Data dataframe.
+
+#Inequity_Data %>%
+  #distinct(NUTS2, GDP_capita_NUTS2, .keep_all = TRUE) %>%
+  #ggplot(aes(x = GDP_capita_NUTS2, y = GINI_NUTS2)) +
+  #geom_point(color = "maroon", alpha = 0.7) +
+  #geom_segment(aes(x = GDP_capita_NUTS2, xend = GDP_capita_NUTS2, y = GINI_NUTS2, #yend = Predicted), color = "red", alpha = 0.4) +
+                 #geom_abline(intercept = b0_GDP_on_GINI, slope = b1_GDP_on_GINI, color = "blue", alpha = 0.5) +
+                 #labs(x = "GDP per Capita", y = "GINI") + theme_bw()
+```
+
+## Other determinants of inequity
+
+```{r}
+
+#defines a function named MRM_NUTS2 that performs multiple linear regression analyses on a given dataset (mlrdata).
+This function creates five linear models, each with different sets of predictors, and then displays the summary of these models in a nicely formatted table using the huxreg function from the huxtable package.
+
+#MRM_NUTS2 <- function(mlrdata){
+  #lm8 = lm(GINI_NUTS2 ~ Tertiary_educ, mlrdata)
+  #lm9 = lm(GINI_NUTS2 ~ Number_of_MWAY_and_RL, mlrdata)
+  #lm10 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2 + Tertiary_educ + Number_of_MWAY_and_RL, #mlrdata)
+  #lm11 = lm(GINI_NUTS2 ~ Life_expectency, mlrdata)
+  #lm12 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2 + Tertiary_educ + Life_expectency, #mlrdata)
+
+  #huxreg(
+    #list("Model 1" = lm8, "Model 2" = lm9, "Model 3" = lm10, "Model 4" = lm11, "Model 5" = lm12), statistics = c(N = "nobs", R2 = "r.squared"), coefs = c(Higher.educ = "Tertiary_educ","Number_of_MWAY_and_RL", Life.expec = "Life_expectency", const.
+= "(Intercept)"),note = "Note: {stars} T statistics in brackets.")
+
+```
+
+### MLR by country
+
+#### Portugal
+
+```{r}
+#| label: MRL Portugal
+#| tbl-cap: "MRM PT"
+#MRM_NUTS2(AT2010)
+```
+
+#### France
+
+```{r}
+#| label: MLR France
+#| tbl-cap: "MRM FR"
+MRM_NUTS2(AT2010)
+```
+
+#### Hungary
+
+```{r}
+#| label: MRL Hungary
+#| tbl-cap: "MRM AT"
+MRM_NUTS2(AT2010)
+```
+
+#### Slovakia
+
+```{r}
+#| label: MRL Slovakia
+#| tbl-cap: "MRM AT"
+MRM_NUTS2(AT2010)
+```
+
+#### Denmark
+
+```{r}
+#| label: MRL Denmark
+#| tbl-cap: "MRM AT"
+MRM_NUTS2(AT2010)
+```
+
 ### 1. Model Estimation

### 2023-11-02 MSB104_GR_1_Final_Assignment_research_article.qmd

@@
 editor: visual
-bibliography: [Regional_inequality_23.bib]
+bibliography: [Regional_inequality_23.bib, econometrics.bib]
 output-file: Assignment1.qmd

### 2023-11-09 MSB104_GR_1_Final_Assignment_research_article.qmd
update 09.11

--- 2023-11-02 MSB104_GR_1_Final_Assignment_research_article.qmd
@@
 editor: visual
-bibliography: [Regional_inequality_23.bib]
-output-file: Assignment1.qmd
+bibliography: [Regional_inequality_23.bib, econometrics.bib]
 page-layout: full
@@
 papersize: a4
+csl: apa.csl
 ---

### 2023-11-09 assignment_1.qmd
update 09.11

--- 2023-10-22 assignment_1.qmd
@@
 library(dineq)
+library(huxtable)
 ```
+
+Må fikse kilder!!
 
@@
 
+... p. 116
+
 ## Discussion
 
-# References
+# Assignment 2: Cross-sectional Estimates
+
+## 2A: Growth and Inequity
+
+### 1. Model Estimation
+
+$\text{GDP per capita = } \beta_{0} + \beta_{1} \cdot GINI + \in$
+
+```{r}
+#| echo: false
+#| output: false
+
+# Making a dataframe that only contains observations from 2010, also filter the data so that gini is bigger than 0.
+
+data2010 <- gini_nuts_2 %>% filter(year == 2010, gini_n2 > 0)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: Descriptive statistics all countries for year 2010
+
+# Finding the natural logarithm of GDP per capita
+data2010 <- data2010 |> 
+  mutate(log_GDP_per_capita = log(GDP_per_capita))
+
+# lm(y ~ x) y = GDP per capita, x = Gini, estimation
+lm_centered <- lm(gini_n2 ~ log_GDP_per_capita, data = data2010)
+
+#Save coefficients
+b0_earn <- lm_centered$coefficients[1]
+b1_earn <- lm_centered$coefficients[2]
+
+# Descriptive statistics
+huxreg(All = lm_centered, statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+```{r}
+#| echo: false
+#| output: false
+# må forklare denne!!
+plot(lm_centered, which = 1:4)
+```
+
+β0​ = 0.0551 (intercept)
+
+β1 = 0.0193 (slope coefficient)
+
+```{r}
+#| echo: false
+#| fig-cap: Descriptive statistics for all countries seperately for the year 2010
+
+# Finding estimates for each country seperately 
+lmSK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "SK"))
+lmDK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "DK"))
+lmHU <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "HU"))
+lmPT <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "PT"))
+lmFR <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "FR"))
+
+# Descriptive statistics
+huxreg(SK = lmSK, DK = lmDK, HU = lmHU, PT = lmPT, FR = lmFR, statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+### 2. Model Diagnostics
+
+Forklaring på statistikk for alle land samlet:
+
+Forklaring på statistikk for land hver for seg:
+
+### 3. Visualization
+
+```{r}
+#| echo: false
+
+# Visualization of the relationship between GDP per capita and gini
+ggplot(data2010, aes(x = gini_n2, y = log_GDP_per_capita)) +
+  geom_point(aes(color = gini_n2), size = 3) +
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  labs(
+    title = "Relationship Between GDP per Capita and Gini",
+    x = "Gini",
+    y = "GDP per capita",
+    caption = "Red line indicates the fitted regression line"
+  ) +
+  theme_minimal()
+
+```
+
+Forklaring på visualisering:
+
+Forklar hvorfor noen observasjoner er langt unna de andre observasjonene.
+
+## 2B: Exploring Other Determinants of Inequity
+
+### 1. Data Acquistion
+
+```{r}
+#| echo: false
+#| output: false
+#| label: data acquistion
+
+# Data for population density
+demo_dens_data <- read_csv("Raw_data/demo_dens.gz")
+
+# Data for education
+educ_data <- read_csv("Raw_data/educ.gz")
+
+# Data for rail network
+rail_data <- read_csv("Raw_data/rail.gz")
+```
+
+```{r}
+#| echo: false
+#| output: false
+
+# Cleaning up data frames
+demo_dens_data <- demo_dens_data |> 
+  select(nuts_2 = geo, pop_density = OBS_VALUE)
+
+educ_data <- educ_data |> 
+  select(nuts_2 = geo, students_percentage = OBS_VALUE)
+
+rail_data <- rail_data |> 
+  select(nuts_2 = geo, rail_km = OBS_VALUE)
+```
+
+```{r}
+merged_data_1 <- demo_dens_data %>%
+  inner_join(educ_data, by = 'nuts_2') %>%
+  inner_join(rail_data, by = 'nuts_2')
+```
+
+#### Population Density
+
+#### Education
+
+#### Rail Network
+
+### 2. Multiple Linear Regression Model
+
+### 3. Model Interpretation

### 2023-11-10 assignment_1.qmd
update 10.11

--- 2023-11-01 assignment_1.qmd
@@
 library(dineq)
+library(huxtable)
 ```
@@
 
-...
+... p. 116
 
 ## Discussion
-
-What does the GINI really tell us? (Look at paper from page 116)
 
@@
 
+$GINI = \beta_{0} + \beta_{1} \cdot \text{GDP per capita} + \epsilon$
+
+Gini is the dependent variable (y), while GDP per capita is the independent variable (x).
+
 ```{r}
-data2010_nuts2 <- gini_nuts_2 %>%
- group_by(id_nuts2, year) %>%
- mutate( gdp_n2 = sum(GDP, na.rm = T),
- pop_n2 = sum(population, na.rm = T),
-GDP_per_capita = gdp_n2/pop_n2) %>%
- ungroup() %>%
- distinct(year, id_nuts0, id_nuts2, GDP_per_capita, gini_n2, gdp_n2, pop_n2) %>% 
+#| echo: false
+#| output: false
+
+# Making a dataframe that only contains observations from 2010, we're also filtering the data so that gini is bigger than 0.
+
+data2010 <- gini_nuts_2 |>
   filter(year == 2010, gini_n2 > 0)
+```
 
-data2010_nuts2 <- data2010_nuts2 %>% 
-  mutate(log_GDP_pc = log(GDP_per_capita))
+```{r}
+#| echo: false
+#| fig-cap: Regression statistics of all countries for the year 2010
+
+# Finding the natural logarithm of GDP per capita
+data2010 <- data2010 |> 
+  mutate(log_GDP_per_capita = log(GDP_per_capita))
+
+# lm(y ~ x) y = gini, x = log of GDP per capita, estimation of linear regression model
+lm <- lm(gini_n2 ~ log_GDP_per_capita, data = data2010)
+
+# Save coefficients
+b0_earn <- lm$coefficients[1]
+b1_earn <- lm$coefficients[2]
+
+# Regression statistics by using huxreg
+huxreg(All = lm, 
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
 ```
+
+```{r}
+#| echo: false
+#| output: false
+
+# Summary
+summary(lm)
+```
+
+Beta 0​ = 0.0551 (intercept)
+
+Beta 1 = 0.0193 (slope coefficient)
+
+```{r}
+#| echo: false
+#| fig-cap: Regression statistics of all countries seperately for the year 2010
+
+# Estimation for each country seperately 
+lmSK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "SK"))
+lmDK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "DK"))
+lmHU <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "HU"))
+lmPT <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "PT"))
+lmFR <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "FR"))
+
+# Regression statistics
+huxreg(SK = lmSK, 
+       DK = lmDK, 
+       HU = lmHU, 
+       PT = lmPT, 
+       FR = lmFR, 
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+```{r}
+#| echo: false
+#| output: false
+
+# Summary
+summary(lmSK)
+summary(lmDK)
+summary(lmPT)
+summary(lmFR)
+summary(lmHU)
+```
+
+### 2. Model Diagnostics
+
+We'll now look at some of the numbers we got from the linear regression model (for all countries combined):
+
+-   Coefficients:
+
+    -   Intercept is 0.0551 (expected value of gini when GDP per capita is 0). Statistically significant (indicated by p-value).
+
+    -   Estimated coefficient for GDP per capita is 0.0193, if the natural logarithm of GDP per capita increase by one, then the gini coefficient will increase by 0.0193.
+The p-value associated with the coefficent is however not statistically significant at 5% level.
+
+-   Goodnes of fit:
+
+    -   Multiple R-squared (0.08028) indicate that around 8% of the variability in the GINI coefficient is explained by the model.
+This is low, which can suggest that the model dosen't explain the variation in gini.
+
+    -   Adjusted R-squared is even lower (0.05474), is therefore expected that the model dosen't really explain the variance in the dependent variable (gini).
+
+-   Model significance:
+
+    -   The F-statistic (3.142) indicate the significance of the regression model. We can see here, that with the p-value of 0.08474, the model isn't significant at a 5% level.
+
+By looking at these numbers, we can see that the model may not reliably predict the gini coefficient. It also suggest that GDP per capita may not be a valid predictor of the gini coefficient.
+
+We also examined our selected countries seperately in order to see how the reliability and validity of the model might vary between countries.
+However, since there are too few observations for most of the countries, it makes it hard to make a conclusion of the reliability.
+What we can see from this examination, is that the models for Denmark, Portugal and Slovakia are not statistically significant.
+France seem however to have significant coefficiants, and Hungary have a moderate R-squared (but lacks significance in the slope).
+
+### 3. Visualization
+
+```{r}
+#| echo: false
+
+# Making two plots, first one is a residuals vs fitted plot, while the other is a normal Q-Q residuals plot
+plot(lm, which = 1:2) 
+```
+
+We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient.
+The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model.
+The residuals should be randomly scattered around the horisontal 0 line, something that can inducate that the variances of the error terms are constant.
+In our plot, the residuals are in some extent randomly distributed, and there is also no clear pattern; this suggest that there is likely no significant issues with heteroscedasticity or non-linearity.
+
+The other plot - normal Q-Q plot - is used to assess if the residuals of the linear model are normally distributed.
+In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed.
+
+In both of these plots, we can see that there are some outliers, these can affect the fit of the model.
+These outliers might come from regions that have a high GDP per capita compared to rest of the regions, like Paris in France that is a financial hub.
+
+The numbers we gathered earlier, as well as these plots, can help us explain and assert whether the classical OLS assumptions holds for our model:
+
+-   Linearity:
+
+    -
+
+-   Independence:
+
+    -
+
+-   Normality:
+
+    -
+
+-   Homoscedasticity:
+
+    -
+
+-   Multicollinearity:
+
+    -
+
+```{r}
+#| echo: false
+#| label: fig_scatterplot-all
+#| cap-location: bottom
+#| fig-cap: "Relationship Between GDP per Capita and Gini"
+
+# Visualization of the relationship between GDP per capita and gini
+ggplot(data2010, aes(x = log_GDP_per_capita, y = gini_n2)) +
+  geom_point(aes(color = log_GDP_per_capita), size = 3) +
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  labs(
+    title = "Relationship Between GDP per Capita and Gini",
+    x = "GDP per capita",
+    y = "Gini",
+    caption = "Red line indicates the fitted regression line"
+  ) +
+  theme_minimal()
+```
+
+In this plot we are visualizing the relationship between GDP per capita and gini by using a mix of the two previous plots. We can here as well, see extreme outliers.
+
+## 2B: Exploring Other Determinants of Inequity
+
+### 1. Data Acquistion
+
+In order to conduct a Multiple Linear Regression model, we need to have some independent variables to use in the model and compare them with the dependent variable.
+The first variable, education, can explain income inequality, since it can influence income distribution in a region.
+If access to education is unequal, then higher education levels might incease income disparities [@rodriguez-pose2008].
+The population density, our second variable, can explain inequality since regions with a higher population density might have different economic behaviours.
+Our last variable, rail network (infrastructure), can influence economic development and accessibility, which also can affect income inequality in a region [@chatterjee2012].
+
+```{r}
+#| echo: false
+#| output: false
+#| label: data acquistion
+
+# Importing datasets
+
+# Data for population density
+demo_dens <- read_csv("Raw_data/demo_dens.gz")
+
+# Data for education
+educ_data <- read_csv("Raw_data/educ.gz")
+
+# Data for rail network
+rail_data <- read_csv("Raw_data/rail.gz")
+
+# Renaming and tidying up the new dataframes 
+demo_dens <- demo_dens %>%
+  rename(year = TIME_PERIOD, id_nuts2 = geo, pop_density = OBS_VALUE)
+
+educ_data <- educ_data %>%
+  rename(year = TIME_PERIOD, id_nuts2 = geo, students_percentage = OBS_VALUE)
+
+rail_data <- rail_data %>%
+  rename(year = TIME_PERIOD, id_nuts2 = geo, rail_km = OBS_VALUE)
+```
+
+```{r}
+#| echo: false
+#| output: false
+
+# The NUTS 2 regions in France changed names in 2018.
+The dataset with education use the old codes while the other datasets use the new codes, we therefore had to change the names in order to be able to merge them together properly.
+
+# Changing NUTS 2 names for France
+nuts2_mapping <- c(
+  FR10 = "FR10",
+  FR21 = "FRF2",
+  FR22 = "FR22",
+  FR23 = "FRD2",
+  FR24 = "FRC1",
+  FR25 = "FRD2",
+  FR26 = "FRF2",
+  FR30 = "FR22",
+  FR41 = "FRF3",
+  FR42 = "FRF1",
+  FR43 = "FRF3",
+  FR51 = "FRG0",
+  FR52 = "FRH0",
+  FR53 = "FRI3",
+  FR61 = "FRI3",
+  FR62 = "FRJ2",
+  FR63 = "FRI3",
+  FR71 = "FRK2",
+  FR72 = "FRK1",
+  FR81 = "FRJ1",
+  FR82 = "FRL0",
+  FR83 = "FRM0"
+  )
+
+# Adding the new names to the dataset
+educ_fixed <- educ_data |>
+  mutate(id_nuts2 = recode(id_nuts2, !!!nuts2_mapping))
+
+# Selecting only the necessary columns
+demo_dens_selected <- select(demo_dens, year, id_nuts2, pop_density)
+educ_selected <- select(educ_fixed, year, id_nuts2, students_percentage)
+rail_selected <- select(rail_data, year, id_nuts2, rail_km)
+
+# Merging the datasets
+merged_datasets <- demo_dens_selected |>
+  left_join(educ_selected, by = c("year", "id_nuts2")) |>
+  left_join(rail_selected, by = c("year", "id_nuts2"))
+
+# Filter by year 2010
+merged_datasets_2010 <- merged_datasets |>
+  filter(year == 2010)
+
+# Merging with dataset containing GDP per capita and gini
+Merged_datasets_gini <- merged_datasets_2010 |>
+  left_join(data2010, by = c("year", "id_nuts2"))
+
+# cleaning dataset and selecting variables
+Merged_datasets_gini <- Merged_datasets_gini |>
+  select(year, id_nuts2, id_nuts0, gini_n2, GDP_per_capita, log_GDP_per_capita, students_percentage, pop_density, rail_km)
+
+# Removing NA values from gini_n2 
+Merged_datasets_gini <- Merged_datasets_gini |>
+  filter(!is.na(gini_n2))
+
+# Remove nuts 2 variables that gets repetead
+cleaned_data <- Merged_datasets_gini |>
+  distinct(id_nuts2, .keep_all = TRUE)
+```
+
+### 2. Multiple Linear Regression Model
+
+We will in this part do a Multiple Linear Regression model by using the variables education (in percentage of pupils and students in education, % of total population), population density and rail network in km.
+This model will tell us if these variables can help explain change in the gini coefficient.
+
+In both our simple linear regression model, and now in our multiple linear regression model, we use the logarithmic function which makes it easier to linearize the relationship between the variables.
+By using it for GDP per capita, we can reflect changes more effectively.
+For rail network and population density, the logarithm function ensure that the model capture proportional changes and deals better with the wide range of values.
+
+```{r}
+#| echo: false
+#| fig-cap: Multiple Linear Regression Model
+
+# Estimations with the help of several independent variables
+
+# To begin with, we'll look at how education can explain the gini coefficient
+lm1 <- lm(gini_n2 ~ log_GDP_per_capita + students_percentage, subset(cleaned_data))
+
+# Here we look at how both population density and education can explain the gini coefficient
+
+lmdens1 <- lm(gini_n2 ~ log_GDP_per_capita + students_percentage + log(pop_density), subset(cleaned_data))
+
+# Lastly, we look at how all of our selected variables, rail, education and population density, can explain the gini coefficient
+
+lm2 <- lm(gini_n2 ~ log_GDP_per_capita + students_percentage + log(pop_density) + log(rail_km), subset(cleaned_data))
+
+# Regression statistics
+huxreg("Model" = lm1, 
+       "Model 2" = lmdens1, 
+       "Model 3" = lm2, 
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+### 3. Model Interpretation
+
+Our first model in the Multiple Linear Regression model examine how education in addition to GDP per capita can help explain the gini coefficient.
+The second model look at both education, GDP per capita, and population density, while the third model examine them all and also add rail network.
+
+## Conclusion

### 2023-11-10 MSB104_GR_1_Final_Assignment_research_article.qmd
update

@@
 
-**#Esempel:Looking at the descriptive statistics we see in average a very low level of inequality within all regions.
-Even at the maximum inequality is modest with 0.35 spacial in comparison with other countries, see for example Lessmann and Seidel ([2017](https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course%20files/html/Assignments_example.html?context_id=10747~22507&context_type=Course&download=1&id=107470000002033066&inline=1#ref-lessmann2017regional)) for such a comparison.Studding the temporal component of our data we see that most regions follow a common trend in there level of inequality with modest changes over time.
-See for example the NUTS2 regions of France in [Figure 4](https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course%20files/html/Assignments_example.html?context_id=10747~22507&context_type=Course&download=1&id=107470000002033066&inline=1#fig-regional_gini_time).
-There are however some severe exceptions ...#**
-
 The next step is, estimating a Linear Regression for All Countries using cross sectional data from 2010, NUTS2.
@@
 
+For our model we have:
+
+$GINI = \beta_{0} + \beta_{1} \cdot \text{GDP per capita} + \epsilon$
+
 #### **Goodness of fit**
@@
 
-**Scale-Location Plot** is another way to check for homoscedasticity.
-The spread of the residuals should be roughly the same across all levels of the fitted values.
-If the spread increases or decreases with the fitted values, it suggests heteroscedasticity.
+```{r}
+#| echo: false
+#| output: false
 
-**Residuals vs.
-Leverage Plot** to identify influential observations that might unduly influence the regression.
-Points with high leverage or large residuals (outliers) can be identified.
-Cook's distance lines may be added to help identify points that have a large influence on the model.
+# Making a dataframe that only contains observations from 2010, we're also filtering the data so that gini is bigger than 0.
 
-## Part 2B: Exploring Other Determinants of Inequity
+data2010 <- gini_nuts_2 |>
+  filter(year == 2010, gini_n2 > 0)
+```
 
-### I. Data Acquisition
+```{r}
+#| echo: false
+#| fig-cap: Regression statistics of all countries for the year 2010
 
-Se data section for more specifications on the added population nut2, transport infrastructure and education.
+# Finding the natural logarithm of GDP per capita
+data2010 <- data2010 |> 
+  mutate(log_GDP_per_capita = log(GDP_per_capita))
 
-### II. Multiple Linear Regression Model
+# lm(y ~ x) y = gini, x = log of GDP per capita, estimation of linear regression model
+lm <- lm(gini_n2 ~ log_GDP_per_capita, data = data2010)
 
-Multiple Linear Regression (MLR) extends simple linear regression to incorporate multiple explanatory variables, allowing us to examine how multiple factors impact a dependent variable.
-Choosing a data set from the year 2010 that consists of various regions (FRANCE?), with data on each region's economic indicators, demographic variables, and other factors.
-Our aim is to understand how these variables collectively affect regional inequality.
+# Save coefficients
+b0_earn <- lm$coefficients[1]
+b1_earn <- lm$coefficients[2]
+
+# Regression statistics by using huxreg
+huxreg(All = lm, 
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+```{r}
+#| echo: false
+#| fig-cap: Regression statistics of all countries seperately for the year 2010
+
+# Estimation for each country seperately 
+lmSK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "SK"))
+lmDK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "DK"))
+lmHU <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "HU"))
+lmPT <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "PT"))
+lmFR <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "FR"))
+
+# Regression statistics
+huxreg(SK = lmSK, 
+       DK = lmDK, 
+       HU = lmHU, 
+       PT = lmPT, 
+       FR = lmFR, 
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+**Model Diagnostics**
+
+We'll now look at some of the numbers we got from the linear regression model (for all countries combined):
+
+-   Coefficients:
+
+    -   Intercept is 0.0551 (expected value of gini when GDP per capita is 0). Statistically significant (indicated by p-value).
+
+    -   Estimated coefficient for GDP per capita is 0.0193, if the natural logarithm of GDP per capita increase by one, then the gini coefficient will increase by 0.0193.
+The p-value associated with the coefficent is however not statistically significant at 5% level.
+
+-   Goodnes of fit:
+
+    -   Multiple R-squared (0.08028) indicate that around 8% of the variability in the GINI coefficient is explained by the model.
+This is low, which can suggest that the model dosen't explain the variation in gini.
+
+    -   Adjusted R-squared is even lower (0.05474), is therefore expected that the model dosen't really explain the variance in the dependent variable (gini).
+
+-   Model significance:
+
+    -   The F-statistic (3.142) indicate the significance of the regression model. We can see here, that with the p-value of 0.08474, the model isn't significant at a 5% level.
+
+By looking at these numbers, we can see that the model may not reliably predict the gini coefficient. It also suggest that GDP per capita may not be a valid predictor of the gini coefficient.
+
+We also examined our selected countries seperately in order to see how the reliability and validity of the model might vary between countries.
+However, since there are too few observations for most of the countries, it makes it hard to make a conclusion of the reliability.
+What we can see from this examination, is that the models for Denmark, Portugal and Slovakia are not statistically significant.
+France seem however to have significant coefficiants, and Hungary have a moderate R-squared (but lacks significance in the slope).
+
+**Ordinary Least Squares (OLS) Estimation**
+
+"The OLS method is employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
+This technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
+The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the data-set, without necessitating iterative procedures.
+
+Furthermore, when the classical linear regression assumptions are met, OLS guarantees that the estimators are BLUE, ensuring their unbiasedness and efficiency.
+This is particularly crucial in econometric analysis, where the precision and reliability of parameter estimates are paramount for policy implications and economic interpretations.
+The foundation assumptions of linear regression, including linearity, independence, homoscedasticity, and normality of residuals, are prerequisites for OLS to attain these desirable properties.
+It is imperative, therefore, to conduct diagnostic tests and assess the validity of these assumptions to ensure the robustness of the OLS estimators used in our analysis."
 
@@
 
+**Visualization**
+
+```{r}
+#| echo: false
+
+# Making two plots, first one is a residuals vs fitted plot, while the other is a normal Q-Q residuals plot
+plot(lm, which = 1:2) 
+```
+
+We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient, and that together with the regression statistics can help with discussing if the classical OLS assumptions hold for the model.
+The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model.
+The residuals should be randomly scattered around the horisontal 0 line, something that can inducate that the variances of the error terms are constant.
+In our plot, the residuals are in some extent randomly distributed, and there is also no clear pattern; this suggest that there is likely no significant issues with heteroscedasticity or non-linearity.
+
+The other plot - normal Q-Q plot - is used to assess if the residuals of the linear model are normally distributed.
+In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed.
+There are however some outliers in the tails, that suggest some variation from normality.
+
+In both of these plots, we can see that there are some outliers, these can affect the fit of the model.
+These outliers might come from regions that have a high GDP per capita compared to rest of the regions, like Paris in France that is a financial hub.
+
+```{r}
+#| echo: false
+#| label: fig_scatterplot-all
+#| cap-location: bottom
+#| fig-cap: "Relationship Between GDP per Capita and Gini"
+
+# Visualization of the relationship between GDP per capita and gini
+ggplot(data2010, aes(x = log_GDP_per_capita, y = gini_n2)) +
+  geom_point(aes(color = log_GDP_per_capita), size = 3) +
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  labs(
+    title = "Relationship Between GDP per Capita and Gini",
+    x = "GDP per capita",
+    y = "Gini",
+    caption = "Red line indicates the fitted regression line"
+  ) +
+  theme_minimal()
+```
+
+In this plot we are visualizing the relationship between GDP per capita and gini by using a mix of the two previous plots. We can here as well, see extreme outliers.
+
+## Part 2B: Exploring Other Determinants of Inequity
+
+### I. Data Acquisition
+
+Se data section for more specifications on the added population nut2, transport infrastructure and education.
+
+In order to conduct a Multiple Linear Regression model, we need to have some independent variables to use in the model and compare them with the dependent variable.
+The first variable, education, can explain income inequality, since it can influence income distribution in a region.
+If access to education is unequal, then higher education levels might incease income disparities [@rodriguez-pose2008].
+The population density, our second variable, can explain inequality since regions with a higher population density might have different economic behaviours.
+Our last variable, rail network (infrastructure), can influence economic development and accessibility, which also can affect income inequality in a region [@chatterjee2012].
+
+```{r}
+#| echo: false
+#| output: false
+#| label: data acquistion
+
+# Importing datasets
+
+# Data for population density
+demo_dens <- read_csv("Raw_data/demo_dens.gz")
+
+# Data for education
+educ_data <- read_csv("Raw_data/educ.gz")
+
+# Data for rail network
+rail_data <- read_csv("Raw_data/rail.gz")
+
+# Renaming and tidying up the new dataframes 
+demo_dens <- demo_dens %>%
+  rename(year = TIME_PERIOD, id_nuts2 = geo, pop_density = OBS_VALUE)
+
+educ_data <- educ_data %>%
+  rename(year = TIME_PERIOD, id_nuts2 = geo, students_percentage = OBS_VALUE)
+
+rail_data <- rail_data %>%
+  rename(year = TIME_PERIOD, id_nuts2 = geo, rail_km = OBS_VALUE)
+```
+
+```{r}
+#| echo: false
+#| output: false
+
+# The NUTS 2 regions in France changed names in 2018.
+The dataset with education use the old codes while the other datasets use the new codes, we therefore had to change the names in order to be able to merge them together properly.
+
+# Changing NUTS 2 names for France
+nuts2_mapping <- c(
+  FR10 = "FR10",
+  FR21 = "FRF2",
+  FR22 = "FR22",
+  FR23 = "FRD2",
+  FR24 = "FRC1",
+  FR25 = "FRD2",
+  FR26 = "FRF2",
+  FR30 = "FR22",
+  FR41 = "FRF3",
+  FR42 = "FRF1",
+  FR43 = "FRF3",
+  FR51 = "FRG0",
+  FR52 = "FRH0",
+  FR53 = "FRI3",
+  FR61 = "FRI3",
+  FR62 = "FRJ2",
+  FR63 = "FRI3",
+  FR71 = "FRK2",
+  FR72 = "FRK1",
+  FR81 = "FRJ1",
+  FR82 = "FRL0",
+  FR83 = "FRM0"
+  )
+
+# Adding the new names to the dataset
+educ_fixed <- educ_data |>
+  mutate(id_nuts2 = recode(id_nuts2, !!!nuts2_mapping))
+
+# Selecting only the necessary columns
+demo_dens_selected <- select(demo_dens, year, id_nuts2, pop_density)
+educ_selected <- select(educ_fixed, year, id_nuts2, students_percentage)
+rail_selected <- select(rail_data, year, id_nuts2, rail_km)
+
+# Merging the datasets
+merged_datasets <- demo_dens_selected |>
+  left_join(educ_selected, by = c("year", "id_nuts2")) |>
+  left_join(rail_selected, by = c("year", "id_nuts2"))
+
+# Filter by year 2010
+merged_datasets_2010 <- merged_datasets |>
+  filter(year == 2010)
+
+# Merging with dataset containing GDP per capita and gini
+Merged_datasets_gini <- merged_datasets_2010 |>
+  left_join(data2010, by = c("year", "id_nuts2"))
+
+# cleaning dataset and selecting variables
+Merged_datasets_gini <- Merged_datasets_gini |>
+  select(year, id_nuts2, id_nuts0, gini_n2, GDP_per_capita, log_GDP_per_capita, students_percentage, pop_density, rail_km)
+
+# Removing NA values from gini_n2 
+Merged_datasets_gini <- Merged_datasets_gini |>
+  filter(!is.na(gini_n2))
+
+# Remove nuts 2 variables that gets repetead
+cleaned_data <- Merged_datasets_gini |>
+  distinct(id_nuts2, .keep_all = TRUE)
+```
+
+### II. Multiple Linear Regression Model
+
+Multiple Linear Regression (MLR) extends simple linear regression to incorporate multiple explanatory variables, allowing us to examine how multiple factors impact a dependent variable.
+Choosing a data set from the year 2010 that consists of various regions, with data on each region's economic indicators, demographic variables, and other factors.
+Our aim is to understand how these variables collectively affect regional inequality.
+
+We will in this part do a Multiple Linear Regression model by using the variables education (in percentage of pupils and students in education, % of total population), population density and rail network in km.
+This model will tell us if these variables can help explain change in the gini coefficient.
+
+In both our simple linear regression model, and now in our multiple linear regression model, we use the logarithmic function which makes it easier to linearize the relationship between the variables.
+By using it for GDP per capita, we can reflect changes more effectively.
+For rail network and population density, the logarithm function ensure that the model capture proportional changes and deals better with the wide range of values.
+
+```{r}
+#| echo: false
+#| fig-cap: Multiple Linear Regression Model
+
+# Estimations with the help of several independent variables
+
+# To begin with, we'll look at how education can explain the gini coefficient
+lm1 <- lm(gini_n2 ~ log_GDP_per_capita + students_percentage, subset(cleaned_data))
+
+# Here we look at how both population density and education can explain the gini coefficient
+
+lmdens1 <- lm(gini_n2 ~ log_GDP_per_capita + students_percentage + log(pop_density), subset(cleaned_data))
+
+# Lastly, we look at how all of our selected variables, rail, education and population density, can explain the gini coefficient
+
+lm2 <- lm(gini_n2 ~ log_GDP_per_capita + students_percentage + log(pop_density) + log(rail_km), subset(cleaned_data))
+
+# Regression statistics
+huxreg("Model" = lm1, 
+       "Model 2" = lmdens1, 
+       "Model 3" = lm2, 
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
 ### Model specification
@@
 
-### Ordinary Least Squares (OLS) Estimation
+### III. Model Interpretation
 
-"The OLS method is employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
-This technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
-The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the data-set, without necessitating iterative procedures.
+Our first model in the Multiple Linear Regression model examine how education in addition to GDP per capita can help explain the gini coefficient.
+The second model look at both education, GDP per capita, and population density, while the third model examine them all and also add rail network.
 
-Furthermore, when the classical linear regression assumptions are met, OLS guarantees that the estimators are BLUE, ensuring their unbiasedness and efficiency.
-This is particularly crucial in econometric analysis, where the precision and reliability of parameter estimates are paramount for policy implications and economic interpretations.
-The foundation assumptions of linear regression, including linearity, independence, homoscedasticity, and normality of residuals, are prerequisites for OLS to attain these desirable properties.
-It is imperative, therefore, to conduct diagnostic tests and assess the validity of these assumptions to ensure the robustness of the OLS estimators used in our analysis."
+For model 1, the adjusted R-squared is 0.033, indicating that the model explains around 3% of the variability in the gini coefficient.
+Since this is relatively low, it suggests that model 1 is not suitable in explaining the variance in gini.
+Both variables have a p-value above 0.05, indicating that they are not statistically significant.
 
-#### **Model Validation with Cross-Sectional Data**
+For model 2, the adjusted R-squared is 0.300, which indicate that the model explains around 30% of the variability in the gini coefficient.
+This is pretty high, suggesting that model 2 is suitable in explaining the variance in gini, meaning that adding population density improves the model's explanatory power.
+The p-value for population density is less than 0.01, which means that its statistically significant.
 
-~~- \*\*Residual Analysis\*\*: Examine residuals for patterns to check for violations of linearity, independence, and homoscedasticity.~~
+The adjusted R-squared for model 3 is 0.370 \~ 37%; this suggest that this model has the greatest fit of all 3 models.
+Population density is significant at a 0.05 level.
+The p-value for rail is not below 0.05, and is therefore not statistically significant.
 
-~~- \*\*Outlier and Influence Diagnostics\*\*: Identify regions with high leverage or large residuals to ensure they are not disproportionately affecting the model.~~
-
-~~- \*\*Multicollinearity\*\*: Given that MLR includes multiple predictors, check for multicollinearity to ensure that the predictors are not too highly correlated with each other.~~
-
-~~\#### \*\*Predictive Power and Goodness of Fit\*\*~~
-
-~~- \*\*\\( R\^2 \\) and Adjusted \\( R\^2 \\)\*\*: Provide information on how well the independent variables explain the variation in the dependent variable.~~
-
-~~- \*\*F-Test\*\*: Tests whether at least one of the independent variables is statistically significant in explaining variation in the dependent variable.~~
-
-~~\#### \*\*Conclusion\*\*~~
-
-~~- The multiple linear regression model, estimated using OLS, allows for the comprehensive analysis of how various factors impact regional inequality.~~
-
-~~- Validating the assumptions of the OLS and properly interpreting the coefficients are vital steps in ensuring the reliability of the model.~~
-
-~~- The use of cross-sectional data from the year 2010 ensures that the analysis is grounded in a specific time context, making the results more interpretable and applicable.~~
-
-#### MLR Portugal
-
-#### MLR Denmark
-
-#### MLR France
-
-#### MLR Hungary
-
-#### MLR Slovakia
-
-### III. Model Interpretation
+To summarize, we can see that the population density is the variable that affects the gini coeffcient the most. The more people that live in an area, the bigger the economic inequality is.
 

### 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.qmd
Small changes in text

@@
 abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
-Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
+Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010.Lastly exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality"
 editor: visual
@@
 
-Lastly, potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions.
+Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).
+
+Lastly, it is worth noting potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions.
 The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017).
@@
 Furthermore,* regional growth and inequality have been important topics of research in Europe for many years.
-In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
+In recent decades, the European Union implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
 Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
@@
 Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014].
-However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia we may expect different growth trajectories, particularly after their EU accession.
 
@@
 EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic.
+In particular Portugal as the crisis hit them harder.
 These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.
@@
 
-@lessmann2017 regional inequality study address the importance of studying regional inequality.
+@lessmann2017 regional inequality study address the importance of studying these aspects.
 Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
@@
 
-Urbanization emerges as a crucial determinant, where bustling, populous around urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
+Urbanization emerges as a crucial determinant, where bustling, populous around urban centers usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
 
@@
 
-[@lessmann2017] discusses the determinant of Education in the context of human capital.
+[@lessmann2017] also discusses the determinant of Education in the context of human capital.
 The study mentions that human capital is the most important determinant of differences in regional development within countries.
-The study also notes that human capital creates positive externalities, which can spill over regional boundaries.
-Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
+The study also notes that human capital creates positive externalizes, which can spill over regional boundaries.
+Moreover, the quality of human capital within countries, as measured by the secondary-school enrollment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
 Lastly, promoting internal migration.
@@
 
-Furthermore, [@lessmann2017] provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
+Furthermore, @lessmann2017 provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
 The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
@@
 The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019].
-Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
+Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
 On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.
@@
 
-# Data
-
-Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files and filted the data by our preferences before downloading it.
-Furthermore, we filtered the dataset by choosing the years 2000 to 2020.
-Then, we selected the NUTS requiered levels region for Portugal, France, Hungary, Slovakia and Denmark.
-Finally, we specified the data to be in million Euro.
+# Data and Descriptive
+
+Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files.
+Selecting our countries of study and timeperiod.
+Selecting the required NUTS level and specified the data to be in million Euro.
 
@@
 Including data on the highest level of education successfully completed by the individuals of a given population.
-Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labour status of young people by years since completion of highest level of education.
-Calculated as annual averages of quarterly EU Labour Force Survey data (EU-LFS).
+Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labor status of young people by years since completion of highest level of education.
+Calculated as annual averages of quarterly EU Labor Force Survey data (EU-LFS).
 
@@
 ```
-
-# Econometric Approach
 
@@
 ```
+
+# Cross sectional estimates
 
@@
 
+# Alternative functional forms and panel estimation
+
+## Part 3A Testing Development effects across subsets
+
+## Part 3B Exploring Alternative Functional Forms
+
+## Part 3C Hetroskedasticity Testing and causality Discussion
+
+## Part 3D Panel Estimates
+
 # Discussion

### 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.qmd
:D

@@
 
+By exploring our pervious estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita.
+Using dummy variables to determine if the effect is significantly different in different subset of our countries.
+
 ## Part 3A Testing Development effects across subsets
@@
 
+### **The cuznets cuve**
+
+**Logarithmic transformation**
+
+The logarithmic transformation for all the variables allows the coefficients to be interpreted as elasticities, which measure the percentage change in the dependent variable associated with a one percent change in the independent variable [@lessmann2017].
+The use of logarithmic transformations is a common practice in econometrics to address issues such as heteroscedasticity and nonlinearity in the data.
+By using logarithmic transformations, we are able to estimate the relationship between regional inequality and development in a more robust and accurate way [@lessmann2017].
+
+**The quadratic term**
+
+Furthermore the quadratic term in the regression model is used to investigate the relationship
+between regional inequality and development [@lessmann2017].
+The results suggest an inverted U-shaped relationship between income and inequality, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
+
+**The cubic function**
+
+The cubic function in the regression model is used to investigate the relationship between regional inequality and development.
+@lessmann2017 suggest that regional inequalities increase again at very high levels of development, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
+The inclusion of a cubic term in a regression model allows the analysis of more complex dynamics, such as increasing or decreasing marginal effects [@wooldridge2020].
+However, interpreting these models can be more challenging, as the effect of a one-unit change in the predictor variable on the dependent variable is no longer constant and depends on the level of the predictor variable.
+
 ## Part 3C Hetroskedasticity Testing and causality Discussion
@@
 
+**Fixed effect models**
+
+Panel data estimation is a method used in econometrics to analyse data that involves observations over multiple time periods.
+In the context of panel data estimation specifics of fixed-effects and random effect are used.
+[@lessmann2017] agrues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function.
+However, the random-effects model allows for time-invariant unobserved heterogeneity across regions and is more appropriate when the fixed-effects model is too restrictive [@lessmann2017].
+
 # Discussion

### 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.qmd
:D

@@
 
+**Causality vs causation**
+
+While correlation indicates a relationship between two variables, it doesn\'t imply causation.
+In econometrics, it's a common pitfall to interpret correlation as causality.
+Panel data helps address this by allowing researchers to control for time-invariant factors and observe changes over time, which can provide stronger evidence for causal relationships.
+
+Panel data analysis helps address endogeneity issues, such as omitted variable bias, which can lead to spurious correlation.
+By tracking the same units over time, it's possible to differentiate whether a relationship is genuinely causal or merely a correlation coinciding in time.
+
+**Fixed-Effects and Random-Effects Models** in panel data analysis help in controlling for unobserved heterogeneity, thereby providing a clearer picture of whether there's a causal relationship.
+
+Overall, panel data analysis is a vital tool in econometrics for distinguishing between causality and correlation.
+By allowing for control over unobserved heterogeneity and providing a dynamic view of data over time, it offers a more nuanced and accurate understanding of the relationships between variables.
+This is crucial in econometric research, where the distinction between causality and mere correlation may have significant implications.
+
 ## Part 3D Panel Estimates

### 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.tex
:D

--- 2023-11-01 MSB104_GR_1_Final_Assignment_research_article.tex
@@
 \usepackage{xcolor}
+\usepackage{soul}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \usepackage{xcolor}
+\usepackage{caption}
+\usepackage{graphicx}
+\usepackage{siunitx}
+\usepackage{hhline}
+\usepackage{calc}
+\usepackage{tabularx}
+\usepackage{adjustbox}
+\usepackage{hyperref}
 \KOMAoption{captions}{tableheading}
@@
 This assignment aims to acquire, process, and analyze sub-national GDP
-and population data for a subset of European countries. Calculate GDP
-per capita and explore regional inequity using various descriptive
-statistics and visualizations.Furthermore examining the relationship
+and population data for a subset of European countries. Firstly
+(assignment 1), Calculateing GDP per capita and exploreing regional
+inequity using various descriptive statistics and
+visualizations.Secondly (assignment 2), examining the relationship
 between regional development and inequality, employing cross sectional
-estimation techniques for the year 2010.
+estimation techniques for the year 2010.Lastly exploring alternative
+functional forms and implementing panel estimation to analzise the
+relationship between regional development inequality
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, breakable, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 can, paradoxically, coexist with pockets of regional deprivation
-(\textbf{lessmann2017?}).
+(Lessmann \& Seidel, 2017).
 
@@
 concern, especially when crafting policies for equitable growth
-(\textbf{lessmann2017?}). A country's macro-level prosperity does not
+(Lessmann \& Seidel, 2017). A country's macro-level prosperity does not
 automatically guarantee that all its regions partake equally in this
@@
 get a more nuanced narrative about the state of regional economic
-disparities (\textbf{lessmann2017?}).
-
-Using time series data and cross sectional observations we investigate
-the GDP per capita and GINI trends in Portugal, France, Hungary,
-Slovenia and Denmark in the time period 2000 to 2020. The aims is to
-acquire, process, and analyze sub-national GDP and population data for a
-subset of European countries. Calculate GDP per capita and explore
-regional inequity using various descriptive statistics and
-visualizations
-
-\hypertarget{literature-review}{%
-\section{Literature review}\label{literature-review}}
-
-Regional growth and inequality have been important topics of research in
-Europe for many years. In recent decades, the European Union has
-implemented policies aimed at reducing regional disparities and
-promoting economic growth across its member states. Furthermore, despite
-these efforts, significant disparities in income and economic
-development persist across regions in Europe (\textbf{lessmann2017?}).
-
-When it comes to economic growth, several studies refer to numerous
-influencing factors. (\textbf{gennaioli2014?}) and
-(\textbf{crescenzi2012?}) states that regional growth, much like
-national growth, is significantly influenced by geography and human
-capital. Geography impacts trade, resource availability, and
-susceptibility to natural events. Human capital, encompassing skills and
-knowledge, drives innovation and productivity. Furthermore,
-(\textbf{gennaioli2014?}) mentions that wealthier nations like Denmark
-in our case, tend to experience faster regional convergence, suggesting
-national prosperity aids regional growth (\textbf{gennaioli2014?}).
-Convergence implies that poorer regions grow faster than richer ones,
-gradually reducing disparities in wealth and income. Also referred to as
-beta convergence in (\textbf{lessmann2017?}) paper.
-
-In the context of Europe, and our selected countries we might expect to
-see a wide variety of regional economic performances, especially within
-Hungary and Slovakia and between the western European countries. Capital
-market regulations also play a fundamental role, with effective
-frameworks promoting quicker regional convergence
-(\textbf{gennaioli2014?}). Taken this in to consideration we might
-expect a faster growth rate after the joining of the EU in 2003 and
-2004. That said the opening of the free labor market might also
-influence the migration of labor in Slovakia and Hungary. Not to mention
-the economic crisis hitting shortly after our eastern European countries
-joined the union.
-
-Whilst their recent EU membership might have posed certain challenges
-for Eastern European countries during the financial crisis, it also
-offered tools and resources that aided in their recovery in Slovakia and
-Hungary. As to the western Europeans such as Denmark, Portugal and
-France we expect to see a sharp decline in the economic growth.
-Likewise, with the recent covid19 pandemic in terms of financial
-aspects.
-
-Regarding the effects of the economic crisis, another paper notes that
-it led to a significant increase in regional unemployment rates across
-Europe. Reporting some regions experiencing unemployment rates as high
-as 30\% (\textbf{iammarino2019?}). The authors also discuss how the
-crisis exposed the vulnerability of regions that were heavily reliant on
-a single industry or sector, and how sheltered economies were better
-able to weather the storm (\textbf{iammarino2019?}). Another affect that
-surprised us was that the economic crisis led to increase inequality.
-The study ``\emph{the effects of financial crisis on regional
-inequality}'' indicate any type of financial crisis results in higher
-income inequality (\textbf{nguyen2022?}).
-
-(\textbf{lessmann2017?}) regional inequality study address the
-importance of studying regional inequality. In particular, its potential
-consequences, such as political tensions that can undermine social and
-political stability (\textbf{lessmann2017?}). The paper also discusses
-the relationship between regional inequality and personal income
-inequality and conflict risk. Furthermore, the paper provides us with
-insights into the inverted U-shaped relationship between regional
-inequality and the level of economic development in different country
-groups (\textbf{lessmann2017?}). The inverted U-shape also referred to
-as Kuznets Curve suggests that as an economy develops, regional
-inequality initially rises, reaches a peak, and then starts to decline.
-(\textbf{lessmann2017?}) paper also finds that there is an N-shaped
+disparities (Lessmann \& Seidel, 2017).
+
+Using time series data and cross-sectional observations we investigate
+Economic growth and Inequality trends in Portugal, France, Hungary,
+Slovenia and Denmark in the time period 2000 to 2020. The aims are to
+acquire, process, and analyse sub-national GDP, GINI and population data
+for a subset of European countries.
+
+The primary objective is to gain knowledge of econometric methods used
+for science in business through our given topic. Subsequently,
+econometric terms and definitions are explained in greater detail than
+it normally would in similar papers. Using GDP per capita and the Gini
+coefficient. Moreover, both are vital indicators in the field of
+economics and are frequently used, in a wide array of studies.
+Furthermore, to analyse and interpret the economic conditions and social
+disparities of different regions or countries. Their analysis provides
+crucial insights for policymakers, economists, and researchers in
+shaping and evaluating the effectiveness of socio-economic policies.
+
+The research question \textbf{\emph{``What are the regional economic
+growth and inequality measurements from 2000 to 2020, in selected
+European countries and how does other}} \textbf{\emph{determinants
+affect the results?{}``}} is key in guiding our study. Setting a solid
+foundation for a comprehensive study that not only measure our given
+parameters, but also a critically examination of the role of
+determinants such as transportation infrastructure and education.
+Lessmann \& Seidel (2017) states that transportation costs play an
+important role in agglomeration and income, and reduced transportation
+costs as a determinant of regional inequality.
+
+The sructure of our study is methodically designed to facilitate this
+exploration. In part 1A and B we calculate GDP per capita and explore
+regional inequality using various descriptive statistics and
+visualizations such as panel data (\textbf{\emph{Assignment\_1}}).
+
+In part 2A and B we use cross sectional method to examine the
+relationship between regional development and inequality, employing
+cross sectional estimation techniques for the year 2010
+(\textbf{\emph{Assignment\_2}}).
+
+Furthermore, exploring alternative functional forms and implementing
+panel estimation to analzise the relationship between regional
+development inequality (\textbf{\emph{Assignment\_3}}).
+
+Lastly, it is worth noting potential limitations may arise due to the
+use of ChatGPT and ChatGPT PDF, which have been utilized as tools to
+gain understanding and extract knowledge from academic papers, as well
+as to explain econometric terms and definitions. The text generated by
+these tools has been cross-checked to the best of our knowledge with our
+textbook and Andre Seidel, lecturer and author of `Regional Inequality,
+Convergence, and Its Determinants -- A View from Outer Space' (2017).
+Despite these precautions, there may have been instances where incorrect
+information was provided, and some errors could have been overlooked.
+The text has been meticulously rewritten to fit our style of writing,
+ensuring consistency and accuracy throughout our study.
+
+\hypertarget{literature-review-regional-economic-growth-inequality-and-determines-2000-2020}{%
+\section{Literature Review: Regional Economic Growth, Inequality and
+determines
+(2000-2020)}\label{literature-review-regional-economic-growth-inequality-and-determines-2000-2020}}
+
+Why some places grow and prosper is a fundamental question in social
+science research. This question motivated Adam Smith's work, \emph{An
+Inquiry into the Nature and Causes of the Wealth of Nations,} (Smith
+\href{https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4\#ref-CR32}{1863})
+and has been a major influence of the \emph{new economic geography}
+(Feldman, 2014)\emph{. Furthermore,} regional growth and inequality have
+been important topics of research in Europe for many years. In recent
+decades, the European Union implemented policies aimed at reducing
+regional disparities and promoting economic growth across its member
+states. Furthermore, despite these efforts, significant disparities in
+income and economic development persist across regions in Europe
+(Lessmann \& Seidel, 2017).
+
+\textbf{Economic Growth and Regional Disparities}
+
+Regional growth in Europe exhibits a complex pattern, influenced
+significantly by factors such as geography, human capital, and economic
+policies. Lessmann and Seidel's research offers insights into these
+dynamics, highlighting the persistence of income disparities, across
+regions despite overarching economic advancement (Lessmann \& Seidel,
+2017). Wealthier nations like Denmark tend to experience quicker
+regional convergence, a phenomenon supported by Gennaioli et al.'s
+findings (Gennaioli et al., 2014).
+
+\textbf{Financial crisis, and EU membership for Slovakia and Hungary}
+
+The 2008 financial crisis significantly disrupted regional growth and
+exacerbated income inequalities across Europe. Nguyen et al.~highlight
+that financial turmoil typically leads to increased income disparity; a
+trend observed globally (Nguyen, 2022).
+
+Focusing on Europe, we expect varied regional economic performances,
+especially when comparing Eastern European countries such as Hungary and
+Slovakia to their Western counterparts. Given the role of capital market
+regulations in promoting regional convergence, outlined by Gennaioli et
+al., we anticipate a potential growth acceleration. In particular for
+Eastern European countries following their EU accession in 2003 and 2004
+(Gennaioli et al., 2014). However, the financial crisis and subsequent
+labour market shifts may also have influenced these trends.
+
+EU membership provided Slovakia and Hungary with resources for economic
+recovery, while Western European countries, including Denmark, Portugal,
+and France, likely experienced significant economic declines due to the
+financial crisis, a pattern possibly mirrored in the recent COVID-19
+pandemic. In particular Portugal as the crisis hit them harder. These
+complexities highlight the importance of a detailed analysis to
+understand regional economic growth and inequality in challenging times.
+
+\textbf{Regional inequality}
+
+Another affect that surprised us was that the economic crisis led to
+increase inequality. The study ``\emph{the effects of financial crisis
+on regional inequality}'' indicate any type of financial crisis results
+in higher income inequality (Nguyen, 2022).
+
+Lessmann \& Seidel (2017) regional inequality study address the
+importance of studying these aspects. Its potential consequences, such
+as political tensions that can undermine social and political stability
+(Lessmann \& Seidel, 2017). The paper also discusses the relationship
+between regional inequality and personal income inequality and conflict
+risk.
+
+\textbf{Determinants of regional inequality}
+
+Geographic characteristics stand out as pivotal, with the natural and
+topographical elements of a region significantly swaying income
+distributions (Lessmann \& Seidel, 2017).
+
+Urbanization emerges as a crucial determinant, where bustling, populous
+around urban centers usually enjoy elevated income levels, contrasting
+starkly with the often-stark disparities witnessed in more rural and
+secluded areas (Lessmann \& Seidel, 2017).
+
+\emph{Transport infrastructure and Education}
+
+Transport infrastructure plays a central role in shaping regional
+economic landscapes. Improved connectivity fosters economic activities,
+reduces travel time, and enhances accessibility to markets and
+resources. Studies indicates that regions with robust transport
+infrastructure tend to attract more investments, generate employment,
+and ultimately contribute to regional prosperity.
+
+(Lessmann \& Seidel, 2017) also discusses the determinant of Education
+in the context of human capital. The study mentions that human capital
+is the most important determinant of differences in regional development
+within countries. The study also notes that human capital creates
+positive externalizes, which can spill over regional boundaries.
+Moreover, the quality of human capital within countries, as measured by
+the secondary-school enrollment rate, can decrease regional inequality
+by facilitating regional spillovers and convergence. Lastly, promoting
+internal migration.
+
+\textbf{Expected Inverted U- and N shaped curves}
+
+Furthermore, Lessmann \& Seidel (2017) provides us with insights into
+the inverted U-shaped relationship between regional inequality and the
+level of economic development in different country groups (Lessmann \&
+Seidel, 2017). The inverted U-shape also referred to as Kuznets Curve
+suggests that as an economy develops, regional inequality initially
+rises, reaches a peak, and then starts to decline.
+
+Lessmann \& Seidel (2017) paper also finds that there is an N-shaped
 relationship between regional inequality and economic development, which
 means that regional inequality increases again after the inverted
-U-shaped pattern has been completed (\textbf{lessmann2017?}).
-
-Lastly, (\textbf{iammarino2019?}) discusses regional inequalities in
-European countries. It presents evidence that inter-regional inequality
-has been increasing in the European Union since the turn of the
-millennium. The authors argue that this is due to the existence of
-several groups of regional economies in Europe that are structurally
-very different from one another (\textbf{iammarino2019?}). Additionally,
-the authors argue that there are some countries in the EU that are more
-evenly developed than others, and that a map of under performance or
-over-performance means less in a high-income but evenly developed
-country, such as in Denmark in our case (\textbf{iammarino2019?}). On
-the other hand, France has been a subject of discussion and policymaking
-for many years, regarding its pockets of extreme wealth. (finner ikke
-kilde)
-
-\hypertarget{data}{%
-\section{Data}\label{data}}
-
-Trough Eurostat, we download the datasets nama\_10r\_3gdp and
-demo\_r\_pjanggr3 as csv files and filted the data by our preferences
-before downloading it. Furthermore, we filtered the dataset by choosing
-the years 2000 to 2020. Then, we selected the NUTS 3 region for
-Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified
-the data to be in million Euro.
+U-shaped pattern has been completed (Lessmann \& Seidel, 2017).
+
+Iammarino et al. (2019) discusses regional inequalities in European
+countries. It presents evidence that inter-regional inequality has been
+increasing in the European Union since the turn of the millennium. The
+authors argue that this is due to the existence of several groups of
+regional economies in Europe that are structurally very different from
+one another (Iammarino et al., 2019). Additionally, the authors argue
+that there are some countries in the EU that are more evenly developed
+than others, and that a map of under performance or over-performance
+means less in a high-income but evenly developed country, such as in
+Denmark in our case (Iammarino et al., 2019). On the other hand, France
+has been a subject of discussion and policy making for many years,
+regarding its well-known pockets of extreme wealth in regions such as
+Cap d'Antibes.
+
+Overall, it is evident that regional economic growth and inequality in
+Europe from 2000 to 2020 have been influenced by a constellation of
+factors, with transport infrastructure and education emerging as
+significant determinants. Earlier studies provide us with a fundamental
+understanding of these dynamics, while also highlighting the nuanced
+impacts of economic crises on regional disparities. The integration of
+transport infrastructure and education into this discourse adds an
+additional layer of complexity, underscoring its potential as a catalyst
+for regional development and equality.
+
+\hypertarget{data-and-descriptive}{%
+\section{Data and Descriptive}\label{data-and-descriptive}}
+
+Trough Eurostat, we download the datasets nama\_10r\_3gdp ,
+demo\_r\_pjanggr3, demo\_r\_mlifexp, tran\_r\_net and edat\_lfse\_04 as
+csv files. Selecting our countries of study and timeperiod. Selecting
+the required NUTS level and specified the data to be in million Euro.
 
 \hypertarget{gdp}{%
-\subsubsection{GDP}\label{gdp}}
+\subsection{GDP}\label{gdp}}
 
@@
 countries. The data is mostly structured by year and region
-(\textbf{eurostat2023?}).
+\emph{Database - Eurostat} (n.d.).
 
@@
 reduced by subsidies on products (which are not allocated to specific
-sectors or industries) (\textbf{eurostat2023c?}). This also balances out
-in the total economy production account. In terms of methodology, while
-national accounts compile GDP from the expenditure side, regional
-accounts don't adopt this approach due to the complexities of accurately
-mapping inter-regional flows of goods and services.
+sectors or industries) \emph{Database - Eurostat} (n.d.). This also
+balances out in the total economy production account. In terms of
+methodology, while national accounts compile GDP from the expenditure
+side, regional accounts don't adopt this approach due to the
+complexities of accurately mapping inter-regional flows of goods and
+services.
 
@@
 and Purchasing Power Standards (PPS), figures per inhabitant and
-relative data compared to the EU Member States average
-(\textbf{eurostat2023c?}).
+relative data compared to the EU Member States average \emph{Database -
+Eurostat} (n.d.).
 
 \hypertarget{population}{%
-\subsubsection{Population}\label{population}}
+\subsection{Population}\label{population}}
 
@@
 also include demographic information on age and gender. Displayed in a
-year-by-region format, with yearly interval updates
-(\textbf{eurostat2023a?}).
+year-by-region format, with yearly interval updates \emph{Database -
+Eurostat} (n.d.).
 
@@
 covers 37 countries and is the central repository for demographic and
-migration-related data (\textbf{eurostat2023a?}). As well as, specific
-metrics gathered under UNIDEMO encompass population counts at the close
-of the calendar year and events such as births and deaths occurring
-within that year (\textbf{eurostat2023a?}). Additionally, data on
-marriages, divorces, and migration flows are recorded .
+migration-related data \emph{Database - Eurostat} (n.d.). As well as,
+specific metrics gathered under UNIDEMO encompass population counts at
+the close of the calendar year and events such as births and deaths
+occurring within that year \emph{Database - Eurostat} (n.d.).
+Additionally, data on marriages, divorces, and migration flows are
+recorded .
 
@@
 NUTS 2016 classification, which provides a detailed breakdown of the
-European Union's territory (\textbf{eurostat2021?}).
+European Union's territory \emph{Database - Eurostat} (n.d.).
+
+\hypertarget{demographic-nuts-2}{%
+\subsection{Demographic (Nuts 2)}\label{demographic-nuts-2}}
+
+Statistics on population change (demo\_r\_mlifexp) and~on population
+structure was collected from Euro Stat (\emph{Database - Eurostat},
+n.d.). Used to monitor demographic behavior within political, economic,
+social and cultural contexts at NUTS 2 level. Same policy and
+demographic classifications as previous population data set applies.
+
+\hypertarget{transportation-infrastructure-nuts-2}{%
+\subsection{Transportation Infrastructure (Nuts
+2)}\label{transportation-infrastructure-nuts-2}}
+
+This data-set (tran\_r\_net) presents recent data on the inland
+transport network of the European Union (EU), EFTA and candidate
+countries presenting motorways, railways and inland waterways
+(\emph{Database - Eurostat}, n.d.). Furthermore used as a varable of of
+interest (MLR) in our section growth and inequality. Moreover, to
+excamine if evolution of the transport network is closely linked to the
+general development of the economy.
+
+The collected data are compiled by the competent national authorities.
+Reported annually be the National Statistical Office, the Civil Aviation
+Authorities (air transport), the Ministry of Transport (inland
+waterways, railway and road networks, victims), and the National
+Maritime Administration (maritime transport) (\emph{Database -
+Eurostat}, n.d.).
+
+\hypertarget{education-nuts-2}{%
+\subsection{Education (Nuts 2)}\label{education-nuts-2}}
+
+This data set (edat\_lfse\_04) contains population by educational
+attainment level, sex and NUTS 2 regions (\%). Including data on the
+highest level of education successfully completed by the individuals of
+a given population. Furthermore, data on young people neither in
+employment nor in education and training -- NEET, early leavers from
+education and training and the labor status of young people by years
+since completion of highest level of education. Calculated as annual
+averages of quarterly EU Labor Force Survey data (EU-LFS).
 
 \hypertarget{nuts-classification}{%
-\subsubsection{NUTS classification}\label{nuts-classification}}
+\subsection{NUTS classification}\label{nuts-classification}}
 
@@
 statistics across Europe. The NUTS regions range from NUTS 0 Country
-level to NUTS 3 small units such as municipalities level
-(\textbf{eurostat2023a?}). ~
-
-\hypertarget{econometric-approach}{%
-\section{Econometric Approach}\label{econometric-approach}}
+level to NUTS 3 small units such as municipalities level Eurostat
+(2023). ~
 
@@
 descriptive statistics. Measuring regional income inequality is
-challenging due to heterogeneity of regions (\textbf{lessmann2017?}).
+challenging due to heterogeneity of regions (Lessmann \& Seidel, 2017).
 The number of regions in our data set varies largely in size and
@@
 becomes a minor issue because the country-level territorial
-heterogeneity is fixed (\textbf{lessmann2017?}).
+heterogeneity is fixed (Lessmann \& Seidel, 2017).
 
@@
 dynamic \emph{measures capture historical trends} (Paneldata)\emph{.}''
-(\textbf{wooldridge2020?})
+(Wooldridge, 2020)
 
@@
 \hypertarget{descriptive-statistics}{%
-\subsubsection{Descriptive statistics}\label{descriptive-statistics}}
+\subsection{Descriptive statistics}\label{descriptive-statistics}}
 
@@
 later steps when building and interpreting regression models
-(\textbf{wooldridge2020?}).
+(Wooldridge, 2020).
 
@@
 disparities. If this measure is small (close to 1), then it would mean
-that the different regions have relatively equal incomes
-(\textbf{wooldridge2020?}). If this measure is large, then the
-interpretation is more problematic, as it does not tell us if the high
-ratio is due to substantial variation in the distribution of per capita
-GDRPs or the presence of outliers. Nevertheless, maximum to minimum
-ratio (MMR) provides a quick, easy to comprehend, and politically
-powerful measure of regional income inequality.
+that the different regions have relatively equal incomes (Wooldridge,
+2020). If this measure is large, then the interpretation is more
+problematic, as it does not tell us if the high ratio is due to
+substantial variation in the distribution of per capita GDRPs or the
+presence of outliers. Nevertheless, maximum to minimum ratio (MMR)
+provides a quick, easy to comprehend, and politically powerful measure
+of regional income inequality.
 
@@
 which is crucial for understanding the reliability of estimations and
-predictions (\textbf{wooldridge2020?}).
+predictions (Wooldridge, 2020).
 
@@
 clearer picture of the ``typical'' value in situations where the data
-distribution is not symmetrical (\textbf{wooldridge2020?}).
+distribution is not symmetrical (Wooldridge, 2020).
 
@@
 \hypertarget{part-1a-sub-national-gdp-and-gdp-per-capita}{%
-\section{Part 1A: Sub-national GDP and GDP per
+\subsection{Part 1A: Sub-national GDP and GDP per
 Capita}\label{part-1a-sub-national-gdp-and-gdp-per-capita}}
@@
 \hypertarget{gdp-per-capita-calculation}{%
-\subsection{GDP per Capita
+\subsubsection{GDP per Capita
 Calculation}\label{gdp-per-capita-calculation}}
@@
 
-\hypertarget{portugal}{%
-\subsubsection{Portugal}\label{portugal}}
+\hypertarget{gdp-per-capita-portugal}{%
+\subsubsection{GDP per capita Portugal}\label{gdp-per-capita-portugal}}
 
@@
 Portugal's regions appears to be fairly consistent. There is however
-some regional variability. We can see that the regions around the big
-cities like Lisbon have a higher GDP per capita compared to some more
-rural areas. Since Lisbon is the capital of Portugal, there is probably
-a higher concentration of industries, making it a economic center (which
-again makes the GDP per capita higher).
+some regional variability. We can see signs of urbanisazation where the
+regions around the big cities like Lisbon have a higher GDP per capita
+compared to some more rural areas. Since Lisbon is the capital of
+Portugal, there is probably a higher concentration of industries, making
+it a economic center (which again makes the GDP per capita higher).
 
@@
 
-\hypertarget{france}{%
-\subsubsection{France}\label{france}}
+\hypertarget{gdp-per-capita-france}{%
+\subsubsection{GDP per capita France}\label{gdp-per-capita-france}}
 
@@
 
-\hypertarget{hungary}{%
-\subsubsection{Hungary}\label{hungary}}
+\hypertarget{gdp-per-capita-hungary}{%
+\subsubsection{GDP per capita Hungary}\label{gdp-per-capita-hungary}}
 
@@
 
-\hypertarget{slovakia}{%
-\subsubsection{Slovakia}\label{slovakia}}
+\hypertarget{gdp-per-capita-slovakia}{%
+\subsubsection{GDP per capita Slovakia}\label{gdp-per-capita-slovakia}}
 
@@
 
-\hypertarget{denmark}{%
-\subsubsection{Denmark}\label{denmark}}
+\hypertarget{gdp-per-capita-denmark}{%
+\subsubsection{GDP per capita Denmark}\label{gdp-per-capita-denmark}}
 
@@
 \hypertarget{part-1b-regional-inequity}{%
-\section{Part 1B: Regional Inequity}\label{part-1b-regional-inequity}}
+\subsection{Part 1B: Regional
+Inequity}\label{part-1b-regional-inequity}}
 
 \hypertarget{gini-coefficient-calculation}{%
-\subsection{Gini Coefficient
+\subsubsection{Gini Coefficient
 Calculation}\label{gini-coefficient-calculation}}
 
-With the use of the NUTS3 GDP per capita data and this formula:
-
-\(GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|\)
-
-we will compute the population-weighted GDP Gini coefficient for each
-European NUTS2 region in our assigned countries.
+In this part we will compute the population-weighted GDP Gini
+coefficient for each European NUTS2 region in our assigned countries.
 
@@
 easier to find the reasons as to why there is a difference between the
-regions (\textbf{hasell2023?}).
+regions Hasell \& Roser (2023).
+
+With the use of the NUTS3 GDP per capita data and this formula:
+
+\(GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|\)
 
@@
 
+\hypertarget{cross-sectional-estimates}{%
+\section{Cross sectional estimates}\label{cross-sectional-estimates}}
+
 \hypertarget{part-2a-growth-and-inequity}{%
-\section{Part 2A: Growth and
+\subsection{Part 2A: Growth and
 Inequity}\label{part-2a-growth-and-inequity}}
@@
 
-\textbf{\#Esempel:Looking at the descriptive statistics we see in
-average a very low level of inequality within all regions. Even at the
-maximum inequality is modest with 0.35 spacial in comparison with other
-countries, see for example Lessmann and Seidel
-(\href{https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course\%20files/html/Assignments_example.html?context_id=10747~22507\&context_type=Course\&download=1\&id=107470000002033066\&inline=1\#ref-lessmann2017regional}{2017})
-for such a comparison.Studding the temporal component of our data we see
-that most regions follow a common trend in there level of inequality
-with modest changes over time. See for example the NUTS2 regions of
-France in
-\href{https://a10747-2033066.cluster215.canvas-user-content.com/courses/10747~22507/files/10747~2033066/course\%20files/html/Assignments_example.html?context_id=10747~22507\&context_type=Course\&download=1\&id=107470000002033066\&inline=1\#fig-regional_gini_time}{Figure~4}.
-There are however some severe exceptions \ldots\#}
-
 The next step is, estimating a Linear Regression for All Countries using
@@
 can be crucial for identifying disparities or differences between the
-regions (\textbf{wooldridge2020?}).
+regions (Wooldridge, 2020).
 
@@
 model}}{Simple linear regression model}}\label{simple-linear-regression-model}}
+
+In this part of the paper, we will carry out a simple regression model
+and explore the effect of regional economic development, represented by
+𝑦𝑗 (GDP per capita), on regional inequality, represented by 𝐺𝐼𝑁 𝐼𝑊𝑗. We
+will do this to gain an understanding of the connection between economic
+growth indicators such as GDP per capita and inequality might have. We
+will explore if higher GDP per capita may lead to less or greater
+inequality, and gain an understanding to what extent these variables are
+related.
+
+Unlike the traditional Gini coefficient, which treats all individuals
+equally regardless of the population size of the region they reside in,
+the weighted Gini considers the population size of each region,
+assigning more weight to regions with larger populations.
+
+In the context of regional inequality, this is particularly important
+because it ensures that the income disparities in more populous regions
+have a proportionally larger impact on the overall measure of
+inequality. For instance, if a country has one region with a very high
+level of income per capita but a small population, and another region
+with a lower level of income per capita but a large population, the
+weighted Gini coefficient would reflect the inequality experienced by a
+larger portion of the country's population, providing a more accurate
+picture of the national income distribution.
 
@@
 
-\textbf{\#\#\#Vis formel\#\#}
-
-Y is the dependent variable (hva som perdikeres).
-
-X is the independent variable (input).
-
-β0\hspace{0pt} is the intercept.
-
-β1\hspace{0pt} is the slope of the line.
-
-ε represents the residuals or error in the prediction.
+\[
+Y_i = β_1 + β_2X_i + ε_i
+\]
+
+\(Y_i\) represent the dependent/ explained variable
+
+\(X_i\) represent the independent/ explanatory variable
+
+\(β_1\) represent the intercept/ constant
+
+\(β_2\) represent the slope coefficient
+
+\(ε_i\) represents the residuals or error in the prediction.
+
+\textbf{Intercept (}β0\hspace{0pt}): represents the value of Y when X is
+0.
+
+\textbf{Slope (}β1\hspace{0pt}): Indicates the change in Y for a
+one-unit change in X.
+
+For our model we have:
+
+\(GINI = \beta_{0} + \beta_{1} \cdot \text{GDP per capita} + \epsilon\)
+
+\hypertarget{goodness-of-fit}{%
+\paragraph{\texorpdfstring{\textbf{Goodness of
+fit}}{Goodness of fit}}\label{goodness-of-fit}}
+
+The goodness of fit in a simple linear regression model measures how
+well the regression line approximates the real data points. The
+regression line that best represents the data according to the least
+squares criterion, which minimizes the sum of the squared vertical
+distances of the points from the line.
+
+R\^{}2 is a statistical measure that represents the proportion of the
+variance for the dependent variable that's explained by the independent
+variable. It ranges from 0 to 1. A higher R\^{}2 value indicates a
+better fit of the model to the data.
+
+R\^{}2 = 0 The model does not explain any of the variability of the
+response data around its mean.
+
+R\^{}2 = 1 The model explains all the variability of the response data
+around its mean.
+
+If the assumptions for a simple linear regression are met, it indicates
+a good fit for the model. \ul{Plotting the data can provide a visual
+indication of the goodness of fit.} The points should fall around a
+straight line without clear patterns in the residuals. As it is crucial
+to verify these assumptions before proceeding with interpreting the
+results of the regression analysis.
+
+\textbf{Residuals vs.~Fitted Values Plot} to check for homoscedasticity
+and linearity. Ideally, this plot shows no pattern; the residuals are
+randomly scattered around the horizontal line at zero. If there's a
+pattern (like a curve or systematic spread of residuals), it suggests
+non-linearity or heteroscedasticity.
+
+\textbf{Normal Q-Q Plot} to check if the residuals are approximately
+normally distributed. The points should fall roughly along a straight
+line. Deviations from a straight line suggest deviations from normality.
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l}
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} All \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.055 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.010)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} log\_GDP\_per\_capita \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.019\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.011)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} r.squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.080\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} adj.r.squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.055\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} statistic \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3.142\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} p.value \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.085\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{2}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
+\hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
+Regression statistics of all countries for the year 2010
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l l l l l}
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} SK \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} DK \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} HU \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} PT \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} FR \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.069\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.127\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.075 ** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.079\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.028\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.010) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.085) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.010)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.015) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.026)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} log\_GDP\_per\_capita \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.018\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.124\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.049\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.008\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.107 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.034) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.059) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.029)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.030) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.026)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} r.squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.221\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.689\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.410\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.067\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.457\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} adj.r.squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.558\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.534\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.262\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.866\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.430\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} statistic \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.284\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 4.432\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.778\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.072\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 16.820\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} p.value \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.689\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.170\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.171\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.833\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.001\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{6}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
+\hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
+Regression statistics of all countries seperately for the year 2010
+
+\textbf{Model Diagnostics}
+
+We'll now look at some of the numbers we got from the linear regression
+model (for all countries combined):
 
@@
 \item
-  \textbf{Intercept (}β0\hspace{0pt}): represents the value of Y when X
-  is 0.
-
-  \textbf{Slope (}β1\hspace{0pt}): Indicates the change in Y for a
-  one-unit change in X.
+  Coefficients:
+
+  \begin{itemize}
+  \item
+    Intercept is 0.0551 (expected value of gini when GDP per capita is
+    0). Statistically significant (indicated by p-value).
+  \item
+    Estimated coefficient for GDP per capita is 0.0193, if the natural
+    logarithm of GDP per capita increase by one, then the gini
+    coefficient will increase by 0.0193. The p-value associated with the
+    coefficent is however not statistically significant at 5\% level.
   \end{itemize}
-
-``In linear regression, the objective is straightforward: determine the
-line that best represents the observed data. Furthermore, this `goodness
-of fit' line is characterized by coefficients β0\hspace{0pt} and β1,
-which represent the intercept and slope, respectively
-(\textbf{wooldridge2020?}). The optimal line is one that minimizes the
-sum of the squared differences between the observed values and the
-predictions made by the model (\textbf{wooldridge2020?}). This approach,
-known as the Ordinary Least Squares (OLS), has a solid mathematical
-foundation (\textbf{wooldridge2020?}). By taking the sum of these
-squared differences and setting their partial derivatives with respect
-to β0\hspace{0pt} and β1 to zero, we can derive the formula that give us
-these coefficients. And once these coefficients are in hand, they
-provide a direct equation for the line that best fits our data set
-(\textbf{wooldridge2020?}). One of the strengths of OLS is its
-closed-form solution, meaning we can directly compute the coefficients
-from the data without iterative procedures (\textbf{wooldridge2020?}).
-Moreover, under the foundamental assumptions of linear regression, OLS
-stands out as the Best Linear Unbiased Estimator (BLUE), emphasizing its
-reliability and efficiency in estimating the true underlying
-relationship (\textbf{wooldridge2020?}).''
-
-\hypertarget{individual-country-cases}{%
-\subsubsection{Individual country
-cases}\label{individual-country-cases}}
-
-\hypertarget{portugal-1}{%
-\paragraph{Portugal}\label{portugal-1}}
-
-\hypertarget{denmark-1}{%
-\paragraph{Denmark}\label{denmark-1}}
-
-\hypertarget{france-1}{%
-\paragraph{France}\label{france-1}}
-
-\hypertarget{hungary-1}{%
-\paragraph{Hungary}\label{hungary-1}}
-
-\hypertarget{slovakia-1}{%
-\paragraph{Slovakia}\label{slovakia-1}}
-
-\hypertarget{outlier-discussion}{%
-\subsubsection{\texorpdfstring{\textbf{Outlier
-discussion}}{Outlier discussion}}\label{outlier-discussion}}
-
-\textbf{\#\#Eksepel\ldots.When looking at the data it is obvious that we
-have too many regions with zero inequality. This seems strange. The
-reason for this is \ldots.. In the subsequent analysis it therefore
-would make scene to exclude these observations to not bias our finding.}
-
-\textbf{Looking at specific countries we observe allso further
-abnormalities like in \ldots. where \ldots{} As thees outliers are not
-do to a bias in our data we shold keep them in our data. However we need
-to test later the sensitivity of our findings regarding the presences of
-these borderline observations. \ldots\#\#\#\#}
+\item
+  Goodnes of fit:
+
+  \begin{itemize}
+  \item
+    Multiple R-squared (0.08028) indicate that around 8\% of the
+    variability in the GINI coefficient is explained by the model. This
+    is low, which can suggest that the model dosen't explain the
+    variation in gini.
+  \item
+    Adjusted R-squared is even lower (0.05474), is therefore expected
+    that the model dosen't really explain the variance in the dependent
+    variable (gini).
+  \end{itemize}
+\item
+  Model significance:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    The F-statistic (3.142) indicate the significance of the regression
+    model. We can see here, that with the p-value of 0.08474, the model
+    isn't significant at a 5\% level.
+  \end{itemize}
+\end{itemize}
+
+By looking at these numbers, we can see that the model may not reliably
+predict the gini coefficient. It also suggest that GDP per capita may
+not be a valid predictor of the gini coefficient.
+
+We also examined our selected countries seperately in order to see how
+the reliability and validity of the model might vary between countries.
+However, since there are too few observations for most of the countries,
+it makes it hard to make a conclusion of the reliability. What we can
+see from this examination, is that the models for Denmark, Portugal and
+Slovakia are not statistically significant. France seem however to have
+significant coefficiants, and Hungary have a moderate R-squared (but
+lacks significance in the slope).
+
+\textbf{Ordinary Least Squares (OLS) Estimation}
+
+``The OLS method is employed to identify the best-fitting linear
+relationship between the dependent and independent variables, aiming to
+minimize the sum of squared residuals. This technique ensures that the
+estimations of the intercept (β0) and slope (β1) yield the least
+possible cumulative discrepancy between the actual and predicted values.
+The strength of OLS lies in its closed-form solution, providing a
+straightforward computation of coefficients directly from the data-set,
+without necessitating iterative procedures.
+
+Furthermore, when the classical linear regression assumptions are met,
+OLS guarantees that the estimators are BLUE, ensuring their unbiasedness
+and efficiency. This is particularly crucial in econometric analysis,
+where the precision and reliability of parameter estimates are paramount
+for policy implications and economic interpretations. The foundation
+assumptions of linear regression, including linearity, independence,
+homoscedasticity, and normality of residuals, are prerequisites for OLS
+to attain these desirable properties. It is imperative, therefore, to
+conduct diagnostic tests and assess the validity of these assumptions to
+ensure the robustness of the OLS estimators used in our analysis.''
+
+\hypertarget{assumptions}{%
+\subsubsection{\texorpdfstring{\textbf{Assumptions}}{Assumptions}}\label{assumptions}}
+
+\textbf{I. Linearity}
+
+The relationship between the independent variable X and the dependent
+variable Y is linear. This means that changes in X are associated with
+proportional changes in Y. A straight line should provide a good fit to
+the data points when plotted on a graph.
+
+You can create a scatter plot of Y versus X and visually inspect whether
+a straight line could well represent the relationship. Alternatively,
+you can plot the residuals versus the fitted values and check for any
+obvious patterns or non-linearity.
+
+\textbf{II. Independence:}
+
+Observations are independent of each other. The value of Y for one
+observation should not depend on the value of Y for any other
+observation.
+
+This assumption is more about study design and data collection. Ensure
+that your observations are not correlated in time or space. For example,
+if you're analyzing economic data across regions, make sure that the
+regions are not influencing each other.
+
+\textbf{III. Homoscedasticity (Equal Variance):}
+
+The variance of the residuals (the errors) is constant across all levels
+of X. This means that the spread or ``width'' of the residuals should
+remain roughly the same across all values of the independent variable.
+
+A plot of residuals versus fitted values should show a random scatter
+and not display any funnel-like shapes (wider at one end).
+
+\textbf{IV. Normality of Residuals:}
+
+The residuals (the differences between observed and predicted values)
+are normally distributed. This assumption is particularly important for
+hypothesis testing and creating confidence intervals.
+
+A Quantile-Quantile (Q-Q) plot of the residuals can show if they follow
+a normal distribution. The points should fall roughly along a straight
+line.
+
+\textbf{V. No Perfect Multicollinearity (Specific to Multiple
+Regression)}
+
+In multiple regression settings, this assumes that no independent
+variable is a perfect linear function of any other independent
+variables. While this is more pertinent to multiple linear regression,
+it is crucial there because high correlation between independent
+variables can lead to unstable coefficient estimates.
+
+Checking the variance inflation factor (VIF) for each variable; a VIF
+above 5-10 indicates a problematic amount of collinearity.
+
+\textbf{Visualization}
+
+\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-29-1.pdf}
+
+\includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/unnamed-chunk-29-2.pdf}
+
+We've made two plots that can help us understand the relationship
+between GDP per capita and the Gini coefficient, and that together with
+the regression statistics can help with discussing if the classical OLS
+assumptions hold for the model. The first one is a plot that shows us
+residuals vs fitted values, which can help us check the homoscedasticity
+assumption of a linear regression model. The residuals should be
+randomly scattered around the horisontal 0 line, something that can
+inducate that the variances of the error terms are constant. In our
+plot, the residuals are in some extent randomly distributed, and there
+is also no clear pattern; this suggest that there is likely no
+significant issues with heteroscedasticity or non-linearity.
+
+The other plot - normal Q-Q plot - is used to assess if the residuals of
+the linear model are normally distributed. In our plot, most of the
+points follow the line closely, suggesting that the residuals are
+normally distributed. There are however some outliers in the tails, that
+suggest some variation from normality.
+
+In both of these plots, we can see that there are some outliers, these
+can affect the fit of the model. These outliers might come from regions
+that have a high GDP per capita compared to rest of the regions, like
+Paris in France that is a financial hub.
+
+\begin{verbatim}
+`geom_smooth()` using formula = 'y ~ x'
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics{MSB104_GR_1_Final_Assignment_research_article_files/figure-pdf/fig_scatterplot-all-1.pdf}
+
+}
+
+\caption{Relationship Between GDP per Capita and Gini}
+
+\end{figure}
+
+In this plot we are visualizing the relationship between GDP per capita
+and gini by using a mix of the two previous plots. We can here as well,
+see extreme outliers.
 
 \hypertarget{part-2b-exploring-other-determinants-of-inequity}{%
-\section{Part 2B: Exploring Other Determinants of
+\subsection{Part 2B: Exploring Other Determinants of
 Inequity}\label{part-2b-exploring-other-determinants-of-inequity}}
@@
 \subsubsection{I. Data Acquisition}\label{i.-data-acquisition}}
+
+Se data section for more specifications on the added population nut2,
+transport infrastructure and education.
+
+In order to conduct a Multiple Linear Regression model, we need to have
+some independent variables to use in the model and compare them with the
+dependent variable. The first variable, education, can explain income
+inequality, since it can influence income distribution in a region. If
+access to education is unequal, then higher education levels might
+incease income disparities (Rodriguez-Pose \& Tselios, 2008). The
+population density, our second variable, can explain inequality since
+regions with a higher population density might have different economic
+behaviours. Our last variable, rail network (infrastructure), can
+influence economic development and accessibility, which also can affect
+income inequality in a region (Chatterjee \& Turnovsky, 2012).
 
@@
 
+Multiple Linear Regression (MLR) extends simple linear regression to
+incorporate multiple explanatory variables, allowing us to examine how
+multiple factors impact a dependent variable. Choosing a data set from
+the year 2010 that consists of various regions, with data on each
+region's economic indicators, demographic variables, and other factors.
+Our aim is to understand how these variables collectively affect
+regional inequality.
+
+We will in this part do a Multiple Linear Regression model by using the
+variables education (in percentage of pupils and students in education,
+\% of total population), population density and rail network in km. This
+model will tell us if these variables can help explain change in the
+gini coefficient.
+
+In both our simple linear regression model, and now in our multiple
+linear regression model, we use the logarithmic function which makes it
+easier to linearize the relationship between the variables. By using it
+for GDP per capita, we can reflect changes more effectively. For rail
+network and population density, the logarithm function ensure that the
+model capture proportional changes and deals better with the wide range
+of values.
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l l l}
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model 2 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model 3 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.050\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.008\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.206\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.091) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.078)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.114)\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} log\_GDP\_per\_capita \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.020\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.011\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.017\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.014) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.012)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.012)\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} students\_percentage \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.000\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.006\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.001\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.004) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.004)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.006)\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} log(pop\_density) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.041 ** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.032 * \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.012)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.014)\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} log(rail\_km) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.020\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.018)\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} r.squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.098\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.370\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.467\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} adj.r.squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.033\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.300\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.370\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} statistic \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1.515\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 5.284\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 4.812\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} p.value \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.237\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.005\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.006\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{4}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
+\hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
+Multiple Linear Regression Model
+
+\hypertarget{model-specification}{%
+\subsubsection{Model specification}\label{model-specification}}
+
+\textbf{Understanding the Coefficients}
+
+Intercept beta\_0 Represents the expected value of the dependent
+variable when all independent variables are set to zero. Interpretation
+is often nonsensical in multiple regression if there is no meaningful
+condition where all predictors are zero.
+
+Slope Coefficients beta\_1, beta\_2, ..., \textbackslash beta\_k
+\textbackslash))**: Represent the expected change in the dependent
+variable for a one-unit change in the respective independent variable,
+holding all other variables constant.
+
 \hypertarget{iii.-model-interpretation}{%
@@
 
+Our first model in the Multiple Linear Regression model examine how
+education in addition to GDP per capita can help explain the gini
+coefficient. The second model look at both education, GDP per capita,
+and population density, while the third model examine them all and also
+add rail network.
+
+For model 1, the adjusted R-squared is 0.033, indicating that the model
+explains around 3\% of the variability in the gini coefficient. Since
+this is relatively low, it suggests that model 1 is not suitable in
+explaining the variance in gini. Both variables have a p-value above
+0.05, indicating that they are not statistically significant.
+
+For model 2, the adjusted R-squared is 0.300, which indicate that the
+model explains around 30\% of the variability in the gini coefficient.
+This is pretty high, suggesting that model 2 is suitable in explaining
+the variance in gini, meaning that adding population density improves
+the model's explanatory power. The p-value for population density is
+less than 0.01, which means that its statistically significant.
+
+The adjusted R-squared for model 3 is 0.370 \textasciitilde{} 37\%; this
+suggest that this model has the greatest fit of all 3 models. Population
+density is significant at a 0.05 level. The p-value for rail is not
+below 0.05, and is therefore not statistically significant.
+
+To summarize, we can see that the population density is the variable
+that affects the gini coeffcient the most. The more people that live in
+an area, the bigger the economic inequality is.
+
+\hypertarget{alternative-functional-forms-and-panel-estimation}{%
+\section{Alternative functional forms and panel
+estimation}\label{alternative-functional-forms-and-panel-estimation}}
+
+By exploring our pervious estimates in the relationship between regional
+development on regional inequality we will estimate the log of GDP per
+capita in a liner regression model, defining the variables as new
+determinants on the effects in regional economic development using
+logarithim of GDP per capita. Using dummy variables to determine if the
+effect is significantly different in different subset of our countries.
+
+\hypertarget{part-3a-testing-development-effects-across-subsets}{%
+\subsection{Part 3A Testing Development effects across
+subsets}\label{part-3a-testing-development-effects-across-subsets}}
+
+\hypertarget{part-3b-exploring-alternative-functional-forms}{%
+\subsection{Part 3B Exploring Alternative Functional
+Forms}\label{part-3b-exploring-alternative-functional-forms}}
+
+\hypertarget{the-cuznets-cuve}{%
+\subsubsection{\texorpdfstring{\textbf{The cuznets
+cuve}}{The cuznets cuve}}\label{the-cuznets-cuve}}
+
+\textbf{Logarithmic transformation}
+
+The logarithmic transformation for all the variables allows the
+coefficients to be interpreted as elasticities, which measure the
+percentage change in the dependent variable associated with a one
+percent change in the independent variable (Lessmann \& Seidel, 2017).
+The use of logarithmic transformations is a common practice in
+econometrics to address issues such as heteroscedasticity and
+nonlinearity in the data. By using logarithmic transformations, we are
+able to estimate the relationship between regional inequality and
+development in a more robust and accurate way (Lessmann \& Seidel,
+2017).
+
+\textbf{The quadratic term}
+
+Furthermore the quadratic term in the regression model is used to
+investigate the relationship between regional inequality and development
+(Lessmann \& Seidel, 2017). The results suggest an inverted U-shaped
+relationship between income and inequality, which implies that the
+relationship between inequality and development looks N-shaped instead
+of U-shaped (Lessmann \& Seidel, 2017).
+
+\textbf{The cubic function}
+
+The cubic function in the regression model is used to investigate the
+relationship between regional inequality and development. Lessmann \&
+Seidel (2017) suggest that regional inequalities increase again at very
+high levels of development, which implies that the relationship between
+inequality and development looks N-shaped instead of U-shaped (Lessmann
+\& Seidel, 2017). The inclusion of a cubic term in a regression model
+allows the analysis of more complex dynamics, such as increasing or
+decreasing marginal effects (Wooldridge, 2020). However, interpreting
+these models can be more challenging, as the effect of a one-unit change
+in the predictor variable on the dependent variable is no longer
+constant and depends on the level of the predictor variable.
+
+\hypertarget{part-3c-hetroskedasticity-testing-and-causality-discussion}{%
+\subsection{Part 3C Hetroskedasticity Testing and causality
+Discussion}\label{part-3c-hetroskedasticity-testing-and-causality-discussion}}
+
+\textbf{Causality vs causation}
+
+While correlation indicates a relationship between two variables, it
+doesn\textquotesingle t imply causation. In econometrics, it's a common
+pitfall to interpret correlation as causality. Panel data helps address
+this by allowing researchers to control for time-invariant factors and
+observe changes over time, which can provide stronger evidence for
+causal relationships.
+
+Panel data analysis helps address endogeneity issues, such as omitted
+variable bias, which can lead to spurious correlation. By tracking the
+same units over time, it's possible to differentiate whether a
+relationship is genuinely causal or merely a correlation coinciding in
+time.
+
+Overall, panel data analysis is a vital tool in econometrics for
+distinguishing between causality and correlation. By allowing for
+control over unobserved heterogeneity and providing a dynamic view of
+data over time, it offers a more nuanced and accurate understanding of
+the relationships between variables. This is crucial in econometric
+research, where the distinction between causality and mere correlation
+may have significant implications.
+
+\textbf{Fixed-Effects and Random-Effects Models} in panel data analysis
+help in controlling for unobserved heterogeneity, thereby providing a
+clearer picture of whether there's a causal relationship.
+
+\hypertarget{part-3d-panel-estimates}{%
+\subsection{Part 3D Panel Estimates}\label{part-3d-panel-estimates}}
+
+\textbf{Fixed effect models}
+
+Panel data estimation is a method used in econometrics to analyse data
+that involves observations over multiple time periods. In the context of
+panel data estimation specifics of fixed-effects and random effect are
+used. (Lessmann \& Seidel, 2017) agrues that the fixed-effects model is
+a reasonable approach when the differences between countries (or
+regions) can be viewed as parametric shifts of the regression function.
+However, the random-effects model allows for time-invariant unobserved
+heterogeneity across regions and is more appropriate when the
+fixed-effects model is too restrictive (Lessmann \& Seidel, 2017).
+
+\textbf{Random-Effects Models}
+
+Lessmann \& Seidel (2017) mention the use of a random-effects model to
+investigate the determinants of within-country changes in inequality.
+The random-effects model controls for several country-level fixed
+factors (national income, number of regions, and area) and fixed effects
+for various country groups (Lessmann \& Seidel, 2017). The advantage of
+the random-effects model is that the expected value of the
+country-specific effect is zero, which means that there is no need to
+apply any arbitrary data imputation procedure for the missing intercepts
+(Lessmann \& Seidel, 2017). However, this approach may come at the cost
+of founding the predictions on a slightly biased coefficient (Lessmann
+\& Seidel, 2017). The text also notes that the major coefficient of
+interest is not sensitive to applying either a fixed-effects model or a
+random-effects model with additional country and region information
+(Lessmann \& Seidel, 2017).
+
 \hypertarget{discussion}{%
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-chatterjee2012}{}}%
+Chatterjee, S., \& Turnovsky, S. J. (2012). Infrastructure and
+inequality. \emph{European Economic Review}, \emph{56}(8), 1730--1745.
+
+\leavevmode\vadjust pre{\hypertarget{ref-database}{}}%
+\emph{Database - eurostat}. (n.d.).
+\url{https://ec.europa.eu/eurostat/data/database}
+
+\leavevmode\vadjust pre{\hypertarget{ref-eurostat2023a}{}}%
+Eurostat. (2023). \emph{Population on 1 {January} by broad age group,
+sex and {NUTS} 3 region}.
+https://ec.europa.eu/eurostat/databrowser/view/demo\_r\_pjanaggr3/default/table?lang=en.
+
+\leavevmode\vadjust pre{\hypertarget{ref-feldman2014}{}}%
+Feldman, M. P. (2014). The character of innovative places:
+Entrepreneurial strategy, economic development, and prosperity.
+\emph{Small Business Economics}, \emph{43}(1), 9--20.
+\url{https://doi.org/10.1007/s11187-014-9574-4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gennaioli2014}{}}%
+Gennaioli, N., La Porta, R., Lopez De Silanes, F., \& Shleifer, A.
+(2014). Growth in regions. \emph{Journal of Economic Growth},
+\emph{19}(3), 259--309. \url{https://doi.org/10.1007/s10887-014-9105-9}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hasell2023}{}}%
+Hasell, J., \& Roser, M. (2023). Measuring inequality: {What} is the
+{Gini} coefficient? \emph{Our World in Data}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-iammarino2019}{}}%
+Iammarino, S., Rodriguez-Pose, A., \& Storper, M. (2019). Regional
+inequality in europe: Evidence, theory and policy implications.
+\emph{Journal of Economic Geography}, \emph{19}(2), 273--298.
+\url{https://doi.org/10.1093/jeg/lby021}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lessmann2017}{}}%
+Lessmann, C., \& Seidel, A. (2017). Regional inequality, convergence,
+and its determinants -- a view from outer space. \emph{European Economic
+Review}, \emph{92}, 110--132.
+\url{https://doi.org/10.1016/j.euroecorev.2016.11.009}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
+Nguyen, T. C. (2022). The effects of financial crisis on income
+inequality. \emph{Development Policy Review}, \emph{40}(6), e12600.
+\url{https://doi.org/10.1111/dpr.12600}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rodriguez-pose2008}{}}%
+Rodriguez-Pose, A., \& Tselios, V. (2008). \emph{{EDUCATION AND INCOME
+INEQUALITY IN THE REGIONS OF THE EUROPEAN UNION}}.
+https://click.endnote.com/viewer?doi=10.1111\%2Fj.1467-9787.2008.00602.x\&token=WzM0MzQ0MzksIjEwLjExMTEvai4xNDY3LTk3ODcuMjAwOC4wMDYwMi54Il0.upIV-Avuk84iH8myPb7qaaaJIQk.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wooldridge2020}{}}%
+Wooldridge, J. M. (2020). \emph{Introductory {Econometrics} - {A Modern
+Approach}} (7th ed.). {Cengage}.
+
+\end{CSLReferences}
 

### 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.qmd
:D

@@
 
-**Fixed-Effects and Random-Effects Models** in panel data analysis help in controlling for unobserved heterogeneity, thereby providing a clearer picture of whether there's a causal relationship.
-
 Overall, panel data analysis is a vital tool in econometrics for distinguishing between causality and correlation.
@@
 This is crucial in econometric research, where the distinction between causality and mere correlation may have significant implications.
+
+**Fixed-Effects and Random-Effects Models** in panel data analysis help in controlling for unobserved heterogeneity, thereby providing a clearer picture of whether there's a causal relationship.
 
@@
 
+**Random-Effects Models**
+
+@lessmann2017 mention the use of a random-effects model to investigate the determinants of
+within-country changes in inequality.
+The random-effects model controls for several country-level fixed factors (national income, number of regions, and area) and fixed effects for various country groups [@lessmann2017].
+The advantage of the random-effects model is that the expected value of the country-specific effect is zero, which means that there is no need to apply any arbitrary data imputation procedure for the missing intercepts [@lessmann2017].
+However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017].
+The text also notes that
+the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
+
 # Discussion

### 2023-11-22 MSB104_GR_1_Final_markdown.qmd
Assignment 3 :)

--- 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.qmd
@@
 abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.
-Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010.Lastly exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality"
+Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
 editor: visual
@@
 
-Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).
-
-Lastly, it is worth noting potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions.
+Lastly, potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions.
 The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017).
@@
 Furthermore,* regional growth and inequality have been important topics of research in Europe for many years.
-In recent decades, the European Union implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
+In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
 Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
@@
 Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014].
+However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia we may expect different growth trajectories, particularly after their EU accession.
 
@@
 EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic.
-In particular Portugal as the crisis hit them harder.
 These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.
@@
 
-@lessmann2017 regional inequality study address the importance of studying these aspects.
+@lessmann2017 regional inequality study address the importance of studying regional inequality.
 Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
@@
 
-Urbanization emerges as a crucial determinant, where bustling, populous around urban centers usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
+Urbanization emerges as a crucial determinant, where bustling, populous around urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
 
@@
 
-[@lessmann2017] also discusses the determinant of Education in the context of human capital.
+[@lessmann2017] discusses the determinant of Education in the context of human capital.
 The study mentions that human capital is the most important determinant of differences in regional development within countries.
-The study also notes that human capital creates positive externalizes, which can spill over regional boundaries.
-Moreover, the quality of human capital within countries, as measured by the secondary-school enrollment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
+The study also notes that human capital creates positive externalities, which can spill over regional boundaries.
+Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
 Lastly, promoting internal migration.
@@
 
-Furthermore, @lessmann2017 provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
+Furthermore, [@lessmann2017] provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
 The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
@@
 The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019].
-Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
+Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
 On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.
@@
 
-# Data and Descriptive
-
-Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files.
-Selecting our countries of study and timeperiod.
-Selecting the required NUTS level and specified the data to be in million Euro.
+# Data
+
+Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files and filted the data by our preferences before downloading it.
+Furthermore, we filtered the dataset by choosing the years 2000 to 2020.
+Then, we selected the NUTS requiered levels region for Portugal, France, Hungary, Slovakia and Denmark.
+Finally, we specified the data to be in million Euro.
 
@@
 Including data on the highest level of education successfully completed by the individuals of a given population.
-Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labor status of young people by years since completion of highest level of education.
-Calculated as annual averages of quarterly EU Labor Force Survey data (EU-LFS).
+Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labour status of young people by years since completion of highest level of education.
+Calculated as annual averages of quarterly EU Labour Force Survey data (EU-LFS).
 
@@
 ```
+
+# Econometric Approach
 
@@
 
-# Cross sectional estimates
-
 ## Part 2A: Growth and Inequity
@@
 The next step is making a data frame of NUTS2 2010 and grouping by NUTS0 country level variables, as well as allocating NUTS2 to selected countries.
-Lastly presenting data in **table XXX** showing a count of NUTS2 regions within each top-level region or country in the 2010 data.
+Lastly presenting data in tables showing a count of NUTS2 regions within each top-level region or country in the 2010 data.
 
@@
 
-**Model assumptions**
+#### **Model assumptions**
 
@@
 
-**Model specification**
+#### **Model specification**
 
@@
 ```
+
+#### **Model Diagnostics**
+
+We'll now look at some of the numbers we got from the linear regression model (for all countries combined):
+
+-   Coefficients:
+
+    -   Intercept is 0.0551 (expected value of gini when GDP per capita is 0). Statistically significant (indicated by p-value).
+
+    -   Estimated coefficient for GDP per capita is 0.0193, if the natural logarithm of GDP per capita increase by one, then the gini coefficient will increase by 0.0193.
+The p-value associated with the coefficent is however not statistically significant at 5% level.
+
+-   Goodnes of fit:
+
+    -   Multiple R-squared (0.08028) indicate that around 8% of the variability in the GINI coefficient is explained by the model.
+This is low, which can suggest that the model dosen't explain the variation in gini.
+
+    -   Adjusted R-squared is even lower (0.05474), is therefore expected that the model dosen't really explain the variance in the dependent variable (gini).
+
+-   Model significance:
+
+    -   The F-statistic (3.142) indicate the significance of the regression model. We can see here, that with the p-value of 0.08474, the model isn't significant at a 5% level.
 
@@
 
-**Model Diagnostics**
-
-We'll now look at some of the numbers we got from the linear regression model (for all countries combined):
-
--   Coefficients:
-
-    -   Intercept is 0.0551 (expected value of gini when GDP per capita is 0). Statistically significant (indicated by p-value).
-
-    -   Estimated coefficient for GDP per capita is 0.0193, if the natural logarithm of GDP per capita increase by one, then the gini coefficient will increase by 0.0193.
-The p-value associated with the coefficent is however not statistically significant at 5% level.
-
--   Goodnes of fit:
-
-    -   Multiple R-squared (0.08028) indicate that around 8% of the variability in the GINI coefficient is explained by the model.
-This is low, which can suggest that the model dosen't explain the variation in gini.
-
-    -   Adjusted R-squared is even lower (0.05474), is therefore expected that the model dosen't really explain the variance in the dependent variable (gini).
-
--   Model significance:
-
-    -   The F-statistic (3.142) indicate the significance of the regression model. We can see here, that with the p-value of 0.08474, the model isn't significant at a 5% level.
-
 By looking at these numbers, we can see that the model may not reliably predict the gini coefficient. It also suggest that GDP per capita may not be a valid predictor of the gini coefficient.
@@
 
-**Ordinary Least Squares (OLS) Estimation**
+### **Ordinary Least Squares (OLS) Estimation**
 
@@
 
-### **Assumptions**
+#### **Assumptions**
 
@@
 
-**Visualization**
+### **Visualization**
 
@@
 
-### I. Data Acquisition
+### Data Acquisition
 
@@
 
-### II. Multiple Linear Regression Model
+### Multiple Linear Regression Model
 
@@
 
-**Understanding the Coefficients**
+#### **Understanding the Coefficients**
 
@@
 
-### III. Model Interpretation
+### Model Interpretation
 
@@
 
-# Alternative functional forms and panel estimation
-
-By exploring our pervious estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita.
-Using dummy variables to determine if the effect is significantly different in different subset of our countries.
-
-## Part 3A Testing Development effects across subsets
-
-## Part 3B Exploring Alternative Functional Forms
-
-### **The cuznets cuve**
-
-**Logarithmic transformation**
-
-The logarithmic transformation for all the variables allows the coefficients to be interpreted as elasticities, which measure the percentage change in the dependent variable associated with a one percent change in the independent variable [@lessmann2017].
-The use of logarithmic transformations is a common practice in econometrics to address issues such as heteroscedasticity and nonlinearity in the data.
-By using logarithmic transformations, we are able to estimate the relationship between regional inequality and development in a more robust and accurate way [@lessmann2017].
-
-**The quadratic term**
-
-Furthermore the quadratic term in the regression model is used to investigate the relationship
-between regional inequality and development [@lessmann2017].
-The results suggest an inverted U-shaped relationship between income and inequality, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
-
-**The cubic function**
-
-The cubic function in the regression model is used to investigate the relationship between regional inequality and development.
-@lessmann2017 suggest that regional inequalities increase again at very high levels of development, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
-The inclusion of a cubic term in a regression model allows the analysis of more complex dynamics, such as increasing or decreasing marginal effects [@wooldridge2020].
-However, interpreting these models can be more challenging, as the effect of a one-unit change in the predictor variable on the dependent variable is no longer constant and depends on the level of the predictor variable.
-
-## Part 3C Hetroskedasticity Testing and causality Discussion
-
-## Part 3D Panel Estimates
-
-**Fixed effect models**
-
-Panel data estimation is a method used in econometrics to analyse data that involves observations over multiple time periods.
-In the context of panel data estimation specifics of fixed-effects and random effect are used.
-[@lessmann2017] agrues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function.
-However, the random-effects model allows for time-invariant unobserved heterogeneity across regions and is more appropriate when the fixed-effects model is too restrictive [@lessmann2017].
-
 # Discussion

### 2023-11-22 assignment_1.qmd
Assignment 3 :)

@@
 library(huxtable)
+library(sandwich)
+library(lmtest)
 ```
@@
 
+**Scale-Location Plot** is another way to check for homoscedasticity.
+The spread of the residuals should be roughly the same across all levels of the fitted values.
+If the spread increases or decreases with the fitted values, it suggests heteroscedasticity.
+
+**Residuals vs.
+Leverage Plot** to identify influential observations that might unduly influence the regression.
+Points with high leverage or large residuals (outliers) can be identified.
+Cook's distance lines may be added to help identify points that have a large influence on the model.
+
 $GINI = \beta_{0} + \beta_{1} \cdot \text{GDP per capita} + \epsilon$
@@
 
-We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient.
+We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient, and that together with the regression statistics can help with discussing if the classical OLS assumptions hold for the model.
 The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model.
@@
 In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed.
+There are however some outliers in the tails, that suggest some variation from normality.
+
+...
 
@@
 These outliers might come from regions that have a high GDP per capita compared to rest of the regions, like Paris in France that is a financial hub.
-
-The numbers we gathered earlier, as well as these plots, can help us explain and assert whether the classical OLS assumptions holds for our model:
-
--   Linearity:
-
-    -
-
--   Independence:
-
-    -
-
--   Normality:
-
-    -
-
--   Homoscedasticity:
-
-    -
-
--   Multicollinearity:
-
-    -
 
@@
 
-## Conclusion
+For model 1, the adjusted R-squared is 0.033, indicating that the model explains around 3% of the variability in the gini coefficient.
+Since this is relatively low, it suggests that model 1 is not suitable in explaining the variance in gini.
+
+# Assignment 3: Alternative Functional Forms and Panel Estimates
+
+## Part A: Testing Development Effects Across Subsets
+
+### Subset Analysis:
+
+To begin the third assignment, we will divide the data into different subsets.
+We have chosen to divide the countries based on size and population; we'll therefore have one subset with France, and one subset with the rest of the countries.
+We have several reasons for the dividing the countries like this:
+
+1.
+France is one of the largest countries in Europe, with a great economic and political influence.
+The larger size and population of France may lead to different economic dyamics compared to the other smaller countries.
+
+2.
+Larger countries might have a greater diversity of economic activities, wich affects the distribution of wealth and income differentely than in smaller countries with less diverse economies (source).
+We can see this by looking at how some of the NUTS 2 regions in France differ far more than in other countries.
+
+3.  The size and population can influence public policy, like social welfare, which impacts income distribution and inequality (source).
+
+4.
+Larger countries often have larger internal markets, which can affect economic development and income distribution differently compared to smaller economies, where ecternal trade might be more crucial.
+
+| Country  | Population (in million) | Size (km\^2) | NUTS 2 observations |
+|----------|-------------------------|--------------|---------------------|
+| France   | 67,75                   | 551 695      |                     |
+| Denmark  | 5,85                    | 42 952       |                     |
+| Slovakia | 5,4                     | 49 000       |                     |
+| Hungary  | 9,71                    | 93 000       |                     |
+| Portugal | 10,33                   | 92 000       |                     |
+
+```{r}
+#| echo: false
+#| output: false
+
+# Dataframe with just france
+gini_nuts_2_FR <- gini_nuts_2 |> 
+  filter(id_nuts0 == "FR", gini_n2 > 0)
+
+# Dataframe with the other countries
+gini_nuts_2_C <- gini_nuts_2 |> 
+  filter(id_nuts0 != "FR", gini_n2 > 0)
+```
+
+```{r}
+#| echo: false
+#| output: false
+
+# Descriptive analysis
+summary(gini_nuts_2_FR)
+summary(gini_nuts_2_C)
+```
+
+```{r}
+#| echo: false
+
+# Regression analysis
+# France
+lmFR <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_nuts_2_FR)
+# Others
+lmC <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_nuts_2_C)
+
+huxreg(France = lmFR,
+       Others = lmC,
+       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
+```
+
+```{r}
+#| echo: false
+#| output: false
+
+# Comparative analysis
+# Welch two sample t-test
+t.test <- t.test(gini_nuts_2_FR$gini_n2, gini_nuts_2_C$gini_n2)
+print(t.test)
+```
+
+### Subset Analysis Discussion:
+
+In order to discuss the differences between the two subsets, we have done different analysis: descriptive, regression and comparative.
+
+France has a mean gini coefficient of 0.06960, while the mean gini coefficient of the other countries is 0.062308.
+France has a slightly higher gini, meaning that they on average have a higher amount of economic inequality compared to the rest.
+The mean of GDP per capita is also higher in france (2.750 \> 1.743), suggesting that they also on average have a higher level of economic development (de har både høyere økonomisk ulikhet og økonomisk utvikling, hva er grunnen til dette?
+Kan se dette på regression analysis!!
+r quared har høy positiv korrelasjon).
+
+Looking at the regression analysis of France, we can see that the intercept is -0.034 with a p-value that is lower than 0.001, suggesting that the intercept is different from zero.
+The coefficient for log of GDP per capita is 0.110 with a p-value lower than 0.001; meaning that as the log of GDP per capita increase, the gini coefficient increase significantly, which suggest that a higher GDP per capita also gives us greater income inequality.
+Laslty, the R-squared is 0.490, meaning that around 49% of the variance in the gini coefficient can be explained by GDP per capita.
+*This is a relatively high value,* and suggest a strong model fit.
+
+## Part B: Exploring Alternative Functional Forms
+
+### Functional Form Exploration:
+
+```{r}
+
+# Logarithmic model
+log_model <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_filtered)
+
+# Quadratic model
+quad_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2), data = gini_filtered)
+
+# Cubic model
+cubic_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2) + I(GDP_per_capita^3), data = gini_filtered)
+
+# Check the summary of each model
+summary(log_model)
+```
+
+```{r}
+summary(quad_model)
+```
+
+```{r}
+summary(cubic_model)
+```
+
+###
+
+### Estimation and Visualization:
+
+```{r}
+# Quadratic model
+
+# Create a new data frame for plotting predicted values
+plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
+plot_data$gini_pred <- predict(quad_model, newdata = plot_data)
+
+# Plotting
+ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
+  geom_point(alpha = 0.5) +  # Plot the actual data points
+  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
+  labs(title = "Quadratic Model of GDP per Capita and Gini Coefficient",
+       x = "GDP per Capita",
+       y = "Gini Coefficient") +
+  theme_minimal()
+```
+
+```{r}
+# Logarithmic model
+
+# Create a new data frame for plotting predicted values
+plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
+plot_data$gini_pred <- predict(log_model, newdata = plot_data)
+
+# Plotting
+ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
+  geom_point(alpha = 0.5) +  # Plot the actual data points
+  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'red') +  # Add the regression line
+  scale_x_log10() +  # Logarithmic scale for the x-axis
+  labs(title = "Logarithmic Model of GDP per Capita and Gini Coefficient",
+       x = "GDP per Capita (log scale)",
+       y = "Gini Coefficient") +
+  theme_minimal()
+
+```
+
+```{r}
+
+# Create a new data frame for plotting predicted values
+plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
+plot_data$gini_pred <- predict(cubic_model, newdata = plot_data)
+
+# Plotting
+ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
+  geom_point(alpha = 0.5) +  # Plot the actual data points
+  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
+  labs(title = "Cubic Model of GDP per Capita and Gini Coefficient",
+       x = "GDP per Capita",
+       y = "Gini Coefficient") +
+  theme_minimal()
+
+```
+
+```{r}
+# Calculate residuals
+residuals_quad <- gini_filtered$gini_n2 - predict(quad_model, gini_filtered)
+residuals_cubic <- gini_filtered$gini_n2 - predict(cubic_model, gini_filtered)
+```
+
+```{r}
+# Create a dataframe for plotting residuals
+residuals_data <- data.frame(GDP_per_capita = gini_filtered$GDP_per_capita, 
+                             Residuals_Quad = residuals_quad, 
+                             Residuals_Cubic = residuals_cubic)
+
+# Plotting residuals
+ggplot(residuals_data) +
+  geom_line(aes(x = GDP_per_capita, y = Residuals_Quad, color = "Quadratic")) +
+  geom_line(aes(x = GDP_per_capita, y = Residuals_Cubic, color = "Cubic")) +
+  labs(title = "Comparison of Residuals",
+       x = "GDP per Capita",
+       y = "Residuals") +
+  theme_minimal() +
+  scale_color_manual(values = c("Quadratic" = "blue", "Cubic" = "red"))
+
+```
+
+```{r}
+sd(residuals_quad)
+sd(residuals_cubic)
+mean(residuals_quad^2)
+mean(residuals_cubic^2)
+```
+
+### Results Interpretation:
+
+## Part C: Heteroskedasticity Testing and Causality Discussion
+
+### Heteroskedasticity Testing:
+
+```{r}
+lm_01 <- lm(gini_n2 ~ GDP_per_capita, data = gini_filtered)
+```
+
+```{r}
+
+# Obtain fitted values and residuals from the model
+fitted_values <- fitted(lm_01)
+residuals <- resid(lm_01)
+
+# Create the plot
+plot(fitted_values, residuals, 
+     xlab = "Fitted Values", 
+     ylab = "Residuals", 
+     main = "Residuals vs Fitted Values for Linear Model",
+     pch = 20)
+abline(h = 0, col = "blue", lwd = 2) # Add a horizontal line at 0
+
+```
+
+```{r}
+
+# Breusch-Pagan test for the original linear model
+bptest_linear <- bptest(lm_01)
+
+# Breusch-Pagan test for the logarithmic model
+bptest_log <- bptest(log_model)
+
+# Breusch-Pagan test for the quadratic model
+bptest_quad <- bptest(quad_model)
+
+# Output the test results
+print(bptest_linear)
+print(bptest_log)
+print(bptest_quad)
+```
+
+The Breusch-Pagan test is used to detect heteroskedasticity in a regression model.
+The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables, which would violate one of the key assumptions of ordinary least squares (OLS) regression.
+
+```{r}
+
+# Calculate robust standard errors for the original linear model
+coeftest_lm_01 <- coeftest(lm_01, vcov = vcovHC(lm_01, type = "HC"))
+
+# Calculate robust standard errors for the logarithmic model
+coeftest_log_model <- coeftest(log_model, vcov = vcovHC(log_model, type = "HC"))
+
+# Calculate robust standard errors for the quadratic model
+coeftest_quad_model <- coeftest(quad_model, vcov = vcovHC(quad_model, type = "HC"))
+
+# Print out the results
+print(coeftest_lm_01)
+print(coeftest_log_model)
+print(coeftest_quad_model)
+
+```
+
+### Causality Discussion:
+
+## Part D: Panel Estimates
+
+### Panel Estimation Task:
+
+### Panel Estimation Analysis:
+
+### Panel Estimation Discussion:

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
:D

@@
 
-Furthermore the quadratic term in the regression model is used to investigate the relationship
-between regional inequality and development [@lessmann2017].
+Furthermore the quadratic term in the regression model is used to investigate the relationship between regional inequality and development [@lessmann2017].
 The results suggest an inverted U-shaped relationship between income and inequality, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
@@
 
-While correlation indicates a relationship between two variables, it doesn\'t imply causation.
+While correlation indicates a relationship between two variables, it doesn't imply causation.
 In econometrics, it's a common pitfall to interpret correlation as causality.
@@
 
-@lessmann2017 mention the use of a random-effects model to investigate the determinants of
-within-country changes in inequality.
+@lessmann2017 mention the use of a random-effects model to investigate the determinants of within-country changes in inequality.
 The random-effects model controls for several country-level fixed factors (national income, number of regions, and area) and fixed effects for various country groups [@lessmann2017].
@@
 However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017].
-The text also notes that
-the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
+The text also notes that the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
 

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
;-)

--- 2023-11-20 MSB104_GR_1_Final_Assignment_research_article.qmd
@@
 library(huxtable)
+library(plm)
 ```
@@
 In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***).
-
 In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).
-
 Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).
@@
 The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017).
-Despite these precautions, there may have been instances where incorrect information was provided, and some errors could have been overlooked.
+Despite these precautions, there may be instances where incorrect information was provided, and some errors could have been overlooked.
 The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.
@@
 
+Since panel data has a two-dimensional structure, with observations on multiple entities across several time periods.
+It enables us a more detailed and complex analyses compared to purely cross-sectional or time-series data.
+The data will then allow us to examine both cross-sectional (differences between subjects at a point in time) and longitudinal (changes within subjects over time) effects.
+For instance, this approach would be apt for studying how policy changes within a region affect its GDP and inequality levels.
+
 **Fixed effect models**
@@
 In the context of panel data estimation specifics of fixed-effects and random effect are used.
-[@lessmann2017] agrues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function.
+[@lessmann2017] argues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function.
 However, the random-effects model allows for time-invariant unobserved heterogeneity across regions and is more appropriate when the fixed-effects model is too restrictive [@lessmann2017].
@@
 However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017].
-The text also notes that
-the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
+@lessmann2017 also notes that the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
 

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
Mer oppdatert

@@
 By exploring our pervious estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita.
-Using dummy variables to determine if the effect is significantly different in different subset of our countries.
+
+Firstly, we will divide the data into different subsets.
+Using France (x) as a binary value often called a dummy variable.
+France is chosen since it is the country with the most observations in 2010.
+Dummy variables are used to explain qualitative variables in regression models [@wooldridge2020].
+
+Choosing France as a dummy variable amongst our group of countries could be justified based on several criteria.
+Firstly, France stands out as one of the largest economies in this group, both in terms of GDP and global economic influence.
+This distinction makes France a unique case compared to the smaller economies of Denmark, Portugal, Hungary, and Slovakia.
+Secondly, France's larger geographical area and population size provide different economic dynamics and complexities, which are not as pronounced in the smaller and more homogenous countries like Denmark, Portugal, Hungary, and Slovakia.
+Thirdly, France's economy is more diversified compared to the other countries listed, which might have more specialized or less varied economic structures.
+This diversity can lead to different patterns in wealth distribution and economic development.
+Lastly, France's policies, particularly in social welfare and economic regulation, and its larger domestic market, might influence economic outcomes differently than in the smaller economies, where external trade and different policy contexts play a more significant role.
 
 ## Part 3A Testing Development effects across subsets
+
+To elucidate the differences between France and other countries, we employed various analytical methods including descriptive, regression, and comparative analyses.
+Descriptively, France exhibits a mean Gini coefficient of 0.06960, marginally higher than the 0.062308 average of the other countries, suggesting more pronounced economic inequality.
+Correspondingly, France's mean GDP per capita is also higher (2.750 compared to 1.743), indicating a higher level of economic development.
+This juxtaposition of heightened economic inequality and development in France raises intriguing questions, addressed further in our regression analysis.
+
+In the regression analysis of France, the intercept is -0.034 with a statistically significant p-value below 0.001, indicating a non-zero intercept.
+More notably, the coefficient for the logarithm of GDP per capita stands at 0.110, with a p-value below 0.001, implying a significant increase in the Gini coefficient with rising GDP per capita.
+This suggests that higher GDP per capita correlates with greater income inequality.
+The R-squared value of 0.490 indicates that about 49% of the variance in the Gini coefficient is accounted for by GDP per capita, reflecting a strong model fit.
+This substantial correlation necessitates further investigation into the dynamics linking economic development with inequality in France.
 
@@
 
+### Functional Form Exploration
+
+```{r}
+#| echo: false
+#| output: false
+# Logarithmic model
+log_model <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_filtered)
+
+# Quadratic model
+quad_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2), data = gini_filtered)
+
+# Cubic model
+cubic_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2) + I(GDP_per_capita^3), data = gini_filtered)
+
+# Check the summary of each model
+summary(log_model)
+```
+
+```{r}
+summary(cubic_model)
+```
+
+```{r}
+summary(quad_model)
+```
+
+### Estimation and Visualization
+
+```{r}
+#| echo: false
+#| output: false
+# Quadratic model
+
+# Create a new data frame for plotting predicted values
+plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
+plot_data$gini_pred <- predict(quad_model, newdata = plot_data)
+
+# Plotting
+ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
+  geom_point(alpha = 0.5) +  # Plot the actual data points
+  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
+  labs(title = "Quadratic Model of GDP per Capita and Gini Coefficient",
+       x = "GDP per Capita",
+       y = "Gini Coefficient") +
+  theme_minimal()
+```
+
+```{r}
+# Logarithmic model
+
+# Create a new data frame for plotting predicted values
+plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
+plot_data$gini_pred <- predict(log_model, newdata = plot_data)
+
+# Plotting
+ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
+  geom_point(alpha = 0.5) +  # Plot the actual data points
+  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'red') +  # Add the regression line
+  scale_x_log10() +  # Logarithmic scale for the x-axis
+  labs(title = "Logarithmic Model of GDP per Capita and Gini Coefficient",
+       x = "GDP per Capita (log scale)",
+       y = "Gini Coefficient") +
+  theme_minimal()
+```
+
+```{r}
+#| echo: false
+#| output: false
+# Create a new data frame for plotting predicted values
+plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
+plot_data$gini_pred <- predict(cubic_model, newdata = plot_data)
+
+# Plotting
+ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
+  geom_point(alpha = 0.5) +  # Plot the actual data points
+  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
+  labs(title = "Cubic Model of GDP per Capita and Gini Coefficient",
+       x = "GDP per Capita",
+       y = "Gini Coefficient") +
+  theme_minimal()
+```
+
+```{r}
+#| echo: false
+#| output: false
+# Calculate residuals
+residuals_quad <- gini_filtered$gini_n2 - predict(quad_model, gini_filtered)
+residuals_cubic <- gini_filtered$gini_n2 - predict(cubic_model, gini_filtered)
+```
+
+```{r}
+#| echo: false
+#| output: false
+# Create a dataframe for plotting residuals
+residuals_data <- data.frame(GDP_per_capita = gini_filtered$GDP_per_capita, 
+                             Residuals_Quad = residuals_quad, 
+                             Residuals_Cubic = residuals_cubic)
+
+# Plotting residuals
+ggplot(residuals_data) +
+  geom_line(aes(x = GDP_per_capita, y = Residuals_Quad, color = "Quadratic")) +
+  geom_line(aes(x = GDP_per_capita, y = Residuals_Cubic, color = "Cubic")) +
+  labs(title = "Comparison of Residuals",
+       x = "GDP per Capita",
+       y = "Residuals") +
+  theme_minimal() +
+  scale_color_manual(values = c("Quadratic" = "blue", "Cubic" = "red"))
+```
+
+```{r}
+#| echo: false
+#| output: false
+sd(residuals_quad)
+sd(residuals_cubic)
+mean(residuals_quad^2)
+mean(residuals_cubic^2)
+```
+
+**Results Interpretation**
+
 ## Part 3C Hetroskedasticity Testing and causality Discussion
@@
 
-**Fixed-Effects and Random-Effects Models** in panel data analysis help in controlling for unobserved heterogeneity, thereby providing a clearer picture of whether there's a causal relationship.
+**Heteroskedasticity Testing**
+
+```{r}
+#| echo: false
+#| output: false
+#hetro
+lm_01 <- lm(gini_n2 ~ GDP_per_capita, data = gini_filtered)
+```
+
+```{r}
+#| echo: false
+#| output: false
+# Obtain fitted values and residuals from the model
+fitted_values <- fitted(lm_01)
+residuals <- resid(lm_01)
+
+# Create the plot
+plot(fitted_values, residuals, 
+     xlab = "Fitted Values", 
+     ylab = "Residuals", 
+     main = "Residuals vs Fitted Values for Linear Model",
+     pch = 20)
+abline(h = 0, col = "blue", lwd = 2) # Add a horizontal line at 0
+```
+
+The Breusch-Pagan test is used to detect heteroskedasticity in a regression model.
+The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables, which would violate one of the key assumptions of ordinary least squares (OLS) regression.
+
+```{r}
+#| echo: false
+#| output: false
+# Breusch-Pagan test for the original linear model
+bptest_linear <- bptest(lm_01)
+
+# Breusch-Pagan test for the logarithmic model
+bptest_log <- bptest(log_model)
+
+# Breusch-Pagan test for the quadratic model
+bptest_quad <- bptest(quad_model)
+
+# Output the test results
+print(bptest_linear)
+print(bptest_log)
+print(bptest_quad)
+```
+
+```{r}
+#| echo: false
+#| output: false
+# Calculate robust standard errors for the original linear model
+coeftest_lm_01 <- coeftest(lm_01, vcov = vcovHC(lm_01, type = "HC"))
+
+# Calculate robust standard errors for the logarithmic model
+coeftest_log_model <- coeftest(log_model, vcov = vcovHC(log_model, type = "HC"))
+
+# Calculate robust standard errors for the quadratic model
+coeftest_quad_model <- coeftest(quad_model, vcov = vcovHC(quad_model, type = "HC"))
+
+# Print out the results
+print(coeftest_lm_01)
+print(coeftest_log_model)
+print(coeftest_quad_model)
+```
 
@@
 For instance, this approach would be apt for studying how policy changes within a region affect its GDP and inequality levels.
+Fixed-Effects and Random-Effects Models in panel data analysis help in controlling for unobserved heterogeneity, thereby providing a clearer picture of whether there's a causal relationship.
 

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
:D

@@
 
-By exploring our pervious estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita.
-
-Firstly, we will divide the data into different subsets.
-Using France (x) as a binary value often called a dummy variable.
-France is chosen since it is the country with the most observations in 2010.
-Dummy variables are used to explain qualitative variables in regression models [@wooldridge2020].
+By exploring our previous estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita.
+
+Firstly, the data is devided into different subsets. Using France (x) as a binary value also called a dummy variable [@wooldridge2020].
 
@@
 ```{r}
+#| echo: false
+#| output: false
 # Logarithmic model

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
hetroskedasketi

@@
 
+The residuals vs.
+fitted values plot indicates potential issues with our linear model, which might include heteroskedasticity and model misspecification.
+The observed patterns and their possible implications, suggests that our model might be missing a nonlinear relationship between the dependent and independent variables.
+This is a sign of model misspecification, indicating that the current linear model may not be the best fit for our data.
+The variance in the residuals first increasing, then decreasing, and increasing again, implies that the error terms have non-constant variance, which is a sign of heteroskedasticity.
+In a well-specified homoscedastic model, the spread of the residuals should be roughly equal across all levels of fitted values.
+
+The sets of disconnected residuals further out could indicate that there are distinct subgroups within our data that are behaving differently.
+It might suggest that the relationship between the variables differs across these groups or that there are influential outliers or leverage points affecting the model differently in these areas.
+
 The Breusch-Pagan test is used to detect heteroskedasticity in a regression model.

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
final update assignment 3

@@
 library(plm)
+library(sandwich)
+library(lmtest)
 ```
@@
 
-# Alternative functional forms and panel estimation
-
-By exploring our previous estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita.
-
-Firstly, the data is devided into different subsets. Using France (x) as a binary value also called a dummy variable [@wooldridge2020].
+# Assignment 3: Alternative Functional Forms and Panel Estimates
+
+## Testing Development Effects Across Subsets
+
+### Subset Analysis
+
+To begin the third assignment, we will explore if the effect of development and economic inequality is significantly different between different subsets in our dataset.
+For our assigned countries, we figured the best way to look at differences between them, is to examine the differences between our biggest country (France) compared to the rest of the countries.
+In order to compare the two different subsets, we will create a dummy variable.
+This dummy variable will help us with explaining qualitative values in the regression model.
 
@@
 This distinction makes France a unique case compared to the smaller economies of Denmark, Portugal, Hungary, and Slovakia.
-Secondly, France's larger geographical area and population size provide different economic dynamics and complexities, which are not as pronounced in the smaller and more homogenous countries like Denmark, Portugal, Hungary, and Slovakia.
+Secondly, France's larger geographical area and population size provide different economic dynamics and complexities, which are not as pronounced in smaller and more homogenous countries.
 Thirdly, France's economy is more diversified compared to the other countries listed, which might have more specialized or less varied economic structures.
@@
 
-## Part 3A Testing Development effects across subsets
-
-To elucidate the differences between France and other countries, we employed various analytical methods including descriptive, regression, and comparative analyses.
-Descriptively, France exhibits a mean Gini coefficient of 0.06960, marginally higher than the 0.062308 average of the other countries, suggesting more pronounced economic inequality.
-Correspondingly, France's mean GDP per capita is also higher (2.750 compared to 1.743), indicating a higher level of economic development.
-This juxtaposition of heightened economic inequality and development in France raises intriguing questions, addressed further in our regression analysis.
-
-In the regression analysis of France, the intercept is -0.034 with a statistically significant p-value below 0.001, indicating a non-zero intercept.
-More notably, the coefficient for the logarithm of GDP per capita stands at 0.110, with a p-value below 0.001, implying a significant increase in the Gini coefficient with rising GDP per capita.
-This suggests that higher GDP per capita correlates with greater income inequality.
-The R-squared value of 0.490 indicates that about 49% of the variance in the Gini coefficient is accounted for by GDP per capita, reflecting a strong model fit.
-This substantial correlation necessitates further investigation into the dynamics linking economic development with inequality in France.
-
-## Part 3B Exploring Alternative Functional Forms
-
-### **The cuznets cuve**
+```{r}
+#| echo: false
+
+# Dummy variable
+gini_filtered$France <- ifelse(gini_filtered$id_nuts0 == "FR", 1,0)
+gini_filtered$Other <- ifelse(gini_filtered$id_nuts0 == "FR", 0,1)
+
+# Regression analysis
+gini_filtered <- gini_filtered
+lmFR = lm(gini_n2 ~ GDP_per_capita, filter(gini_filtered, France == 1))
+lmC = lm(gini_n2 ~ GDP_per_capita, filter(gini_filtered, France == 0))
+
+huxreg(list("France" = lmFR, "All other countries" = lmC), statistics = c(N = "nobs", R2 = "r.squared", FStat = "statistic", PValue = "p.value"), note = "Note: {stars} T statistics in brackets.")
+```
+
+### Subset Analysis Discussion
+
+For France, the regression model shows us a significant and positive relationship between GDP per capita and the Gini coefficient.
+The intercept is statistically significant, but negative.
+The positive coefficient of GDP per capita of 0.028 is highly significant with a p-value lower than 0.001; this suggests that when GDP per capita in France rises, then economic inequality rises correspondingly as well.
+We can also see this by looking at the R-squared of 0.637, indicating that around 63.7% of the variation in the economic inequality can be explained by changes in economic development (GDP per capita).
+To summarize our findings for France, we can see that economic growth in France can be associated with rising economic inequality, something that could possibly be due to a concentration of wealth in some regions (like Paris for example).
+
+When looking at out other model with the other countries, we can see a different situation.
+The relationship between GDP per capita and Gini is not statistically significant.
+The model has a positive, but non-significant coefficient for GDP per capita, which can suggest that variations in this variable dosent have a significant impact on economic inequality.
+This is also shown by looking at the low R-squared of 0.004 \~ 0.4%, this may mean that GDP per capita only explains a rather small portion of the variance in inequality.
+
+As we can see, there is a large gap between France and the other countries when it comes to the relationship between economic development and economic inequality.
+
+## Exploring Alternative Functional Forms
+
+### Functional Form Exploration
+
+In addition to the linear model, we can also use other different models in order to explain the relationship between regional development and economic inequality.
+By using these additional models, we can test different transformations of the variables in order to find the best representation of the data.
+Some of the models we can use are:
 
@@
 
-### Functional Form Exploration
-
-```{r}
-#| echo: false
-#| output: false
+In our analysis, we have chosen to focus on two models: the logarithmic transformation model and the quadratic model.
+The choice of these models is rooted in both their methodological robustness and their ability to capture the complex nature of the relationship between regional development and income inequality.
+
+### Estimation and Visualization
+
+```{r}
+#| echo: false
+
+# Comparing the different models with the linear model
+
+# Linear model
+linear_modelFR <- lm((gini_n2 ~ GDP_per_capita), filter(gini_filtered, France == 1))
+linear_modelC <- lm((gini_n2 ~ GDP_per_capita), filter(gini_filtered, France == 0))
+
 # Logarithmic model
-log_model <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_filtered)
+log_modelFR <- lm(gini_n2 ~ log(GDP_per_capita), filter(gini_filtered, France == 1))
+log_modelC <- lm(gini_n2 ~ log(GDP_per_capita), filter(gini_filtered, France == 0))
 
 # Quadratic model
-quad_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2), data = gini_filtered)
-
-# Cubic model
-cubic_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2) + I(GDP_per_capita^3), data = gini_filtered)
-
-# Check the summary of each model
-summary(log_model)
-```
-
-```{r}
-summary(cubic_model)
-```
-
-```{r}
-summary(quad_model)
-```
-
-### Estimation and Visualization
-
-```{r}
-#| echo: false
-#| output: false
-# Quadratic model
-
-# Create a new data frame for plotting predicted values
-plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
-plot_data$gini_pred <- predict(quad_model, newdata = plot_data)
-
-# Plotting
-ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
-  geom_point(alpha = 0.5) +  # Plot the actual data points
-  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
-  labs(title = "Quadratic Model of GDP per Capita and Gini Coefficient",
-       x = "GDP per Capita",
-       y = "Gini Coefficient") +
+quadratic_modelFR <- lm(gini_n2 ~ log(GDP_per_capita) + I(GDP_per_capita^2), filter(gini_filtered, France == 1))
+quadratic_modelC <- lm(gini_n2 ~ log(GDP_per_capita) + I(GDP_per_capita^2), filter(gini_filtered, France == 0))
+
+huxreg(list("LinearFR" = linear_modelFR, "LinearC" = linear_modelC, "LogFR" = log_modelFR, "LogC" = log_modelC, "QuadraticFR" = quadratic_modelFR, "QuadraticC" = quadratic_modelC), statistics = c(N = "nobs", R2 = "r.squared", FStat = "statistic", PValue = "p.value"), note = "Note: {stars} T statistics in brackets.")
+```
+
+We have already seen in the linear regression model earlier that there is a significant positive correlation between GDP per capita and Gini for France, while its non-significant for the rest of the countries.
+In the logarithmic model, France also show that there is a positive relationship, with an R-squared of 0.490.
+The other countries still dosent have a significant relationship, and have a R-squared that is zero.
+
+For the quadratic model, France have a positive coefficient for squared term of GDP per capita, something that can suggest a U-shaped relationship between the variables.
+The R-squared for France is 0.682, higher than the R-squared for the other models.
+The other countries also indicates a U-shaped relationship, and has a R-squared of 0.061; this means that the quadratic model is a better fit than the other models to explain the variance in income inequality.
+
+#### Logarithmic Model
+
+```{r}
+#| echo: false
+
+# Visaulization of Logarithmic model for France
+
+france_data <- gini_filtered[gini_filtered$id_nuts0 == "FR", ]
+
+france_data$predicted_gini_FR = predict(log_modelFR, newdata = france_data)
+
+ggplot(france_data, aes(x = log(GDP_per_capita), y = gini_n2)) + 
+  geom_point(aes(color = "Actual"), alpha = 0.5) + 
+  geom_line(aes(y = predicted_gini_FR, color = "Predicted")) +
+  labs(title = "Logarithmic Model for France: Predicted vs Actual Gini Coefficients",
+       x = "Log of GDP per Capita",
+       y = "Gini Coefficient",
+       color = "Legend") +
   theme_minimal()
@@
 #| echo: false
-#| output: false
-# Logarithmic model
-
-# Create a new data frame for plotting predicted values
-plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
-plot_data$gini_pred <- predict(log_model, newdata = plot_data)
-
-# Plotting
-ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
-  geom_point(alpha = 0.5) +  # Plot the actual data points
-  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'red') +  # Add the regression line
-  scale_x_log10() +  # Logarithmic scale for the x-axis
-  labs(title = "Logarithmic Model of GDP per Capita and Gini Coefficient",
-       x = "GDP per Capita (log scale)",
-       y = "Gini Coefficient") +
+
+# Visaulization of Logarithmic model for the other countries
+
+other_countries_data <- gini_filtered[gini_filtered$id_nuts0 != "FR", ]
+
+other_countries_data$predicted_gini_C <- predict(log_modelC, newdata = other_countries_data)
+
+ggplot(other_countries_data, aes(x = log(GDP_per_capita), y = gini_n2)) + 
+  geom_point(aes(color = "Actual"), alpha = 0.5) + 
+  geom_line(aes(y = predicted_gini_C, color = "Predicted")) +
+  labs(title = "Logarithmic Model for Other Countries: Predicted vs Actual Gini Coefficients",
+       x = "Log of GDP per Capita",
+       y = "Gini Coefficient",
+       color = "Legend") +
   theme_minimal()
@@
 
-```{r}
-#| echo: false
-#| output: false
-# Create a new data frame for plotting predicted values
-plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
-plot_data$gini_pred <- predict(cubic_model, newdata = plot_data)
-
-# Plotting
-ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
-  geom_point(alpha = 0.5) +  # Plot the actual data points
-  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
-  labs(title = "Cubic Model of GDP per Capita and Gini Coefficient",
-       x = "GDP per Capita",
-       y = "Gini Coefficient") +
+In the plot for France, we can see a upward trend, which indicated that as the log of GDP per capita increases, the gini coefficient also increases, something that suggests higher income inequality.
+The data points are scattered around the predicted regression line, which goes upward.
+This pattern also supports the regression model that shows a significant positive realtionship between the variables.
+There are some extreme outliers, suggestion that some regions have a higher amount of economic activity (like Paris that is a economic center in France).
+This also shows that economic growth may lead to increased inequality because of the concentration of wealth in some regions.
+
+The model for the other countries show that there is not a significant relationship between the variables.
+The data points also dosen't cluster around the predicted line, which shows that the log of GDP does not explain the variance in the Gini coefficient.
+
+When analyzing both plots, it's clear that the relationship between economic development and income inequality behaves differently across these regions.
+The plot for France suggests that economic growth is associated with rising inequality, while the plot for the other countries indicates that economic growth has little to no effect on inequality
+
+#### Quadratic Model
+
+```{r}
+#| echo: false
+
+# Quadratic model for France
+
+france_data$predicted_gini_FR = predict(quadratic_modelFR, newdata = france_data)
+
+ggplot(france_data, aes(x = log(GDP_per_capita), y = gini_n2)) + 
+  geom_point(aes(color = "Actual"), alpha = 0.5) + 
+  geom_line(aes(y = predicted_gini_FR, color = "Predicted")) +
+  labs(title = "Quadratic Model for France: Predicted vs Actual Gini Coefficients",
+       x = "Log of GDP per Capita",
+       y = "Gini Coefficient",
+       color = "Legend") +
   theme_minimal()
@@
 #| echo: false
-#| output: false
-# Calculate residuals
-residuals_quad <- gini_filtered$gini_n2 - predict(quad_model, gini_filtered)
-residuals_cubic <- gini_filtered$gini_n2 - predict(cubic_model, gini_filtered)
-```
-
-```{r}
-#| echo: false
-#| output: false
-# Create a dataframe for plotting residuals
-residuals_data <- data.frame(GDP_per_capita = gini_filtered$GDP_per_capita, 
-                             Residuals_Quad = residuals_quad, 
-                             Residuals_Cubic = residuals_cubic)
-
-# Plotting residuals
-ggplot(residuals_data) +
-  geom_line(aes(x = GDP_per_capita, y = Residuals_Quad, color = "Quadratic")) +
-  geom_line(aes(x = GDP_per_capita, y = Residuals_Cubic, color = "Cubic")) +
-  labs(title = "Comparison of Residuals",
-       x = "GDP per Capita",
-       y = "Residuals") +
-  theme_minimal() +
-  scale_color_manual(values = c("Quadratic" = "blue", "Cubic" = "red"))
-```
-
-```{r}
-#| echo: false
-#| output: false
-sd(residuals_quad)
-sd(residuals_cubic)
-mean(residuals_quad^2)
-mean(residuals_cubic^2)
-```
-
-**Results Interpretation**
-
-## Part 3C Hetroskedasticity Testing and causality Discussion
-
-**Causality vs causation**
-
-While correlation indicates a relationship between two variables, it doesn't imply causation.
-In econometrics, it's a common pitfall to interpret correlation as causality.
-Panel data helps address this by allowing researchers to control for time-invariant factors and observe changes over time, which can provide stronger evidence for causal relationships.
-
-Panel data analysis helps address endogeneity issues, such as omitted variable bias, which can lead to spurious correlation.
-By tracking the same units over time, it's possible to differentiate whether a relationship is genuinely causal or merely a correlation coinciding in time.
-
-Overall, panel data analysis is a vital tool in econometrics for distinguishing between causality and correlation.
-By allowing for control over unobserved heterogeneity and providing a dynamic view of data over time, it offers a more nuanced and accurate understanding of the relationships between variables.
-This is crucial in econometric research, where the distinction between causality and mere correlation may have significant implications.
-
-**Heteroskedasticity Testing**
-
-```{r}
-#| echo: false
-#| output: false
-#hetro
+
+# Quadratic model for the other countries
+
+other_countries_data$predicted_gini_C <- predict(quadratic_modelC, newdata = other_countries_data)
+
+ggplot(other_countries_data, aes(x = log(GDP_per_capita), y = gini_n2)) + 
+  geom_point(aes(color = "Actual"), alpha = 0.5) + 
+  geom_line(aes(y = predicted_gini_C, color = "Predicted")) +
+  labs(title = "Quadratic Model for Other Countries: Predicted vs Actual Gini Coefficients",
+       x = "Log of GDP per Capita",
+       y = "Gini Coefficient",
+       color = "Legend") +
+  theme_minimal()
+```
+
+For France, the Gini coefficients are initially concentrated at lower levels of GDP per capita, presenting stability or a slight decrease in inequality, which then transitions into a sharp upward trend at higher GDP levels.
+This is indicative of an inverted U-shaped curve, a characteristic of the Kuznets curve hypothesis, suggesting that inequality escalates significantly with further economic growth [@ota2017].
+The quadratic model appears to fit the lower and middle range of the data well, yet it deviates from the actual data points at the higher end of GDP per capita, signaling that other factors may come into play as the economy grows.
+
+The quadratic model for the other countries shows a more subtle U-shaped relationship, with the initial part of the curve being relatively flat.
+This indicates that changes in GDP per capita have a limited effect on income inequality at first.
+However, a steep increase is observed at the higher end of GDP per capita, which may be attributed to outliers or specific country conditions not captured by the model.
+
+Incorporating these observations, both plots still signify a non-linear relationship between GDP per capita and income inequality, but with varying patterns and intensities between France and the other countries.
+The pronounced curve for France supports a stronger Kuznets curve effect, while for the other countries, the effect is less pronounced and the connection between economic growth and inequality appears to be weaker and more influenced by external or unmodeled factors.
+
+### Results Interpretation
+
+The linear model, while providing a significant fit for France, falls short for the other countries, suggesting that the economic inequality's relationship to GDP per capita may not be linear or might be influenced by factors not captured in a simple linear model.
+The logarithmic model's reduced explanatory power compared to the linear model for France suggests that the relationship between GDP and inequality may not be strictly proportional across all levels of GDP per capita.
+
+The quadratic model's U-shaped curve implies that as countries develop, inequality may first decrease and then increase, supporting the Kuznets curve hypothesis.
+
+In conclusion, while the linear model provides a baseline understanding of the relationship between GDP per capita and income inequality, the logarithmic and quadratic models offer more nuanced insights that can lead to a deeper understanding of the underlying economic mechanisms.
+These alternative models highlight the importance of considering non-linear dynamics when analyzing economic relationships.
+
+## Heteroskedasticity Testing and Causality Discussion
+
+### Heteroskedasticity Testing
+
+The Breusch-Pagan test is used to detect heteroskedasticity in a regression model.
+The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables, which would violate one of the key assumptions of ordinary least squares (OLS) regression.
+
+```{r}
+#| echo: false
+#| output: false
+
+# Breusch-Pagan test for the original linear model
+bptest_linearFR <- bptest(linear_modelFR)
+bptest_linearC <- bptest(linear_modelC)
+
+# Breusch-Pagan test for the logarithmic model
+bptest_logFR <- bptest(log_modelFR)
+bptest_logC <- bptest(log_modelC)
+
+# Breusch-Pagan test for the quadratic model
+bptest_quadFR <- bptest(quadratic_modelFR)
+bptest_quadC <- bptest(quadratic_modelC)
+
+# Output of test results
+print(bptest_linearFR)
+print(bptest_linearC)
+print(bptest_logFR)
+print(bptest_logC)
+print(bptest_quadFR)
+print(bptest_quadC)
+```
+
+The Breusch-Pagan test results suggest that the linear and logarithmic models for the other countries and the quadratic models for both France and the other countries exhibit heteroskedasticity.
+This could have implications for the standard errors and confidence intervals of the coefficients, potentially leading to incorrect inferences about the significance of predictors.
+The linear model for France does not appear to suffer from this issue.
+
+Given these findings, it may be necessary to consider heteroskedasticity-robust standard errors or to transform the data or model specification to address this issue.
+It is essential to correct for heteroskedasticity to ensure the reliability of the regression analysis.
+
+```{r}
+#| echo: false
+
+# Residuals vs fitted values for linear model, all countries combined (including France)
+
 lm_01 <- lm(gini_n2 ~ GDP_per_capita, data = gini_filtered)
-```
-
-```{r}
-#| echo: false
-#| output: false
-# Obtain fitted values and residuals from the model
+
+# Obtaining fitted values and residuals from the model
 fitted_values <- fitted(lm_01)
@@
 
-# Create the plot
 plot(fitted_values, residuals, 
@@
      pch = 20)
-abline(h = 0, col = "blue", lwd = 2) # Add a horizontal line at 0
+abline(h = 0, col = "blue", lwd = 2)
 ```
@@
 
-The sets of disconnected residuals further out could indicate that there are distinct subgroups within our data that are behaving differently.
-It might suggest that the relationship between the variables differs across these groups or that there are influential outliers or leverage points affecting the model differently in these areas.
-
-The Breusch-Pagan test is used to detect heteroskedasticity in a regression model.
-The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables, which would violate one of the key assumptions of ordinary least squares (OLS) regression.
-
-```{r}
-#| echo: false
-#| output: false
-# Breusch-Pagan test for the original linear model
-bptest_linear <- bptest(lm_01)
-
-# Breusch-Pagan test for the logarithmic model
-bptest_log <- bptest(log_model)
-
-# Breusch-Pagan test for the quadratic model
-bptest_quad <- bptest(quad_model)
-
-# Output the test results
-print(bptest_linear)
-print(bptest_log)
-print(bptest_quad)
-```
-
-```{r}
-#| echo: false
-#| output: false
-# Calculate robust standard errors for the original linear model
-coeftest_lm_01 <- coeftest(lm_01, vcov = vcovHC(lm_01, type = "HC"))
-
-# Calculate robust standard errors for the logarithmic model
-coeftest_log_model <- coeftest(log_model, vcov = vcovHC(log_model, type = "HC"))
-
-# Calculate robust standard errors for the quadratic model
-coeftest_quad_model <- coeftest(quad_model, vcov = vcovHC(quad_model, type = "HC"))
-
-# Print out the results
-print(coeftest_lm_01)
-print(coeftest_log_model)
-print(coeftest_quad_model)
-```
-
-## Part 3D Panel Estimates
+### Causality Discussion:
+
+The Breusch-Pagan tests have confirmed the presence of heteroskedasticity in most of our models, especially for the other countries.
+This suggests that the assumption of homoscedasticity is violated in these models.
+Heteroskedasticity and patterns in the residuals can lead to inefficient estimates and undermine the causal interpretation of the models.
+When the error variance changes across observations, the standard errors of the coefficient estimates may be biased, leading to incorrect conclusions about the significance of predictors.
+
+The best strategies to address these issues might be to use heteroskedasticity-robust standard errors (White's standard errors), apply transformations to the dependent variable, or using weighted least squares (WLS).
+These strategies can help mitigate issues and lead to more robust conclusions.
+
+## Panel Estimates
 
@@
 However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017].
-@lessmann2017 also notes that the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
-
-# Discussion
+@lessmann2017 also notes that the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information.
+
+### Panel Estimation Task:
+
+```{r}
+#| echo: false
+#| output: false
+
+# Panel estiamtion
+
+plm1 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("id_nuts0"), model = "within",effect = "twoway") |> 
+  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
+
+plm2 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("year"), model = "within",effect = "twoway") |> 
+  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
+
+plm3 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("id_nuts2"), model = "within",effect = "twoway") |> 
+  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
+
+plm4 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("year","id_nuts2"), model = "within",effect = "twoway") |> 
+  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
+
+huxreg(plm1, plm2, plm3, plm4, statistics = c("nobs","r.squared","adj.r.squared", "statistic", "p.value"), 
+       stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
+       note = "Note): {stars} T statistics in brackets.")
+```
+
+### Panel Estimation Analysis
+
+### Panel Estimation Discussion
 

### 2023-12-06 MSB104_Submission_Group1.qmd
More updates, Including numberes in the decription and putting the printouts in Appendix

--- 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
@@
 
-Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files.
-Selecting our countries of study and timeperiod.
-Selecting the required NUTS level and specified the data to be in million Euro.
-
-## GDP
-
-The GDP dataset provides insights into GDP at regional level using the NUTS classification system.
-It furnishes GDP values in both current prices and adjusted for inflation, with figures given in purchasing power standards (PPS) to account for price level differences between countries.
-The data is mostly structured by year and region @database.
-
-The GDP at market prices represents the final result of production activities of resident producer units within a region or nation.
-It is calculated as the sum of the gross value added across various institutional sectors or industries.
-Furthermore, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries) @database.
-This also balances out in the total economy production account.
-In terms of methodology, while national accounts compile GDP from the expenditure side, regional accounts don't adopt this approach due to the complexities of accurately mapping inter-regional flows of goods and services.
-
-The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @database.
-
-## Population
-
-Eurostat's records annual population data with NUTS classification.
-Our dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender.
-Displayed in a year-by-region format, with yearly interval updates @database.
-
-Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project.
-The project covers 37 countries and is the central repository for demographic and migration-related data @database.
-As well as, specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year @database.
-Additionally, data on marriages, divorces, and migration flows are recorded .
-
-For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @database.
-
-## Demographic (Nuts 2)
-
-Statistics on population change (demo_r_mlifexp) and on population structure was collected from Euro Stat [@database].
-Used to monitor demographic behavior within political, economic, social and cultural contexts at NUTS 2 level.
-Same policy and demographic classifications as previous population data set applies.
-
-## Transportation Infrastructure (Nuts 2)
-
-This data-set (tran_r_net) presents recent data on the inland transport network of the European Union (EU), EFTA and candidate countries presenting motorways, railways and inland waterways [@database].
-Furthermore used as a varable of of interest (MLR) in our section growth and inequality.
-Moreover, to excamine if evolution of the transport network is closely linked to the general development of the economy.
-
-The collected data are compiled by the competent national authorities.
-Reported annually be the National Statistical Office, the Civil Aviation Authorities (air transport), the Ministry of Transport (inland waterways, railway and road networks, victims), and the National Maritime Administration (maritime transport) [@database].
-
-## Education (Nuts 2)
-
-This data set (edat_lfse_04) contains population by educational attainment level, sex and NUTS 2 regions (%).
-Including data on the highest level of education successfully completed by the individuals of a given population.
-Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labor status of young people by years since completion of highest level of education.
-Calculated as annual averages of quarterly EU Labor Force Survey data (EU-LFS).
-
-## NUTS classification
-
-The Nomenclature of Territorial Units for Statistics (NUTS) offers a stratified system to segment the economic territory of the EU and UK to facilitate the consistent collection and harmonization of regional statistics across Europe.
-The NUTS regions range from NUTS 0 Country level to NUTS 3 small units such as municipalities level @eurostat2023a.
+## **Data collection**
+
+All dataset used in this paper is from Eurostat and based on the Nomenclature of Territorial Units for Statistics (NUTS) classification.
+The NUTS system provides a hierarchical classification of the EU's economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).
+
+The collected data for GDP per capita includes values in current prices and those adjusted for purchasing power parity (PPP), enabling comparisons by accounting for price level differences.
+GDP at market prices measures the total economic output within regions, adjusted by taxes, subsidies, and not delineated by specific sectors.
+
+Furthermore, population data includes annual figures on births, deaths, net migration, along with demographic details.
+Eurostat update these dataset annually.
+Furthermore, transportation infrastructure data encompasses the extent of the inland transport network, including motorways, railways, and waterways, within the EU, EFTA, and candidate countries.
+It is reported annually by various national authorities and used to explore the link between infrastructure and regional development.
+Educational attainment data reflects the highest level of education completed by individuals within regions, with additional metrics on employment and education status, particularly among young people.
+This data is derived from annual averages of the EU Labor Force Survey.
 
@@
 
-In this part we will report and interpret different types of essential descriptive statistics.
-Measuring regional income inequality is challenging due to heterogeneity of regions [@lessmann2017].
-The number of regions in our data set varies largely in size and population.
-Since the focus of this paper is purely growth and changes in inequities over time, the variations of size and population density becomes a minor issue because the country-level territorial heterogeneity is fixed [@lessmann2017].
-
-"Interest in income inequality has led to the development of several ways of measuring it.
-Two types of measures are of interest in this paper----static and dynamic.
-*Static measures* provide a snapshot (cross sectional) of these inequalities at a point of time whereas the dynamic *measures capture historical trends* (Paneldata)*.*" [@wooldridge2020]
-
-By using figures, we can visualize the GDP per capita, and look at how it varies among the different regions. In these figures, a line represent one NUTS 3 region.
+In this section, we explore a variety of descriptive statistics, crucial for understanding regional income inequality, a concept that presents challenges due to the heterogeneity of regions.
+Our dataset encompasses a diverse range of regions, each varying significantly in size and population.
+
+This paper focuses on two distinct types of measures: static and dynamic.
+Static measures offer a cross-sectional snapshot, providing a perspective on these inequalities at a specific point in time.
+In contrast, dynamic measures track historical trends, offering insights into how these disparities evolve over time, a concept well-articulated in [@wooldridge2020].
+
+Moreover, to effectively convey our findings, we utilize graphical representations, particularly focusing on GDP per capita across various regions.
+These visualizations not only aid in understanding the current state of regional disparities but also in tracing their progression over time.
 
@@
 
-Comparing the GRDP (gross regional domestic product) per capita of the region with the highest income to the region with the lowest income (minimum per capita GRDP) provides a measure of the range of these disparities.
+Comparing the GRP per capita of the region with the highest income to the region with the lowest income (minimum GDP per capita) provides a measure of the range of these disparities.
 If this measure is small (close to 1), then it would mean that the different regions have relatively equal incomes [@wooldridge2020].
-If this measure is large, then the interpretation is more problematic, as it does not tell us if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers.
+If this measure is large, then the interpretation is more problematic, as it does not tell if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers.
 Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to comprehend, and politically powerful measure of regional income inequality.
@@
 Unlike the mean, the median is not influenced by extreme values and, thus, can provide a clearer picture of the "typical" value in situations where the data distribution is not symmetrical [@wooldridge2020].
-
-### **Time series (Panel data)**
-
-### Cross sectional analysis
-
-### Simple linear regression
-
-### Multiple linear regression
 
@@
 
-After calculating the GDP per capita for all NUTS 3 regions within the chosen countries, we can see that there is a large spread between the figures for the various regions.
-We want to look at regional inequity; in order to do this in a valuable way, countries are divide.
-Furthermore, gaining important insights on regional differences and utilize.
-Later to discuss national policy on equity and sustainable economic development in regions.
+When exploring our data we will look at different NUTS levels aswell as countries both together and seperated.
 
@@
 
-```{r}
-#| echo: false
+HER NÅ VI BESKRIVE TRENDER
+
+```{r}
+#| echo: false
+#| output: false
 
@@
 
-# Print the summary statistics
+# Print the summary statistics #APPENDIX
 print(summary_stats_portugal)
@@
 
-By looking at figure for Portugal, we can see that the GDP per capita in Portugal's regions appears to be fairly consistent.
-There is however some regional variability.
-We can see signs of urbanisazation where the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas.
-Since Lisbon is the capital of Portugal, there is probably a higher concentration of industries, making it a economic center (which again makes the GDP per capita higher).
-
-To continue, we can see that the mean is a little higher that the median, something that might indicate that regions like Lisbon are pulling up the average.
-If we compare the standard derivation for Portugal with the other countries, we'll see that is fairly low in comparison.
-This might mean that there is not a lot of variability between the GDP per capita across different regions in Portugal.
-The gap between minimum and maximum is also low compared to other countries, something that'll also show us that the economic disparity in Portugal might not be as high as it is in other countries.
+Portugal's regional GDP per capita is characterized by relatively consistent values across the country.
+The descriptive statistics of these values show a mean of approximately 1.42 and a median of 1.39.
+This close alignment between the mean and median suggests a balanced distribution of economic value across regions.
+
+Furthermore, the standard deviation is 0.37, indicating a modest variation in GDP per capita across different regions.
+This is comparatively low when viewed against similar statistics from other countries, suggesting a more uniform economic landscape within Portugal.
+
+Examining the range, the minimum GDP per capita is 0.65, and the maximum is 2.72.
+The narrow gap between these extremes further supports the observation of limited economic disparity across the regions.
+While urban areas like Lisbon have higher GDP per capita, indicative of economic concentration, the overall disparity across the country remains relatively low.
 
@@
 #| echo: false
+#| output: false
 
@@
 
-When looking at the figure for France, we see some regions have a much higher GDP per capita compared to the other regions.
-The regions with the highest GDP per capita for all years is the île-de-France region, one that also includes Paris.
-This significant difference between the regions with the highest GDP per capita and the lowest, shows us that there is a high concentration of economic activity and wealth in a few urban regions.
-Similar to Portugal, we see differences between urban and rural regions.
-
-Just as in Portugal, there is also a higher mean in France as well.
-On the contary the data in France has higher standard derivation, and the difference between minimum and maximum is larger.
-This strengthens earlier figures showing, some regions having a high concentration of wealth.
+GDP per capita statistics from France show that the mean is 2.63, indicating an overall high economic status across regions.
+The median, at 2.44, is lower than the mean, suggesting a distribution skewed towards wealthier areas.
+The standard deviation is 1.05, highlighting considerable regional economic variability.
+The range is notably wide, with the minimum GDP per capita at 0.83 and the maximum at 11.58, underscoring stark contrasts.
+Regions like Île-de-France, encompassing Paris, and affluent areas like the southern Riviera, likely contribute to the high maximum value, illustrating a significant economic divide between urban centres and rural areas.
 
@@
 
-```{r}
-#| echo: false
+HVA SKJER HER!
+
+```{r}
+#| echo: false
+#| output: false
 
@@
 
-In Hungary, most of the regions have similar GDP per Capita. One region that sticks out by having a higher value, is the region of Budapest, the capital.
-
-This case also record the mean as higher vale than the median, high standard derivation, and a large gap between minimum and maximum.
+In Hungary, the GDP per capita data shows notable regional disparities.
+With a mean of 0.8598 and a median of 0.765, the higher mean suggests Budapest's significant economic influence.
+The standard deviation of 0.406 and the wide range from a minimum of 0.32 to a maximum of 3.13 further highlight Budapest's economic dominance compared to other regions.
 
@@
 #| echo: false
+#| print: false
 
@@
 
-Additionally, Slovakia, also have one region with much higher GDP per capita than the rest of the regions.
-This region is Bratislava, which is the biggest city and the capital, something that might point to this city being the economic centre of Slovakia as well.
-
-Slovakia follows the trend with higher mean than the median, and a large gap between minimum and maximum.
-In addition, the standard derivation is significantly high, meaning that there is some regions (or one region in this case) that is further away from the rest of the regions in terms of economic development.
+In Slovakia, the GDP per capita statistics reveals significant regional economic differences.
+The mean GDP per capita is 1.2502, higher than the median of 1.095, indicating that Bratislava, the capital and largest city, may substantially boost the national average due to its economic prominence.
+
+The large standard deviation of 0.8018 and the extensive range from a minimum of 0.30 to a maximum of 4.02 highlighting urbanization in areas such as Bratislava's distinct economic status.
+These figures indicate that while most regions have moderate economic levels, Bratislava stand out with a much higher GDP per capita, with its role as the economic centre of Slovakia.
 
@@
 
-Lastly, we see similar pattern in Denmark, with the capital Copenhagen being one of the regions with the highest GDP per capita.
-
-Whith mean higher than the median, showing that regions like Copenhagen possible dragging the mean up by population volume.
+Lastly, In Denmark, with its higher mean of 4.4192 compared to the median of 4.0000 suggests regions like Copenhagen, with its significant economic activity, are likely elevating the national average.
+The standard deviation of 1.3439 indicates moderate regional economic disparities, while the range from 2.76 to 8.42 reflects Copenhagen's economic prominence relative to other regions.
 
@@
 
-To begin the third assignment, we will explore if the effect of development and economic inequality is significantly different between different subsets in our dataset.
-For our assigned countries, we figured the best way to look at differences between them, is to examine the differences between our biggest country (France) compared to the rest of the countries.
-In order to compare the two different subsets, we will create a dummy variable.
-This dummy variable will help us with explaining qualitative values in the regression model.
+When exploring whether the effect of development and economic inequality is significantly different between different subsets.
+We look at the variantions between our biggest country (France) compared to the rest of the countries.
 
@@
 
-When looking at out other model with the other countries, we can see a different situation.
-The relationship between GDP per capita and Gini is not statistically significant.
-The model has a positive, but non-significant coefficient for GDP per capita, which can suggest that variations in this variable dosent have a significant impact on economic inequality.
+When comparing France to the rest of the countries, we observe a different situation.
+Showing the relationship between GDP per capita and Gini is not statistically significant.
+This model has a positive, but non-significant coefficient for GDP per capita, which can suggest that variations in this variable does not have a significant impact on economic inequality.
 This is also shown by looking at the low R-squared of 0.004 \~ 0.4%, this may mean that GDP per capita only explains a rather small portion of the variance in inequality.
 
-As we can see, there is a large gap between France and the other countries when it comes to the relationship between economic development and economic inequality.
+Overall we record major differences between France and the other countries when it comes to the relationship between economic development and economic inequality.
 
@@
 
-In addition to the linear model, we can also use other different models in order to explain the relationship between regional development and economic inequality.
+In addition to the linear model, other models such as logarithmic transformation, the quadratic term and the cubic are useful to explain the relationship between regional development and economic inequality .
 By using these additional models, we can test different transformations of the variables in order to find the best representation of the data.
-Some of the models we can use are:
+The addditional models are explained in more details below.
 
@@
 
-The logarithmic transformation for all the variables allows the coefficients to be interpreted as elasticities, which measure the percentage change in the dependent variable associated with a one percent change in the independent variable [@lessmann2017].
-The use of logarithmic transformations is a common practice in econometrics to address issues such as heteroscedasticity and nonlinearity in the data.
+The logarithmic transformation for all the variables allows the coefficients to be interpreted as elasticizes, which measure the percentage change in the dependent variable associated with a one percent change in the independent variable [@lessmann2017].
+The use of logarithmic transformations is a common practice in econometrics to address issues such as heteroscedasticity and non-linearity in the data.
 By using logarithmic transformations, we are able to estimate the relationship between regional inequality and development in a more robust and accurate way [@lessmann2017].
@@
 
-Furthermore the quadratic term in the regression model is used to investigate the relationship between regional inequality and development [@lessmann2017].
-The results suggest an inverted U-shaped relationship between income and inequality, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
+The quadratic term in the regression model is used to investigate the relationship between regional inequality and development.
+The results suggest an inverted U-shaped relationship between income and inequality, as indicated by a negative coefficient for the quadratic term [@lessmann2017].
+This implies that initially, as income rises, inequality increases, but after reaching a certain threshold, further increases in income lead to a decrease in inequality [@lessmann2017].
 
@@
 
-The cubic function in the regression model is used to investigate the relationship between regional inequality and development.
-@lessmann2017 suggest that regional inequalities increase again at very high levels of development, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].
-The inclusion of a cubic term in a regression model allows the analysis of more complex dynamics, such as increasing or decreasing marginal effects [@wooldridge2020].
-However, interpreting these models can be more challenging, as the effect of a one-unit change in the predictor variable on the dependent variable is no longer constant and depends on the level of the predictor variable.
-
-In our analysis, we have chosen to focus on two models: the logarithmic transformation model and the quadratic model.
-The choice of these models is rooted in both their methodological robustness and their ability to capture the complex nature of the relationship between regional development and income inequality.
+The cubic function allows for the exploration of an N-shaped relationship between two variables.
+As suggested by @lessmann2017, this can be particularly relevant when studying the link between regional inequality and development.
+In such a scenario, inequality might initially decrease with development, then increase, and eventually decrease again at very high levels of development.
+
+Including a cubic term in a regression model enables the analysis of intricate dynamics, such as changing marginal effects.
+This is crucial in understanding how the impact of a predictor variable on the outcome variable varies at different levels of that predictor.
 
@@
 
-We have already seen in the linear regression model earlier that there is a significant positive correlation between GDP per capita and Gini for France, while its non-significant for the rest of the countries.
-In the logarithmic model, France also show that there is a positive relationship, with an R-squared of 0.490.
-The other countries still dosent have a significant relationship, and have a R-squared that is zero.
-
-For the quadratic model, France have a positive coefficient for squared term of GDP per capita, something that can suggest a U-shaped relationship between the variables.
+As earlier predicted there is a significant positive correlation between GDP per capita and Gini for France, while its non-significant for the rest of the countries.
+In the logarithmic model, R-squared of 0.490 in the French data set demonstrate this.
+Showing that the model does not explain much of the variance in the dependent variable.
+Other variables not included in the model might have a significant impact.
+While the model may still be statistically significant, the low R-squared value suggests that its predictive power is limited.
+However, the other countries still does not show a significant relationship, with results showing R-squared at zero.
+Indicates that the independent variables provide very little information about the variation in the dependent variable.
+This suggests that the model is not a good predictor of the dependent variable.
+It might be necessary to re-evaluate the model, consider additional variables, or review the underlying assumptions.
+
+For the quadratic model, France show positive coefficient for squared term of GDP per capita, something that can suggest a U-shaped relationship between the variables.
 The R-squared for France is 0.682, higher than the R-squared for the other models.
-The other countries also indicates a U-shaped relationship, and has a R-squared of 0.061; this means that the quadratic model is a better fit than the other models to explain the variance in income inequality.
+This implies that the model explains a moderate amount of the variance in the dependent variable.
+It is a resanoble fit and such values are common in social sciences and economics where the behavior or phenomena being modeled are influenced by many unobserved factors.
+
+The other countries also indicates a U-shaped relationship, despite R-squared levels as low as 0.061 in other words only about 6.1% of the variation in the dependent variable is explained by the independent variable(s) in our model.
+This indicates a low level of explanatory power.
+Overall, the quadratic model is a better fit than the other models to explain the variance in income inequality.
 
@@
 
+Even with a logarithmic model, extreme outliers such as those spotted above, at the end of the line can indicate that the relationship captured by the model does not hold as well in these atypical circumstances.
+It suggests that the economic dynamics of these extremely wealthy regions differ significantly from the average.
+It's important to carefully assess how these outliers influence our model and consider additional or alternative approaches to capture the complex nature of the relationship between GDP per capita and the Gini coefficient accurately.
+
 ```{r}
@@
 
-In the plot for France, we can see a upward trend, which indicated that as the log of GDP per capita increases, the gini coefficient also increases, something that suggests higher income inequality.
+In the plot for France, we can see an upward trend, which indicated that as the log of GDP per capita increases, the gini coefficient also increases, something that suggests higher income inequality.
 The data points are scattered around the predicted regression line, which goes upward.
 This pattern also supports the regression model that shows a significant positive realtionship between the variables.
-There are some extreme outliers, suggestion that some regions have a higher amount of economic activity (like Paris that is a economic center in France).
+There are some extreme outliers, suggesting some regions have a higher amount of economic activity (like Paris that is a economic center in France).
 This also shows that economic growth may lead to increased inequality because of the concentration of wealth in some regions.
@@
 
-The Breusch-Pagan test is used to detect heteroskedasticity in a regression model.
-The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables, which would violate one of the key assumptions of ordinary least squares (OLS) regression.
+Heteroskedasticity is according to @wooldridge2020 an important factor to consider in econometrics and statistical modelling, especially in the context of regression analysis.
+Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables \[@wooldridge2020\].
+In simpler terms, the spread of the residuals varies at different points in the regression model.
+Moreover, heteroskedasticity can lead to inefficient estimators and unreliable hypothesis tests.
+Leading to the assumption of homoscedastic in MRL not being met.
+In other words the variance of the error terms is not constant across all levels if the independent variables when heteroskedasticity is present.
+
+By conducting a Breusch-Pagan we can detect heteroskedasticity in our regression model.
+The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables.
+A higher BP value indicates a stronger presence of heteroskedasticity.
+Degrees of Freedom refers to the number of independent variables in the model.
+P-value determines the significance of the test result.
+A low p-value (typically less than 0.05) suggests that the null hypothesis of homoskedasticity (constant variance of errors) can be rejected, indicating heteroskedasticity.
+If heteroskedasticity is detected, the standard errors of the regression coefficients may be biased, which can lead to incorrect inferences in hypothesis testing.
 
@@
 
-The Breusch-Pagan test results suggest that the linear and logarithmic models for the other countries and the quadratic models for both France and the other countries exhibit heteroskedasticity.
-This could have implications for the standard errors and confidence intervals of the coefficients, potentially leading to incorrect inferences about the significance of predictors.
-The linear model for France does not appear to suffer from this issue.
-
-Given these findings, it may be necessary to consider heteroskedasticity-robust standard errors or to transform the data or model specification to address this issue.
+In the case of the France, the BP value is 0.25123 with 1 degree of freedom, and a p-value of 0.6162, suggesting no significant evidence of heteroskedasticity, as the p-value is well above the typical threshold of 0.05.
+However, for the rest of the countries, the BP value is much higher at 83.806, with 1 degree of freedom, and the p-value is less than 2.2e-16, indicating a strong presence of heteroskedasticity.
+This is also the case for the log model for France and for the rest of the countries.
+The log model for France has a BP value of 22.971 and a p-value of 1.645e-06, while the log model for the rest shows a BP value of 60.814 and a p-value of 6.273e-15.
+Both indicate significant heteroskedasticity.
+The quadratic models exhibit similar trends.
+The quadratic model for France shows a BP value of 59.658 with 2 degrees of freedom and a p-value of 1.11e-13, which clearly indicates heteroskedasticity.
+Similarly, the quadratic model for the rest, with a BP value of 18.4, 2 degrees of freedom, and a p-value of 0.0001011, suggests notable heteroskedasticity.
+
+Overall, these results, particularly the low p-values in most models, suggest that the null hypothesis of homoskedasticity (constant variance of errors) can be rejected, revealing heteroskedasticity.
 It is essential to correct for heteroskedasticity to ensure the reliability of the regression analysis.
+There are several methods to deal with heteroskedasticity, depending on the nature of the data and the specific requirements of the analysis.
+Residuals vs.
+Fitted is another way of detecting heteroskedasticity.
+However, it provides a deeper understanding of the heteroskedasticity's nature and extent
+
+**Residuals vs. Fitted**
+
+The residuals versus fitted values plot is another diagnostic tool that helps in visually identifying issues in a regression model, aiding in improving model accuracy and reliability [@wooldridge2020].
+Even after confirming heteroskedasticity.
+According to @wooldridge2020 It provides a deeper understanding of the heteroskedasticity's nature and extent, which is crucial for effectively addressing it in our regression analysis.
+Furthermore, the interpretation of different patterns in the plot, like random dispersion indicating a good model fit or systematic patterns indicating issues.
+Subsiquently, ensuring robust and reliable modeling visual diagnostics, complement statistical tests \[@wooldridge2020\].
 
@@
 
-The residuals vs.
-fitted values plot indicates potential issues with our linear model, which might include heteroskedasticity and model misspecification.
+There appears to be a pattern in the residuals; they are not randomly dispersed around the horizontal line but show a curve-like trajectory as the fitted values increase.
+This pattern suggests non-random error variance, which is indicative of non-linearity in the relationship between the independent and dependent variables, or it could also suggest heteroskedasticity.
+Ideally, if the model was a perfect fit, the residuals would be randomly scattered around the zero line (the horizontal blue line in your plot) without any discernible pattern.
+This would indicate homoskedasticity and a good linear fit.
+The pattern we are seeing in our residuals plot suggests that our linear model may not be the best representation of the relationship between the variables.
+It could be that a non-linear model would provide a better fit or that there are other variables or transformations that should be considered.
+Understanding this plot is important because it directly relates to the assumptions underlying linear regression.
+
 The observed patterns and their possible implications, suggests that our model might be missing a nonlinear relationship between the dependent and independent variables.
-This is a sign of model misspecification, indicating that the current linear model may not be the best fit for our data.
+This is a sign of model misspecification, confirming that the current linear model may not be the best fit for our data.
 The variance in the residuals first increasing, then decreasing, and increasing again, implies that the error terms have non-constant variance, which is a sign of heteroskedasticity.
-In a well-specified homoscedastic model, the spread of the residuals should be roughly equal across all levels of fitted values.
 
@@
 
-The Breusch-Pagan tests have confirmed the presence of heteroskedasticity in most of our models, especially for the other countries.
-This suggests that the assumption of homoscedasticity is violated in these models.
-Heteroskedasticity and patterns in the residuals can lead to inefficient estimates and undermine the causal interpretation of the models.
-When the error variance changes across observations, the standard errors of the coefficient estimates may be biased, leading to incorrect conclusions about the significance of predictors.
-
-The best strategies to address these issues might be to use heteroskedasticity-robust standard errors (White's standard errors), apply transformations to the dependent variable, or using weighted least squares (WLS).
-These strategies can help mitigate issues and lead to more robust conclusions.
+The statistical relationship between two variables does not imply that one causes the other.
+Correlation can be due to a causative relationship or due to both variables being influenced by a third factor.
+Establishing causation means demonstrating that changes in one variable directly result in changes in another.
+To claim causality, econometricians typically need to demonstrate three things.
+Firstly, temporal Precedence meaning the cause must precede the effect in time.
+Secondly, covariation of the Cause and Effect.
+When the cause happens, the effect happens; when the cause doesn't happen, the effect doesn't happen.
+Lastly, no Plausible Alternative Explanations meaning there should be no other factors that can explain the relationship between the cause and effect.
 
@@
 # References
+
+# Appendix
+
+```{r summary-stats-appendix, echo=FALSE}
+# Descriptive statistics GDP per capita, mean, min, max, SD
+
+## Summary Statistics
+# Print the summary statistics for Portugal
+
+cat("Summary Statistics for Portugal's GDP Per Capita:\n")
+print(summary_stats_portugal)
+
+# Print the summary statistics for France
+cat("\nSummary Statistics for France's GDP Per Capita:\n")
+print(summary_stats_france)
+
+# Print the summary statistics for Hungary
+cat("\nSummary Statistics for Hungary's GDP Per Capita:\n")
+print(summary_stats_hungary)
+
+# Print the summary statistics for Slovakia
+cat("\nSummary Statistics for Slovakia's GDP Per Capita:\n")
+print(summary_stats_slovakia)
+
+# Print the summary statistics for Denmark
+cat("\nSummary Statistics for Slovakia's GDP Per Capita:\n")
+print(summary_stats_denmark)
+```

### 2023-12-06 MSB104_Submission_Group1.qmd
Bar chart interpetation :D

@@
 All dataset used in this paper is from Eurostat and based on the Nomenclature of Territorial Units for Statistics (NUTS) classification.
-The NUTS system provides a hierarchical classification of the EU's economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).
+The NUTS system provides a hierarchical classification of the EU\'s economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).
 
@@
 Furthermore, population data includes annual figures on births, deaths, net migration, along with demographic details.
-Eurostat update these dataset annually.
-Furthermore, transportation infrastructure data encompasses the extent of the inland transport network, including motorways, railways, and waterways, within the EU, EFTA, and candidate countries.
-It is reported annually by various national authorities and used to explore the link between infrastructure and regional development.
-Educational attainment data reflects the highest level of education completed by individuals within regions, with additional metrics on employment and education status, particularly among young people.
+Eurostat update these datasets annually.
+Transportation infrastructure data encompasses the extent of the inland transport network, including motorways, railways, and waterways, within the EU, EFTA, and candidate countries.
+The data is reported annually by various national authorities and used to explore the link between infrastructure and regional development.
+Educational attainment data reflects the highest level of education completed by individuals within regions, with additional metrics employment and education status, particularly among young people.
 This data is derived from annual averages of the EU Labor Force Survey.
+
+In the next section, we explore the data using descriptive statistics, crucial for understanding regional income inequality, a concept that presents challenges due to the heterogeneity of regions.
 
@@
 
+## Descriptive statistics
+
 In this section, we explore a variety of descriptive statistics, crucial for understanding regional income inequality, a concept that presents challenges due to the heterogeneity of regions.
@@
 
-This paper focuses on two distinct types of measures: static and dynamic.
+This paper focuses on two distinct types of statistic measures: static and dynamic.
 Static measures offer a cross-sectional snapshot, providing a perspective on these inequalities at a specific point in time.
-In contrast, dynamic measures track historical trends, offering insights into how these disparities evolve over time, a concept well-articulated in [@wooldridge2020].
+In contrast, dynamic measures track historical trends, offering insights into how these disparities evolve over time, a concept well-articulated in by @wooldridge2020.
 
@@
 
-## Descriptive statistics
-
 **Mean**
@@
 
-Comparing the GRP per capita of the region with the highest income to the region with the lowest income (minimum GDP per capita) provides a measure of the range of these disparities.
-If this measure is small (close to 1), then it would mean that the different regions have relatively equal incomes [@wooldridge2020].
-If this measure is large, then the interpretation is more problematic, as it does not tell if the high ratio is due to substantial variation in the distribution of per capita GDRPs or the presence of outliers.
-Nevertheless, maximum to minimum ratio (MMR) provides a quick, easy to comprehend, and politically powerful measure of regional income inequality.
+Comparing the GDP per capita of the region with the highest income to the region with the lowest income (minimum GDP per capita) provides a measure of the range of disparities.
+If this measure is small (close to 1), the different regions have relatively equal incomes [@wooldridge2020].
+If this measure is large, then the interpretation is more problematic, as it does not tell if the high ratio is due to substantial variation in the distribution or the presence of outliers.
+Nevertheless, maximum to minimum ratio (MMR) provides a quick, comprehensible, and powerful measure of regional income inequality.
 
@@
 
-Calculating SD to quantify the dispersion or variability of a data set around its mean.
-Helping us assess the degree of uncertainty, variability, or risk associated with an economic variable or parameter, which is crucial for understanding the reliability of estimations and predictions [@wooldridge2020].
+Calculating the standard deviation (SD) quantifies the dispersion or variability of a dataset around its mean, thereby aiding in assessing the degree of uncertainty, variability, or risk associated with an economic variable or parameter.
+This is crucial for evaluating the reliability of estimations and predictions in economic analysis, as highlighted by @wooldridge2020.
 
@@
 
-The formula for calculating GDP per Capita is as follows:
+The formula for calculating GDP per Capita below:
 
@@
 
-When exploring our data we will look at different NUTS levels aswell as countries both together and seperated.
+In the following sections we will investigate our data at various NUTS levels and consider countries both collectively and individually.
+This approach allows us to discern broader regional trends, while also capturing unique national characteristics.
+Such a multifaceted examination is essential for a nuanced understanding of the economic dynamics within our chosen countries.
 
@@
 
-HER NÅ VI BESKRIVE TRENDER
+HER kommer BESKRIVE TRENDER
 
@@
 The range is notably wide, with the minimum GDP per capita at 0.83 and the maximum at 11.58, underscoring stark contrasts.
-Regions like Île-de-France, encompassing Paris, and affluent areas like the southern Riviera, likely contribute to the high maximum value, illustrating a significant economic divide between urban centres and rural areas.
+As observed in our time series, regions like Île-de-France, encompassing Paris, and affluent areas like the southern Riviera, likely contribute to the high maximum value, illustrating a significant economic divide between urban centres and rural areas.
 
@@
 
+Beskrivelse kommer! :D
+
 ```{r}
@@
 
-Lastly, In Denmark, with its higher mean of 4.4192 compared to the median of 4.0000 suggests regions like Copenhagen, with its significant economic activity, are likely elevating the national average.
+Lastly, in Denmark we observe a higher mean of 4.4192 compared to the median of 4.0000 suggesting regions like Copenhagen, with its significant economic activity, are likely elevating the national average.
 The standard deviation of 1.3439 indicates moderate regional economic disparities, while the range from 2.76 to 8.42 reflects Copenhagen's economic prominence relative to other regions.
@@
 
+In this part we compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.
+The GINI coefficient measures inequality in our distribution, a useful tool when looking at regional inequality.
+The closer the GINI coefficient is to 1, the bigger the inequality; a number closer to 0 equals equality.
+Furthermore, looking at the GINI coefficient for NUTS 2 regions, will provide a better overview over differences in income between different regions, making it easier to address difference between the regions @hasell2023.
+
 ### Gini Coefficient Calculation
 
-In this part we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.
-
-The gini coefficient can help us measure inequality in a distribution, as is therefore a useful tool for us to use when we look at regional inequity.
-The closer the gini coefficient is to 1, the bigger the inequality is; a number closer to 0 equals equality.
-When looking at the gini coefficient for NUTS 2 regions, we also get a better overview over differences in income between different regions, and it also makes it easier to find the reasons as to why there is a difference between the regions @hasell2023.
-
-With the use of the NUTS3 GDP per capita data and this formula:
+Calculation the NUTS3 GDP per capita data with the following formula:
 
 $GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|$
+
+The GINI serves as the dependent variable Y, reflecting the extent of income inequality within a specific region.
+Furthermore, the weights Pi / pi and Pj / pj adjust the calculation based on the proportion of individual's income, to the total population.
+The equation considers the population size of each area to give more accurate importance to areas with larger populations.
+This way, income disparities in more populated areas have a proportionally greater impact on the Gini coefficient than those in less populated areas, reflecting a more realistic picture of regional inequality.
 
@@
 
-After calculating the gini coefficients, we can see that there are some similarities to the data we got from GDP per capita for NUTS 3 regions.
-In order to see these similarities better, as well as look for other important aspects that can be provided trough the calculations, we will visualize the data in three different ways.
+After calculating the GINI, we observe similarities to the GDP per capita data for NUTS 3 regions.
+By adopting different approaches to visualization such as the one provided in the bar chart below, we get a better understanding of these similarities.
+Moreover, looking for other important aspects.
 
@@
 ```
+
+The chart underscores the varied landscape of income inequality within European NUTS 2 regions.
+France's Île-de-France (FR10), Champagne-Ardenne (FRF2), Midi-Pyrénées (FRJ2), and Auvergne (FRK2) show significant disparities, likely linked to urban wealth concentration and policy.
+Denmark differs, with Copenhagen (DK01) and Central Jutland (DK03) displaying lower inequality, and Zealand (DK02) and North Jutland (DK04) even lower, reflecting strong equal opportunities practices.
+Lorraine (FRF3) in France suggests similar equity.
+Hungary's inequality is uniformly high except for Southern Transdanubia (HU23).
+Portugal's Lisbon (PT11) stands out for high inequality, while Slovakia's regions show lower disparity, akin to Denmark's average but not its lowest.
+Denmark emerges as having the least inequality, with France showing the greatest and Hungary and Slovakia presenting more uniform distributions.
 

### 2023-12-06 MSB104_Submission_Group1.qmd
Overskrifter er iallefall på plass nå ;-)

@@
 The sructure of our study is methodically designed to facilitate this exploration.
-In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***).
-In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).
-Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).
+In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data.
+In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010.
+Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality.
 
@@
 
-Why some places grow and prosper is a fundamental question in social science research.
-This question motivated Adam Smith's work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A.
-(1863).
-An inquiry into the nature and causes of the wealth of nations.
-Edinburgh: A.
-& C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*.
+Why some places grow and prosper is a fundamental question in the field of social science.
+This question motivated Adam Smith's \[1863\] work and has been a major influence of the *new economic geography* [@feldman2014]*.
 Furthermore,* regional growth and inequality have been important topics of research in Europe for many years.
-In recent decades, the European Union implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
-Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
+Despite the European commission\'s efforts to decrease income gaps, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
 
@@
 
-Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
-Lessmann and Seidel's research offers insights into these dynamics, highlighting the persistence of income disparities, across regions despite overarching economic advancement [@lessmann2017].
-Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014].
+Regional growth in Europe reveals a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
+Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by \[@gennaioli2014\] findings.
 
@@
 
-The 2008 financial crisis significantly disrupted regional growth and exacerbated income inequalities across Europe.
+The 2008 financial crisis disrupted regional growth and worsened income inequalities across Europe.
 Nguyen et al.
-highlight that financial turmoil typically leads to increased income disparity; a trend observed globally [@nguyen2022].
-
+(2022) highlight that financial instability typically leads to increased income disparity; a trend observed globally.
 Focusing on Europe, we expect varied regional economic performances, especially when comparing Eastern European countries such as Hungary and Slovakia to their Western counterparts.
-Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential growth acceleration.
-In particular for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014].
-However, the financial crisis and subsequent labour market shifts may also have influenced these trends.
-
-EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic.
-In particular Portugal as the crisis hit them harder.
+Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al (2014), we anticipate a potential growth acceleration after submission.
+ EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis.
+This pattern possibly mirrored in the recent COVID-19 pandemic.
 These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.
@@
 
-Another affect that surprised us was that the economic crisis led to increase inequality.
-The study "*the effects of financial crisis on regional inequality*" indicate any type of financial crisis results in higher income inequality [@nguyen2022].
-
-@lessmann2017 regional inequality study address the importance of studying these aspects.
-Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
-The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.
+The effect that surprised us most was that the economic crisis led to increase inequality\[@nguyen2022\].
+indicate that any type of financial crisis results in higher income inequality.
+Furthermore,  @lessmann2017 address the importance of studying these aspects.
+Moreover, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
 
@@
 
-Geographic characteristics stand out as pivotal, with the natural and topographical elements of a region significantly swaying income distributions [@lessmann2017].
-
-Urbanization emerges as a crucial determinant, where bustling, populous around urban centers usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
-
-*Transport infrastructure and Education*
+Geographic characteristics stand out as fundamental, with the natural and topographical elements of a region shifting income distributions [@lessmann2017].
+Urbanization emerges as a crucial determinant, where active, populations around urban centres usually enjoy raised income levels, contrasting with the disparities witnessed in more rural and secluded areas [@lessmann2017].
 
@@
 The study mentions that human capital is the most important determinant of differences in regional development within countries.
-The study also notes that human capital creates positive externalizes, which can spill over regional boundaries.
-Moreover, the quality of human capital within countries, as measured by the secondary-school enrollment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
+Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
 Lastly, promoting internal migration.
@@
 
-Furthermore, @lessmann2017 provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
+@lessmann2017 provides insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
 The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
-
-@lessmann2017 paper also finds that there is an N-shaped relationship between regional inequality and economic development, which means that regional inequality increases again after the inverted U-shaped pattern has been completed [@lessmann2017].
-
-@iammarino2019 discusses regional inequalities in European countries.
-It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium.
-The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019].
-Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
+Additionally, there is an N-shaped relationship between regional inequality and economic development according to \@lessmann2017.
+Subsequently, regional inequality increases again after the inverted U-shaped pattern has been completed [@lessmann2017].
+
+Lastly, [@iammarino2019] argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
 On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.
 
-Overall, it is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure and education emerging as significant determinants.
-Earlier studies provide us with a fundamental understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities.
-The integration of transport infrastructure and education into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
-
-# Data and Descriptive
+Overall, it is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by several factors, with transport infrastructure and education emerging as significant determinants.
+Earlier studies provide us with a fundamental understanding of these dynamics, while also highlighting the impacts of economic crises on regional disparities.
+The integration of transport infrastructure and education adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
+
+# Data and Methodology
 
@@
 #| echo: false
-#| print: false
+#| output: false
 
@@
 ggplot(gini_filtered, aes(x = id_nuts0, y = GDP_per_capita)) +
+  geom_boxplot() +
+  ggtitle("Plot 1: Distribution of GDP per Capita by Country") +
+  xlab("Country") +
+  ylab("GDP per Capita")
+
+# Saving the plot for referencing in the appendix
+g <- ggplot(gini_filtered, aes(x = id_nuts0, y = GDP_per_capita)) +
   geom_boxplot() +
@@
   ylab("GDP per Capita")
-```
-
-# Cross sectional estimates
+
+# Save the plot as an image file
+ggsave("GDP_per_Capita_Boxplot.png", plot = g, width = 10, height = 6)
+```
+
+The box plots (in the appendix) provide a distinct perspective on the data, independent of the assumptions typically associated with other statistical distributions.
+This method of visualization highlights Denmark as having the highest overall GDP per capita across its regions.
+France, while not having the overall highest GDP per capita, demonstrates significant variability, especially when considering its outliers.
+Hungary, in contrast, is still presenting the lowest GDP per capita.
+These insights, derived from the box plot analysis, augment our understanding from previous descriptive statistics, bar plots, and time series analyses, offering a more nuanced view of the GDP per capita distribution among these countries.
+
+# Empirical Findings
+
+## Cross sectional estimates
 
 ## Part 2A: Growth and Inequity
-
-### **Data acquisition**
-
-Firstly, the data is summerised for the year 2010 at the NUTS2 regional level, focusing on economic metrics.
-Processing the dataset containing the NUTS3 level.
-Thereafter, computing the total GDP and total population for each NUTS2 region and year.
-Then, derive the GDP per capita at the NUTS2 level by dividing the total GDP by the total population again.
-Furthermore, narrowing down the dataset to observations from the year 2010 where the Gini coefficient is positive.
-Then adding, logarithmic transformations to linearize relationships or to reduce the impact of extreme values.
-
-The next step is making a data frame of NUTS2 2010 and grouping by NUTS0 country level variables, as well as allocating NUTS2 to selected countries.
-Lastly presenting data in **table XXX** showing a count of NUTS2 regions within each top-level region or country in the 2010 data.
-
-The next step is, estimating a Linear Regression for All Countries using cross sectional data from 2010, NUTS2.
 
@@
 
-The goodness of fit in a simple linear regression model measures how well the regression line approximates the real data points.
-The regression line that best represents the data according to the least squares criterion, which minimizes the sum of the squared vertical distances of the points from the line.
-
-R\^2 is a statistical measure that represents the proportion of the variance for the dependent variable that's explained by the independent variable.
-It ranges from 0 to 1.
-A higher R\^2 value indicates a better fit of the model to the data.
-
-R\^2 = 0 The model does not explain any of the variability of the response data around its mean.
-
-R\^2 = 1 The model explains all the variability of the response data around its mean.
-
-If the assumptions for a simple linear regression are met, it indicates a good fit for the model.
-[Plotting the data can provide a visual indication of the goodness of fit.]{.underline} The points should fall around a straight line without clear patterns in the residuals.
-As it is crucial to verify these assumptions before proceeding with interpreting the results of the regression analysis.
+The goodness of fit in simple linear regression, evaluated by R\^2, gauges how well the regression line fits the data.
+This line is determined using the least squares criterion, minimizing squared differences between data points and the line.
+R\^2 varies from 0 to 1, where 0 indicates no explanatory power and 1 signifies perfect fit.
+A higher R\^2 suggests a better model fit.
+For an effective model, data should linearly align with minimal patterns in residuals, assuming all linear regression prerequisites are met.
 
 **Residuals vs.
-Fitted Values Plot** to check for homoscedasticity and linearity.
+Fitted Values Plot** is used to check for homoscedasticity and linearity.
 Ideally, this plot shows no pattern; the residuals are randomly scattered around the horizontal line at zero.
@@
 #| echo: false
-#| fig-cap: Regression statistics of all countries for the year 2010
-
+#| output: false
+
+# Made a copy for appendix
 # Finding the natural logarithm of GDP per capita
@@
 #| echo: false
+#| output: false
 #| fig-cap: Regression statistics of all countries seperately for the year 2010
 
+#Appendix 
 # Estimation for each country seperately 
@@
 
-"The OLS method is employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
-This technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
-The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the data-set, without necessitating iterative procedures.
-
-Furthermore, when the classical linear regression assumptions are met, OLS guarantees that the estimators are BLUE, ensuring their unbiasedness and efficiency.
-This is particularly crucial in econometric analysis, where the precision and reliability of parameter estimates are paramount for policy implications and economic interpretations.
-The foundation assumptions of linear regression, including linearity, independence, homoscedasticity, and normality of residuals, are prerequisites for OLS to attain these desirable properties.
-It is imperative, therefore, to conduct diagnostic tests and assess the validity of these assumptions to ensure the robustness of the OLS estimators used in our analysis."
-
-### **Assumptions**
-
-**I. Linearity**
-
-The relationship between the independent variable X and the dependent variable Y is linear.
-This means that changes in X are associated with proportional changes in Y.
-A straight line should provide a good fit to the data points when plotted on a graph.
-
-You can create a scatter plot of Y versus X and visually inspect whether a straight line could well represent the relationship.
-Alternatively, you can plot the residuals versus the fitted values and check for any obvious patterns or non-linearity.
-
-**II. Independence:**
-
-Observations are independent of each other. The value of Y for one observation should not depend on the value of Y for any other observation.
-
-This assumption is more about study design and data collection.
-Ensure that your observations are not correlated in time or space.
-For example, if you're analyzing economic data across regions, make sure that the regions are not influencing each other.
-
-**III. Homoscedasticity (Equal Variance):**
-
-The variance of the residuals (the errors) is constant across all levels of X.
-This means that the spread or "width" of the residuals should remain roughly the same across all values of the independent variable.
-
-A plot of residuals versus fitted values should show a random scatter and not display any funnel-like shapes (wider at one end).
-
-**IV. Normality of Residuals:**
-
-The residuals (the differences between observed and predicted values) are normally distributed. This assumption is particularly important for hypothesis testing and creating confidence intervals.
-
-A Quantile-Quantile (Q-Q) plot of the residuals can show if they follow a normal distribution. The points should fall roughly along a straight line.
-
-**V. No Perfect Multicollinearity (Specific to Multiple Regression)**
-
-In multiple regression settings, this assumes that no independent variable is a perfect linear function of any other independent variables.
-While this is more pertinent to multiple linear regression, it is crucial there because high correlation between independent variables can lead to unstable coefficient estimates.
-
-Checking the variance inflation factor (VIF) for each variable; a VIF above 5-10 indicates a problematic amount of collinearity.
-
-**Visualization**
-
-```{r}
-#| echo: false
+The OLS method is according to @wooldridge2020 employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
+Furthermore, this technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
+The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the dataset, without necessitating iterative procedures.
+According to @wooldridge2020 the following assumtions need to be met to achieve BLUE.
+
+1\. ***Linear in coefficients and error term***: The model is linear in its terms and the error term.
+
+2\. ***Error term has zero mean***: The error term averages zero, ensuring unbiasedness.
+
+3\. ***No correlation between independent variables and error term***: Independent variables should not predict the error term.
+
+4\. ***Observations of error term are uncorrelated*****:** Each error term observation doesn't predict the next.
+
+5\. ***Homoscedasticity*****:** The error term's variance is consistent across observations.
+
+6\. ***No perfect multicollinearity***: No independent variable is a perfect linear function of others.
+
+7\. ***Normally distributed error term (Optional*****)**: This aids in statistical testing, although not required for unbiased OLS estimates
+
+```{r}
+#| echo: false
+#| output: false
 
 # Making two plots, first one is a residuals vs fitted plot, while the other is a normal Q-Q residuals plot
-plot(lm, which = 1:2) 
-```
-
-We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient, and that together with the regression statistics can help with discussing if the classical OLS assumptions hold for the model.
-The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model.
-The residuals should be randomly scattered around the horisontal 0 line, something that can inducate that the variances of the error terms are constant.
+#plot(lm, which = 1:2) 
+```
+
+The residuals vs fitted values and the Q-plot, combined with the regression statistics may help us find out if classical OLS assumptions are met.
+The first help us check for homoscedasticity assumption of a linear regression model.
+The residuals should be randomly scattered around the horizontal 0 line, something that can indicate that the variances of the error terms are constant.
 In our plot, the residuals are in some extent randomly distributed, and there is also no clear pattern; this suggest that there is likely no significant issues with heteroscedasticity or non-linearity.
-
-The other plot - normal Q-Q plot - is used to assess if the residuals of the linear model are normally distributed.
+The latter is used to assess if the residuals of the linear model are normally distributed.
 In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed.
 There are however some outliers in the tails, that suggest some variation from normality.
-
-In both of these plots, we can see that there are some outliers, these can affect the fit of the model.
-These outliers might come from regions that have a high GDP per capita compared to rest of the regions, like Paris in France that is a financial hub.
-
-```{r}
-#| echo: false
+Both plots are presented in appendix.
+
+```{r}
+#| echo: false
+#| output: false
 #| label: fig_scatterplot-all
@@
 
-## Part 2B: Exploring Other Determinants of Inequity
-
-### I. Data Acquisition
+### Part 2B: Exploring Other Determinants of Inequity
+
+#### I. Data Acquisition
 
@@
 
-### II. Multiple Linear Regression Model
+#### II. Multiple Linear Regression Model
 
@@
 #| echo: false
+#| output: false
 #| fig-cap: Multiple Linear Regression Model
@@
 
-### Model specification
+#### II. Model specification
 
@@
 
-### III. Model Interpretation
+#### III. Model Interpretation
 
@@
 
-# Assignment 3: Alternative Functional Forms and Panel Estimates
-
-## Testing Development Effects Across Subsets
-
-### Subset Analysis
+## Alternative Functional Forms and Panel Estimates
+
+### Testing Development Effects Across Subsets
+
+#### Subset Analysis
 
@@
 
-### Subset Analysis Discussion
+#### Subset Analysis Discussion
 
@@
 
-### Results Interpretation
+#### Results Interpretation
 
@@
 
-## Heteroskedasticity Testing and Causality Discussion
-
-### Heteroskedasticity Testing
+### Heteroskedasticity Testing and Causality Discussion
+
+#### Heteroskedasticity Testing
 
@@
 
-# References
+# Discussion
+
+# Limitations and Future Research
+
+# Conclusion
 
 # Appendix
+
+1.  **Descriptive statistics**
 
@@
 ```
+
+2.  **Boxplot**
+
+![](GDP_pc_Boxplot.png)
+
+3.  **Regression statistics of all countries separately for the year 2010**
+
+```{r}
+#| echo: false
+#| fig-cap: Regression statistics of all countries for the year 2010
+
+cat("Intercept (Coefficient 1):", b0_earn, "\n")
+cat("Slope of log GDP per capita (Coefficient 2):", b1_earn, "\n")
+```
+
+![](images/Regression_all_countries2010-01.png)
+
+**Testing for OLS assumtions**
+
+```{r}
+#| echo: false
+#|
+
+# Making two plots, first one is a residuals vs fitted plot, while the other is a normal Q-Q residuals plot
+plot(lm, which = 1:2) 
+```
+
+# References

### 2023-12-06 MSB104_Submission_Group1.qmd
:D

@@
 
-While national GDP and GDP per capita are vital indicators of a country's aggregate economic health, they do not shed light on how wealth or income is distributed among its residents.
-A high national GDP can, paradoxically, coexist with pockets of regional deprivation [@lessmann2017].
-
+While national GDP and GDP per capita are vital indicators of a country's collective economic health, they do not shed light on how wealth or income is distributed among its residents.
+A high national GDP can, unexpectedly, coexist with pockets of regional deprivation [@lessmann2017].
 The truth of this statement becomes more evident when taking a closer look at sub-national data.
 Regional wealth disparities are of prime concern, especially when crafting policies for equitable growth [@lessmann2017].
-A country's macro-level prosperity does not automatically guarantee that all its regions partake equally in this wealth.
-By studying smaller regions within a country, it is possible to get a more nuanced narrative about the state of regional economic disparities [@lessmann2017].
-
-Using time series data and cross-sectional observations we investigate Economic growth and Inequality trends in Portugal, France, Hungary, Slovenia and Denmark in the time period 2000 to 2020.
+Furthermore, a country's macro-level prosperity does not automatically guarantee, that all its regions partake equally in this wealth.
+By studying smaller regions within a country, it is possible to get a healthier narrative about the state of regional economic disparities [@lessmann2017].
+
 The aims are to acquire, process, and analyse sub-national GDP, GINI and population data for a subset of European countries.
-
-The primary objective is to gain knowledge of econometric methods used for science in business through our given topic.
+Using regression models with time series data, cross-sectional and panel data observations we investigate economic growth and inequality trends in; Portugal, France, Hungary, Slovakia and Denmark in the time period 2000 to 2020.
+
+Moreover, the primary objective is to gain knowledge of econometric methods used for science in business through our given topic.
 Subsequently, econometric terms and definitions are explained in greater detail than it normally would in similar papers.
-Using GDP per capita and the Gini coefficient.
-Moreover, both are vital indicators in the field of economics and are frequently used, in a wide array of studies.
-Furthermore, to analyse and interpret the economic conditions and social disparities of different regions or countries.
-Their analysis provides crucial insights for policymakers, economists, and researchers in shaping and evaluating the effectiveness of socio-economic policies.
 
@@
 Setting a solid foundation for a comprehensive study that not only measure our given parameters, but also a critically examination of the role of determinants such as transportation infrastructure and education.
-@lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs as a determinant of regional inequality.
-
-The sructure of our study is methodically designed to facilitate this exploration.
+@lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs.
+
+The structure of our study is methodically designed to facilitate this exploration.
 In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data.
 In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010.
-Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality.
+Furthermore, exploring alternative functional forms and implementing panel estimation to analyse the relationship between regional development inequality.
 
 Lastly, it is worth noting potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions.
-The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017).
+The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer, and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017).
 Despite these precautions, there may be instances where incorrect information was provided, and some errors could have been overlooked.
-The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.
 
@@
 
-HER kommer BESKRIVE TRENDER
+The graph depicts a 20-year upward trend in GDP per capita across Portugal's regions, illustrating resilience and economic recovery despite various setbacks.
+Three regions have noticeable higher GDP per capita than the rest among those Alentejo Central (top, green) and Algarve (top, purple) as well as areas around the capital Lisbon (top, blue).
+The first two regions experienced significant growth, bolstered by factors like human capital development and a thriving tourism industry.
+This growth, however, was followed by a notable dip around 2008, reflecting these regions' sensitivity to global economic shifts and their reliance on tourism.
+The Metropolitan Area of Lisbon (Top Blue), while also impacted, showed less fluctuation during this period, indicative of its more diverse economic base.
+
+Post-financial crisis, there was a clear trend of recovery and return to growth across the regions, demonstrating their ability to adapt and rebound from economic challenges.
+However, the onset of the COVID-19 pandemic introduced a universal decline in all regions, underscoring the extensive impact of global health crises on regional economies.
+This pandemic-induced downturn mirrors global economic trends, where financial instability, as noted by Nguyen et al.
+(2022), often exacerbates regional economic hardships.
 

### 2023-12-06 MSB104_Submission_Group1.qmd
M

@@
 Furthermore,* regional growth and inequality have been important topics of research in Europe for many years.
-Despite the European commission\'s efforts to decrease income gaps, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
+Despite the European commission's efforts to decrease income gaps, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
 
@@
 Regional growth in Europe reveals a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
-Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by \[@gennaioli2014\] findings.
+Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by [@gennaioli2014] findings.
 
@@
 
-The effect that surprised us most was that the economic crisis led to increase inequality\[@nguyen2022\].
+The effect that surprised us most was that the economic crisis led to increase inequality[@nguyen2022].
 indicate that any type of financial crisis results in higher income inequality.
@@
 All dataset used in this paper is from Eurostat and based on the Nomenclature of Territorial Units for Statistics (NUTS) classification.
-The NUTS system provides a hierarchical classification of the EU\'s economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).
+The NUTS system provides a hierarchical classification of the EU's economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).
 
@@
 Heteroskedasticity is according to @wooldridge2020 an important factor to consider in econometrics and statistical modelling, especially in the context of regression analysis.
-Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables \[@wooldridge2020\].
+Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables [@wooldridge2020].
 In simpler terms, the spread of the residuals varies at different points in the regression model.
@@
 Furthermore, the interpretation of different patterns in the plot, like random dispersion indicating a good model fit or systematic patterns indicating issues.
-Subsiquently, ensuring robust and reliable modeling visual diagnostics, complement statistical tests \[@wooldridge2020\].
+Subsiquently, ensuring robust and reliable modeling visual diagnostics, complement statistical tests [@wooldridge2020].
 

### 2023-12-06 MSB104_Submission_Group1.qmd
mmm
hey
.

--- 2023-12-06 MSB104_Submission_Group1.qmd
@@
 The sructure of our study is methodically designed to facilitate this exploration.
-In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data.
-In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010.
-Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality.
+In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***).
+In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).
+Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).
 
@@
 
-Why some places grow and prosper is a fundamental question in the field of social science.
-This question motivated Adam Smith's \[1863\] work and has been a major influence of the *new economic geography* [@feldman2014]*.
+Why some places grow and prosper is a fundamental question in social science research.
+This question motivated Adam Smith's work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A.
+(1863).
+An inquiry into the nature and causes of the wealth of nations.
+Edinburgh: A.
+& C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*.
 Furthermore,* regional growth and inequality have been important topics of research in Europe for many years.
-Despite the European commission\'s efforts to decrease income gaps, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
+In recent decades, the European Union implemented policies aimed at reducing regional disparities and promoting economic growth across its member states.
+Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].
 
@@
 
-Regional growth in Europe reveals a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
-Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by \[@gennaioli2014\] findings.
+Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies.
+Lessmann and Seidel's research offers insights into these dynamics, highlighting the persistence of income disparities, across regions despite overarching economic advancement [@lessmann2017].
+Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014].
 
@@
 
-The 2008 financial crisis disrupted regional growth and worsened income inequalities across Europe.
+The 2008 financial crisis significantly disrupted regional growth and exacerbated income inequalities across Europe.
 Nguyen et al.
-(2022) highlight that financial instability typically leads to increased income disparity; a trend observed globally.
+highlight that financial turmoil typically leads to increased income disparity; a trend observed globally [@nguyen2022].
+
 Focusing on Europe, we expect varied regional economic performances, especially when comparing Eastern European countries such as Hungary and Slovakia to their Western counterparts.
-Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al (2014), we anticipate a potential growth acceleration after submission.
- EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis.
-This pattern possibly mirrored in the recent COVID-19 pandemic.
+Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential growth acceleration.
+In particular for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014].
+However, the financial crisis and subsequent labour market shifts may also have influenced these trends.
+
+EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic.
+In particular Portugal as the crisis hit them harder.
 These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.
@@
 
-The effect that surprised us most was that the economic crisis led to increase inequality\[@nguyen2022\].
-indicate that any type of financial crisis results in higher income inequality.
-Furthermore,  @lessmann2017 address the importance of studying these aspects.
-Moreover, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
+Another affect that surprised us was that the economic crisis led to increase inequality.
+The study "*the effects of financial crisis on regional inequality*" indicate any type of financial crisis results in higher income inequality [@nguyen2022].
+
+@lessmann2017 regional inequality study address the importance of studying these aspects.
+Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].
+The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.
 
@@
 
-Geographic characteristics stand out as fundamental, with the natural and topographical elements of a region shifting income distributions [@lessmann2017].
-Urbanization emerges as a crucial determinant, where active, populations around urban centres usually enjoy raised income levels, contrasting with the disparities witnessed in more rural and secluded areas [@lessmann2017].
+Geographic characteristics stand out as pivotal, with the natural and topographical elements of a region significantly swaying income distributions [@lessmann2017].
+
+Urbanization emerges as a crucial determinant, where bustling, populous around urban centers usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].
+
+*Transport infrastructure and Education*
 
@@
 The study mentions that human capital is the most important determinant of differences in regional development within countries.
-Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
+The study also notes that human capital creates positive externalizes, which can spill over regional boundaries.
+Moreover, the quality of human capital within countries, as measured by the secondary-school enrollment rate, can decrease regional inequality by facilitating regional spillovers and convergence.
 Lastly, promoting internal migration.
@@
 
-@lessmann2017 provides insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
+Furthermore, @lessmann2017 provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017].
 The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.
-Additionally, there is an N-shaped relationship between regional inequality and economic development according to \@lessmann2017.
-Subsequently, regional inequality increases again after the inverted U-shaped pattern has been completed [@lessmann2017].
-
-Lastly, [@iammarino2019] argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
+
+@lessmann2017 paper also finds that there is an N-shaped relationship between regional inequality and economic development, which means that regional inequality increases again after the inverted U-shaped pattern has been completed [@lessmann2017].
+
+@iammarino2019 discusses regional inequalities in European countries.
+It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium.
+The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019].
+Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019].
 On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.
 
-Overall, it is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by several factors, with transport infrastructure and education emerging as significant determinants.
-Earlier studies provide us with a fundamental understanding of these dynamics, while also highlighting the impacts of economic crises on regional disparities.
-The integration of transport infrastructure and education adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
-
-# Data and Methodology
+Overall, it is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure and education emerging as significant determinants.
+Earlier studies provide us with a fundamental understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities.
+The integration of transport infrastructure and education into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.
+
+# Data and Descriptive
 
@@
 #| echo: false
-#| output: false
 
@@
 #| echo: false
-#| output: false
+#| print: false
 
@@
 ggplot(gini_filtered, aes(x = id_nuts0, y = GDP_per_capita)) +
-  geom_boxplot() +
-  ggtitle("Plot 1: Distribution of GDP per Capita by Country") +
-  xlab("Country") +
-  ylab("GDP per Capita")
-
-# Saving the plot for referencing in the appendix
-g <- ggplot(gini_filtered, aes(x = id_nuts0, y = GDP_per_capita)) +
   geom_boxplot() +
@@
   ylab("GDP per Capita")
-
-# Save the plot as an image file
-ggsave("GDP_per_Capita_Boxplot.png", plot = g, width = 10, height = 6)
-```
-
-The box plots (in the appendix) provide a distinct perspective on the data, independent of the assumptions typically associated with other statistical distributions.
-This method of visualization highlights Denmark as having the highest overall GDP per capita across its regions.
-France, while not having the overall highest GDP per capita, demonstrates significant variability, especially when considering its outliers.
-Hungary, in contrast, is still presenting the lowest GDP per capita.
-These insights, derived from the box plot analysis, augment our understanding from previous descriptive statistics, bar plots, and time series analyses, offering a more nuanced view of the GDP per capita distribution among these countries.
-
-# Empirical Findings
-
-## Cross sectional estimates
+```
+
+# Cross sectional estimates
 
 ## Part 2A: Growth and Inequity
+
+### **Data acquisition**
+
+Firstly, the data is summerised for the year 2010 at the NUTS2 regional level, focusing on economic metrics.
+Processing the dataset containing the NUTS3 level.
+Thereafter, computing the total GDP and total population for each NUTS2 region and year.
+Then, derive the GDP per capita at the NUTS2 level by dividing the total GDP by the total population again.
+Furthermore, narrowing down the dataset to observations from the year 2010 where the Gini coefficient is positive.
+Then adding, logarithmic transformations to linearize relationships or to reduce the impact of extreme values.
+
+The next step is making a data frame of NUTS2 2010 and grouping by NUTS0 country level variables, as well as allocating NUTS2 to selected countries.
+Lastly presenting data in **table XXX** showing a count of NUTS2 regions within each top-level region or country in the 2010 data.
+
+The next step is, estimating a Linear Regression for All Countries using cross sectional data from 2010, NUTS2.
 
@@
 
-The goodness of fit in simple linear regression, evaluated by R\^2, gauges how well the regression line fits the data.
-This line is determined using the least squares criterion, minimizing squared differences between data points and the line.
-R\^2 varies from 0 to 1, where 0 indicates no explanatory power and 1 signifies perfect fit.
-A higher R\^2 suggests a better model fit.
-For an effective model, data should linearly align with minimal patterns in residuals, assuming all linear regression prerequisites are met.
+The goodness of fit in a simple linear regression model measures how well the regression line approximates the real data points.
+The regression line that best represents the data according to the least squares criterion, which minimizes the sum of the squared vertical distances of the points from the line.
+
+R\^2 is a statistical measure that represents the proportion of the variance for the dependent variable that's explained by the independent variable.
+It ranges from 0 to 1.
+A higher R\^2 value indicates a better fit of the model to the data.
+
+R\^2 = 0 The model does not explain any of the variability of the response data around its mean.
+
+R\^2 = 1 The model explains all the variability of the response data around its mean.
+
+If the assumptions for a simple linear regression are met, it indicates a good fit for the model.
+[Plotting the data can provide a visual indication of the goodness of fit.]{.underline} The points should fall around a straight line without clear patterns in the residuals.
+As it is crucial to verify these assumptions before proceeding with interpreting the results of the regression analysis.
 
 **Residuals vs.
-Fitted Values Plot** is used to check for homoscedasticity and linearity.
+Fitted Values Plot** to check for homoscedasticity and linearity.
 Ideally, this plot shows no pattern; the residuals are randomly scattered around the horizontal line at zero.
@@
 #| echo: false
-#| output: false
-
-# Made a copy for appendix
+#| fig-cap: Regression statistics of all countries for the year 2010
+
 # Finding the natural logarithm of GDP per capita
@@
 #| echo: false
-#| output: false
 #| fig-cap: Regression statistics of all countries seperately for the year 2010
 
-#Appendix 
 # Estimation for each country seperately 
@@
 
-The OLS method is according to @wooldridge2020 employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
-Furthermore, this technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
-The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the dataset, without necessitating iterative procedures.
-According to @wooldridge2020 the following assumtions need to be met to achieve BLUE.
-
-1\. ***Linear in coefficients and error term***: The model is linear in its terms and the error term.
-
-2\. ***Error term has zero mean***: The error term averages zero, ensuring unbiasedness.
-
-3\. ***No correlation between independent variables and error term***: Independent variables should not predict the error term.
-
-4\. ***Observations of error term are uncorrelated*****:** Each error term observation doesn't predict the next.
-
-5\. ***Homoscedasticity*****:** The error term's variance is consistent across observations.
-
-6\. ***No perfect multicollinearity***: No independent variable is a perfect linear function of others.
-
-7\. ***Normally distributed error term (Optional*****)**: This aids in statistical testing, although not required for unbiased OLS estimates
-
-```{r}
-#| echo: false
-#| output: false
+"The OLS method is employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals.
+This technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values.
+The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the data-set, without necessitating iterative procedures.
+
+Furthermore, when the classical linear regression assumptions are met, OLS guarantees that the estimators are BLUE, ensuring their unbiasedness and efficiency.
+This is particularly crucial in econometric analysis, where the precision and reliability of parameter estimates are paramount for policy implications and economic interpretations.
+The foundation assumptions of linear regression, including linearity, independence, homoscedasticity, and normality of residuals, are prerequisites for OLS to attain these desirable properties.
+It is imperative, therefore, to conduct diagnostic tests and assess the validity of these assumptions to ensure the robustness of the OLS estimators used in our analysis."
+
+### **Assumptions**
+
+**I. Linearity**
+
+The relationship between the independent variable X and the dependent variable Y is linear.
+This means that changes in X are associated with proportional changes in Y.
+A straight line should provide a good fit to the data points when plotted on a graph.
+
+You can create a scatter plot of Y versus X and visually inspect whether a straight line could well represent the relationship.
+Alternatively, you can plot the residuals versus the fitted values and check for any obvious patterns or non-linearity.
+
+**II. Independence:**
+
+Observations are independent of each other. The value of Y for one observation should not depend on the value of Y for any other observation.
+
+This assumption is more about study design and data collection.
+Ensure that your observations are not correlated in time or space.
+For example, if you're analyzing economic data across regions, make sure that the regions are not influencing each other.
+
+**III. Homoscedasticity (Equal Variance):**
+
+The variance of the residuals (the errors) is constant across all levels of X.
+This means that the spread or "width" of the residuals should remain roughly the same across all values of the independent variable.
+
+A plot of residuals versus fitted values should show a random scatter and not display any funnel-like shapes (wider at one end).
+
+**IV. Normality of Residuals:**
+
+The residuals (the differences between observed and predicted values) are normally distributed. This assumption is particularly important for hypothesis testing and creating confidence intervals.
+
+A Quantile-Quantile (Q-Q) plot of the residuals can show if they follow a normal distribution. The points should fall roughly along a straight line.
+
+**V. No Perfect Multicollinearity (Specific to Multiple Regression)**
+
+In multiple regression settings, this assumes that no independent variable is a perfect linear function of any other independent variables.
+While this is more pertinent to multiple linear regression, it is crucial there because high correlation between independent variables can lead to unstable coefficient estimates.
+
+Checking the variance inflation factor (VIF) for each variable; a VIF above 5-10 indicates a problematic amount of collinearity.
+
+**Visualization**
+
+```{r}
+#| echo: false
 
 # Making two plots, first one is a residuals vs fitted plot, while the other is a normal Q-Q residuals plot
-#plot(lm, which = 1:2) 
-```
-
-The residuals vs fitted values and the Q-plot, combined with the regression statistics may help us find out if classical OLS assumptions are met.
-The first help us check for homoscedasticity assumption of a linear regression model.
-The residuals should be randomly scattered around the horizontal 0 line, something that can indicate that the variances of the error terms are constant.
+plot(lm, which = 1:2) 
+```
+
+We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient, and that together with the regression statistics can help with discussing if the classical OLS assumptions hold for the model.
+The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model.
+The residuals should be randomly scattered around the horisontal 0 line, something that can inducate that the variances of the error terms are constant.
 In our plot, the residuals are in some extent randomly distributed, and there is also no clear pattern; this suggest that there is likely no significant issues with heteroscedasticity or non-linearity.
-The latter is used to assess if the residuals of the linear model are normally distributed.
+
+The other plot - normal Q-Q plot - is used to assess if the residuals of the linear model are normally distributed.
 In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed.
 There are however some outliers in the tails, that suggest some variation from normality.
-Both plots are presented in appendix.
-
-```{r}
-#| echo: false
-#| output: false
+
+In both of these plots, we can see that there are some outliers, these can affect the fit of the model.
+These outliers might come from regions that have a high GDP per capita compared to rest of the regions, like Paris in France that is a financial hub.
+
+```{r}
+#| echo: false
 #| label: fig_scatterplot-all
@@
 
-### Part 2B: Exploring Other Determinants of Inequity
-
-#### I. Data Acquisition
+## Part 2B: Exploring Other Determinants of Inequity
+
+### I. Data Acquisition
 
@@
 
-#### II. Multiple Linear Regression Model
+### II. Multiple Linear Regression Model
 
@@
 #| echo: false
-#| output: false
 #| fig-cap: Multiple Linear Regression Model
@@
 
-#### II. Model specification
+### Model specification
 
@@
 
-#### III. Model Interpretation
+### III. Model Interpretation
 
@@
 
-## Alternative Functional Forms and Panel Estimates
-
-### Testing Development Effects Across Subsets
-
-#### Subset Analysis
+# Assignment 3: Alternative Functional Forms and Panel Estimates
+
+## Testing Development Effects Across Subsets
+
+### Subset Analysis
 
@@
 
-#### Subset Analysis Discussion
+### Subset Analysis Discussion
 
@@
 
-#### Results Interpretation
+### Results Interpretation
 
@@
 
-### Heteroskedasticity Testing and Causality Discussion
-
-#### Heteroskedasticity Testing
+## Heteroskedasticity Testing and Causality Discussion
+
+### Heteroskedasticity Testing
 
 Heteroskedasticity is according to @wooldridge2020 an important factor to consider in econometrics and statistical modelling, especially in the context of regression analysis.
-Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables \[@wooldridge2020\].
+Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables [@wooldridge2020].
 In simpler terms, the spread of the residuals varies at different points in the regression model.
@@
 
-# Discussion
-
-# Limitations and Future Research
-
-# Conclusion
+# References
 
 # Appendix
-
-1.  **Descriptive statistics**
 
@@
 ```
-
-2.  **Boxplot**
-
-![](GDP_pc_Boxplot.png)
-
-3.  **Regression statistics of all countries separately for the year 2010**
-
-```{r}
-#| echo: false
-#| fig-cap: Regression statistics of all countries for the year 2010
-
-cat("Intercept (Coefficient 1):", b0_earn, "\n")
-cat("Slope of log GDP per capita (Coefficient 2):", b1_earn, "\n")
-```
-
-![](images/Regression_all_countries2010-01.png)
-
-**Testing for OLS assumtions**
-
-```{r}
-#| echo: false
-#|
-
-# Making two plots, first one is a residuals vs fitted plot, while the other is a normal Q-Q residuals plot
-plot(lm, which = 1:2) 
-```
-
-# References

### 2023-12-07 MSB104_Submission_Group1.qmd

--- 2023-12-06 MSB104_Submission_Group1.qmd
@@
 
-**Expected Inverted U- and N shaped curves**
+**Expected Inverted U- and N shaped curves (er dette riktig?!?, se tilbakemelding)**
 
@@
 #| echo: false
+#| label: fig-1
+#| fig-cap: "GDP per capita Portugal"
 
@@
 #| echo: false
+#| #| label: fig-2
+#| fig-cap: "GDP per capita France"
 
@@
 #| echo: false
+#| #| label: fig-3
+#| fig-cap: "GDP per capita Hungary"
 
@@
 #| echo: false
-#| output: false
+#| label: fig-4
+#| fig-cap: "GDP per capita Slovakia"
 
@@
 #| echo: false
+#| label: fig-5
+#| fig-cap: "GDP per capita Denmark"
 
@@
 #| echo: false
+#| label: fig-6
+#| fig-cap: "Bar chart"
 
@@
 #| echo: false
+#| label: fig-7
+#| fig-cap: "Time-series plot"
 
@@
 
-```{r}
-#| echo: false
-#| output: false
+Ska me ha med denne? er det egentlig nødvendig??
+
+```{r}
+#| echo: false
+#| output: false
+#| label: fig-8
+#| fig-cap: "Box plot"
 
@@
 
+bør kanskje skrive noe mer om hva gini er, tror det står på side 116 eller 106 i artikkel
+
 # Empirical Findings
@@
 #| output: false
+#| label: tbl-1
+#| fig-cap: "Regression analysis all countries combined for the year 2010"
 
@@
 #| output: false
-#| fig-cap: Regression statistics of all countries seperately for the year 2010
+#| label: tbl-2
+#| fig-cap: "Regression statistics of all countries seperately for the year 2010"
 
@@
 #| output: false
+#| label: fig-9
+#| fig-cap: "Residuals vs fitted plot and normal Q-Q residuals plot"
 
@@
 #| output: false
-#| label: fig_scatterplot-all
+#| label: fig-10
 #| cap-location: bottom
@@
 In this plot we are visualizing the relationship between GDP per capita and gini by using a mix of the two previous plots. We can here as well, see extreme outliers.
+
+MÅ SKRIVE MER HER!!
 
@@
 #| output: false
-#| fig-cap: Multiple Linear Regression Model
+#| label: tbl-3
+#| fig-cap: "Multiple Linear Regression Model"
 
@@
 #| echo: false
+#| #| label: tbl-4
+#| fig-cap: "Regression analysis France vs other countries"
 
@@
 #| echo: false
+#| label: tbl-5
+#| fig-cap: "Linear, logarithmic and quadratic model"
 
@@
 #| echo: false
+#| label: fig-11
+#| fig-cap: "Figure of logarithmic model for France: Predicted vs actual gini coefficients"
 
@@
 #| echo: false
+#| label: fig-12
+#| fig-cap: "Figure of logarithmic model for other countries: Predicted vs actual gini coefficients"
 
@@
 #| echo: false
+#| label: fig-13
+#| fig-cap: "Figure of quuadratic model for France: Predicted vs actual gini coefficients"
 
@@
 #| echo: false
+#| label: fig-14
+#| fig-cap: "Figure of quuadratic model for the other countries: Predicted vs actual gini coefficients"
 
@@
 
+MÅ FIKSE PÅ NOE MED N OG U SHAPE (SE TILBAKEMELDING)
+
 ### Heteroskedasticity Testing and Causality Discussion
@@
 #| echo: false
+#| label: fig-15
+#| fig-cap: "Residuals vs fitted values for linear model, all countries combined"
 
@@
 #| output: false
-
-# Panel estiamtion
-
-plm1 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("id_nuts0"), model = "within",effect = "twoway") |> 
+# filtering data 
+fgini_nuts_2 <- gini_nuts_2 |> 
+  filter(!is.na(population),
+         population != 0,
+         GDP_per_capita != 0, 
+         GDP_per_capita != "NA"
+  ) 
+```
+
+```{r}
+#| echo: false
+#| output: false
+#| label: tbl-6
+#| fig-cap: "Panel estimations"
+
+# Panel estimations
+
+plm1 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("id_nuts0"), model = "within",effect = "twoway") |> 
   coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
 
-plm2 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("year"), model = "within",effect = "twoway") |> 
+plm2 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("year"), model = "within",effect = "twoway") |> 
   coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
 
-plm3 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("id_nuts2"), model = "within",effect = "twoway") |> 
+plm3 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("id_nuts2"), model = "within",effect = "twoway") |> 
   coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
 
-plm4 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("year","id_nuts2"), model = "within",effect = "twoway") |> 
+plm4 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("year","id_nuts2"), model = "within",effect = "twoway") |> 
   coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)
@@
 
+By looking at the results of the panel estimation, we can see a distinct relationship between GDP per capita and regional inequality, with a significant variability observed across the different models.
+Model 1 and 2 shows a significant positive correlation between GDP per capita and regional inequality, which suggests that higher economic output per capita is associated with greater inequality.
+The negatice coefficients for the log of GDP per capita imply that relative increases in economic prosperity could lead to an reduction in inequality; however, models 3 and 4 show that this association becomes less strong or irrelevant, suggesting that the influence of economic forces on inequality depends on regional or temporal characteristics.
+
+These variations underscore the influence of fixed effects, showing that country-specific, temporal and regional characteristics can alter our understanding of economic disparity.
+When we compare it to cross-sectional estimates, we can see the panel data's ability to caputre temporal changes and control for heterogenity, offers a more comprenehsive perspective.
+This is in contrast with the static nature of cross-sectional analysis, which might overlook the complexities and nature of economic inequality.
+The panel approach provides deeper insights into the development of regional disparities and factors influencing them.
+
 ### Panel Estimation Discussion
+
+The findings from the panel estimation offer insightful revelations about the relationship between GDP per capita and regional inequality; but we also have to look at potential methodological limitations and biases.
+The models sensitivity to different fixed effects suggests a nuanced, but potentially inconsistent understanding of inequality.
+This variability reflect the complexity of economic factors at play, and raises concerns about omitted variable bias, where key inlfuencers like industrial composition or demographic changes might be missing.
+
+The inherent challenges in accurately measuring GDP per capita and inequality could introduce measurement errors, affecting the reliability of the results.
+While panel data provides a dynamic view over cross-sectional data, establishing causality remains elusive, and the correlations observed should be interpreted with caution.
+The study's generalizability is also limited with its focus on a specific set of countries and times, which may not represent broader global dynamics.
+There is a low R-squared in some of the models, something that might've occured due to unexplained variances, suggesting that other significant, but unaccounted factors might be influencing regional inequality.
+Additionally, the economic significance on the finding needs careful consideration in order to understand the-world impact of these economic relationships.
+Overall the findings should be interpreted within the context of these potential biases and limitations, and they underscore the complexity of modeling and understanding economic phenomenas such as regional inequality.
 
